---
output: github_document
---

<!-- README.md is generated from README.Rmd by devtools::build_readme(). Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BetaDivMultifun

<!-- badges: start -->
<!-- badges: end -->

The goal of BetaDivMultifun is to ...

## Installation

For installation, please consider the vignette `how-to-use-this-package.Rmd`.

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("allanecology/BetaDivMultifun")
```

## Content

- The scripts you need are in the folder vignettes

#TODO : 
- add a content description (TOC)
- add the script overview image

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = F}
library(BetaDivMultifun)
## basic example code
```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
