---
title: "prepare functions"
author: "noelle"
date: "November 14, 2018"
output: html_document
---


# Loading file paths
The paths of file locations are stored externally in "info_data.csv" where all information is stored.
Load this file externally and store in variable "info_data". More information is given in "nonpublic.R", 
at server where the whole analysis is stored.

*note* : to read in data, you need to be connected to server. 

*note2* : if info_data.xls is changed, it needs to be saved as .csv before the changes apply here!

```{r}
# only show lines which contain path info
info_data <- info_data[grep("txt", info_data$path_unix),c("ColumnName", "path_unix")]
```


# Plot, exploratories
```{r}
PlotExplo <- data.table::fread(info_data$path_unix[grep("Biomass2011", info_data$ColumnName)], header=T, sep='\t')
PlotExplo <- PlotExplo[,c("EpPlotID", "explo")]
data.table::setnames(PlotExplo, old=c("EpPlotID", "explo"), new=c("Plot", "Explo"))
# generate old plot names
Plot <- c(paste("AEG", seq(1,50), sep=""), paste("HEG", seq(1,50), sep=""), paste("SEG", seq(1,50), sep=""))
# generate new plot names
Plotn <- c(paste("AEG", sprintf("%02d", 1:50), sep=""), paste("HEG", sprintf("%02d", 1:50), sep=""), paste("SEG", sprintf("%02d", 1:50), sep=""))
PlotExplo <- merge(PlotExplo, cbind(Plot, Plotn), by = "Plot")

rm(Plot); rm(Plotn)
```

# raw dataset
Generate a dataset which stores all raw data. Will be assembled with the script `2calc_raw_dataset.R`. Datasets that can not be easily included as raw datasets into `raw_functions` are assembled (calculations are done) in `2calc_raw_dataset.R` and then integrated into raw_functions as well (and assembled functions variable, of corse).
```{r}
# subset for raw dataset construction
raw_functions <- data.table::copy(PlotExplo)
# generate list for datasets that will be constructed in 2calc_raw_dataset.R script directly.
list_raw_functions <- list()
```

# Temperature and moisture
```{r}
#TODO where is this dataset?
```


# Biomass
years 2008 to 2017. **caution**, 2008 might not be reliable.

all biomasses in g/m^2

2009 : Biomass was clipped in each two subplots per plot. In each subplot we clipped four 50 x 50 cm plots 2-3 cm above-ground.
two treatments, K1 and K2, K2 is taken in synthesis. 149 plots only.
```{r}
#TODO check if units are the same in each year

# 2008
Biomass2008 <- data.table::fread(info_data$path_unix[grep("Biomass2008", info_data$ColumnName)], header=T, sep='\t')
Biomass2008 <- Biomass2008[, c("PlotID", "biomass_g_m_2")]
data.table::setnames(Biomass2008, old= c("PlotID", "biomass_g_m_2"), new = c("Plot", "Biomass2008"))
raw_functions <- merge(raw_functions, Biomass2008, by="Plot", all=T)
rm(Biomass2008)

# 2009
Biomass2009 <- data.table::fread(info_data$path_unix[grep("Biomass2009", info_data$ColumnName)], header=T, sep='\t')
Biomass2009 <- Biomass2009[Treatment == "K2",]
Biomass2009 <- Biomass2009[,c("GpPlotID", "Biomass")]
data.table::setnames(Biomass2009, old=c("GpPlotID", "Biomass"), new=c("Plot", "Biomass2009"))
# raw_functions <- merge(raw_functions, Biomass2009, by="Plot", all.x=T)
# list_raw_functions[["Biomass2009"]] <- Biomass2009
# TODO there is something wrong with Biomass2009. is it same area as other years?
#TODO linear model day vs. biomass and take fitted values

# 2010
Biomass2010 <- data.table::fread(info_data$path_unix[grep("Biomass2010", info_data$ColumnName)], header=T, sep='\t')
Biomass2010 <- Biomass2010[, c("PlotID", "Biomasse")]
data.table::setnames(Biomass2010, old= c("PlotID", "Biomasse"), new = c("Plot", "Biomass2010"))
raw_functions <- merge(raw_functions, Biomass2010, by="Plot", all=T)
rm(Biomass2010)

# 2011
Biomass2011 <- data.table::fread(info_data$path_unix[grep("Biomass2011", info_data$ColumnName)], header=T, sep='\t')
Biomass2011 <- Biomass2011[,c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2011, old = c("EpPlotID", "biomass_g"), new = c("Plot", "Biomass2011"))
raw_functions <- merge(raw_functions, Biomass2011, by="Plot", all=T)
rm(Biomass2011)

# 2012
Biomass2012 <- data.table::fread(info_data$path_unix[grep("Biomass2012", info_data$ColumnName)], header=T, sep='\t')
Biomass2012 <- Biomass2012[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2012, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2012"))
raw_functions <- merge(raw_functions, Biomass2012, by="Plot", all=T)
rm(Biomass2012)

# 2013
Biomass2013 <- data.table::fread(info_data$path_unix[grep("Biomass2013", info_data$ColumnName)], header=T, sep='\t')
Biomass2013 <- Biomass2013[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2013, old= c("EpPlotID", "biomass_g"), new= c("Plot", "Biomass2013"))
raw_functions <- merge(raw_functions, Biomass2013, by="Plot", all=T); rm(Biomass2013)

# 2014
Biomass2014 <- data.table::fread(info_data$path_unix[grep("Biomass2014", info_data$ColumnName)], header=T)
Biomass2014 <- Biomass2014[,c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2014, old = c("EpPlotID", "biomass_g"), new = c("Plot", "Biomass2014"))
raw_functions <- merge(raw_functions, Biomass2014, by="Plot", all=T); rm(Biomass2014)

# 2015
Biomass2015 <- data.table::fread(info_data$path_unix[grep("Biomass2015", info_data$ColumnName)], header=T, quote="", drop=19)
Biomass2015 <- Biomass2015[,c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2015, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2015"))
raw_functions <- merge(raw_functions, Biomass2015, by="Plot", all=T); rm(Biomass2015)

# 2016
Biomass2016 <- data.table::fread(info_data$path_unix[grep("Biomass2016", info_data$ColumnName)], header=T)
Biomass2016 <- Biomass2016[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2016, old= c("EpPlotID", "biomass_g"), new= c("Plot", "Biomass2016"))
raw_functions <- merge(raw_functions, Biomass2016, by="Plot", all=T); rm(Biomass2016)

# 2017
Biomass2017 <- data.table::fread(info_data$path_unix[grep("Biomass2017", info_data$ColumnName)], header=T)
Biomass2017 <- Biomass2017[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2017, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2017"))
raw_functions <- merge(raw_functions, Biomass2017, by="Plot", all=T); rm(Biomass2017)
```



# Soil Organic C
2011
```{r}
SoilOrganicC <- data.table::fread(info_data$path_unix[grep("SoilOrganicC", info_data$ColumnName)], header=T, sep='\t')
SoilOrganicC <- SoilOrganicC[which(SoilOrganicC$Type == "G"), c("EP_Plotid","Organic_C")]
data.table::setnames(SoilOrganicC, old=c("EP_Plotid", "Organic_C"), new=c("Plot", "SoilOrganicC"))
raw_functions <- merge(raw_functions, SoilOrganicC, by="Plot", all=T); rm(SoilOrganicC)
```


# Root biomass
2011, 
Includes fine and coarse biomass, will add them together.
```{r}
Root.biomass <- data.table::fread(info_data$path_unix[grep("Root.biomass", info_data$ColumnName)], header=T, sep='\t')
Root.biomass <- Root.biomass[Type == "G"]
Root.biomass <- Root.biomass[,c("EP_Plotid", "Fine_Roots_Biomass", "Coarse_Roots_Biomass")]
data.table::setnames(Root.biomass, old="EP_Plotid", new="Plot")
list_raw_functions[["Root.biomass"]] <- Root.biomass ; rm(Root.biomass)
```

# Flower cover
**leave out**
We can go from "total_influorescence_area_in_cm" of each species to a total flower cover. The total area measured was : 600m² per grassland plot.
TODO : left open as not sure if I include it
```{r, eval=F}
Total.flower.cover <- data.table::fread(info_data$path_unix[grep("Total.flower.cover", info_data$ColumnName)], header=T, sep='\t')
```

# PRI - Phosphorous retention index
## Microbial phosphorous in soil
2011 and 2014

(mg/kg)

used to calculate PRI
```{r}
Pmic.2011 <- data.table::fread(info_data$path_unix[grep("Pmic.2011", info_data$ColumnName)], header=T, sep='\t')
Pmic.2011 <- Pmic.2011[grep("G", Pmic.2011$EP)]
data.table::setnames(Pmic.2011, old=c("EP", "Pmic"), new=c("Plot", "Pmic.2011"))
raw_functions <- merge(raw_functions, Pmic.2011, by="Plot", all=T); rm(Pmic.2011)
```

```{r}
Pmic.2014 <- data.table::fread(info_data$path_unix[grep("Pmic.2014", info_data$ColumnName)], header=T, sep='\t')
Pmic.2014 <- Pmic.2014[,c("EP", "Pmic")]
data.table::setnames(Pmic.2014, old=c("EP", "Pmic"), new=c("Plot", "Pmic.2014"))
raw_functions <- merge(raw_functions, Pmic.2014, by="Plot", all=T) ; rm(Pmic.2014)
```

## Soil Phosphorous fractions
used to calculate PRI

years : 2008, 2009

NaHCO3_Pi is used
```{r}
np <- data.table::fread(info_data$path_unix[grep("NaHCO3.Pi", info_data$ColumnName)], header=T, sep='\t') 
np <- np[grep("G", np$EP)]
np <- np[, c("EP", "NaHCO3_Pi")]
data.table::setnames(np, old="EP", new="Plot")
raw_functions <- merge(raw_functions, np, by="Plot", all=T) ; rm(np)
```


## Plant Nutrients (Pshoot)
2009, 2010, 2011, 2012, 2013

contains: Nshoot and Pshoot of all years

near infrared spectroscopy (NIRS), 

note : there are **values for 2008** in dataset 5401 (used for Biomass2008)
```{r}
nutrients <- data.table::fread(info_data$path_unix[grep("Nshoot.2009", info_data$ColumnName)])
# 2009
nutrients2009 <- nutrients[year == "2009", c("Plot", "N", "P")]
data.table::setnames(nutrients2009, old=c("N", "P"), new=c("Nshoot2009percent", "Pshoot2009percent"))
# 2010
nutrients2010 <- nutrients[year == "2010", c("Plot", "N", "P")]
data.table::setnames(nutrients2010, old=c("N", "P"), new=c("Nshoot2010percent", "Pshoot2010percent"))
# 2011
nutrients2011 <- nutrients[year == "2011", c("Plot", "N", "P")]
data.table::setnames(nutrients2011, old=c("N", "P"), new=c("Nshoot2011percent", "Pshoot2011percent"))
# 2012
nutrients2012 <- nutrients[year == "2012", c("Plot", "N", "P")]
data.table::setnames(nutrients2012, old=c("N", "P"), new=c("Nshoot2012percent", "Pshoot2012percent"))
# 2013
nutrients2013 <- nutrients[year == "2013", c("Plot", "N", "P")]
data.table::setnames(nutrients2013, old=c("N", "P"), new=c("Nshoot2013percent", "Pshoot2013percent"))
# take nutrients together for list_raw_functions
nutrients <- Reduce(function(x, y) merge(x, y, all=TRUE), list(nutrients2009, nutrients2010, nutrients2011, nutrients2012, nutrients2013))
list_raw_functions[["nutrients"]] <- nutrients
rm(nutrients); rm(nutrients2009); rm(nutrients2010); rm(nutrients2011); rm(nutrients2012); rm(nutrients2013)
```



# Root decomposition
Decomposition measurements of fine roots (<2mm) within the upper 10 cm of the mineral soil. Determination of mass loss after 6 months. Litterbag method. Standardized control material: beech roots for forest sites and herbaceous roots for grassland sites. Numbers in [%]

mass "loss" because decomposed roots are lost.
```{r}
Root.decomposition <- data.table::fread(info_data$path_unix[grep("Root.decomposition", info_data$ColumnName)], header=T, sep='\t')
Root.decomposition <- Root.decomposition[Type == "G"]
Root.decomposition <- data.table::setnames(Root.decomposition[,c("EP_Plotid", "Mass_loss_October_2012")], old=c("EP_Plotid", "Mass_loss_October_2012"), new=c("Plot", "Root.decomposition"))
raw_functions <- merge(raw_functions, Root.decomposition, by="Plot", all=T); rm(Root.decomposition)
```


# Bulk density and soil C stock
**Bulk density** : Increased use of machine compact the soil. At a given point, the root growth is impaired.

**soil C stock** : C in soil

Determination bulk density and calculated CN stocks for the upper 10 cm of the mineral soil. Dry combustion of soil at a temperature of 1100°C and subsequent determination of evolving CO2 and N2 with a Thermal Conductivity Detector (TCD).

[g/cm^3]

**data is from 2011, not 2014!**

```{r}
Bulk.density.soil.C.stock.2011 <- data.table::fread(info_data$path_unix[grep("Soil.C.stock", info_data$ColumnName)], header=T, sep='\t')
Bulk.density.soil.C.stock.2011 <- Bulk.density.soil.C.stock.2011[Type == "G"]
Bulk.density.soil.C.stock.2011 <- data.table::setnames(Bulk.density.soil.C.stock.2011[,c("EP_Plotid", "BD", "OC_stock")], old=c("EP_Plotid", "BD", "OC_stock"), new=c("Plot", "Bulk.density2011", "Soil.C.stock"))
raw_functions <- merge(raw_functions, Bulk.density.soil.C.stock.2011, by="Plot", all=T); rm(Bulk.density.soil.C.stock.2011)
```

# Herbivory

## 2013
invertebrate herbivory in 146 managed temperate grasslands. Year: 2013

increased land-use - less invertebrates - less herbivory

Visually estimation of herbivory by eye using templates and measurement of leaf area left after feeding of the herbivores using LI-COR area meter

2 samples per plot in 145 cases, 1 sample for :  "AEG6"  "HEG16" "HEG43" "SEG19" "SEG33" and 3 samples for "SEG45". 2 Columns are included in the raw_functions dataset, 1 for each sample.

SEG45 : The mean of all 3 samples was already taken and stored as sample2. Taking the mean of samples2 first and then with sample1 would lead to different results than in the synthesis dataset.
```{r}
herb <- data.table::fread(info_data$path_unix[grep("herbivory.2013", info_data$ColumnName)], header=T, sep='\t')
data.table::setorder(herb, EP_Plotid)
list_raw_functions[["Herbivory.2013"]] <- herb ; rm(herb)
```

## 2017 and 2018
plant material was collected before the first mowing (May 2017/2018). For each plant, we visually estimated the area damaged by herbivorous insects and gall mites on 12 - 200 leaves (depending on leaf size) and measured total leaf area using a leaf area meter. The deduced herbivory rates (% damaged area) per plant species were than summarised to community-level herbivory rates based on the respective plant cover values in the vegetation records of the survey years (2017 for AEG, AEW, HEW, SEW; 2018 for HEG, SEG from data set 24247 (pathogen infection)).

*Noelle :* 
prop_a_herb : proportion of leaf area damaged of the 10 (8) most abundant species. Serves as estimate for the whole plot and can therefore be taken directly.

AEG plots were measured in 2017, SEG/HEG in 2018. 49 AEG plots, 48 HEG plots and 50 SEG plots.

```{r}
herb <- data.table::fread(info_data$path_unix[grep("herbivory.20172018", info_data$ColumnName)], header=T, sep='\t')
data.table::setnames(herb, old=c("plotid_withzero", "plotid"), new=c("Plotn", "Plot"))
herb <- herb[grep("G", herb$Plotn)]
data.table::setorder(herb, Plotn)
raw_functions <- merge(raw_functions, herb[,c("Plot", "prop_a_herb")], by="Plot", all=T); rm(herb)
```



# Forage quality - NOT READY YET
"Futter qualität", year : 2009
forage quality (index based on crude protein concentration and relative forage value) - Eric Ecology Letters LUI

RFV : Relative Feed Value
RFQ : Relative Forage Quality
ADL : acid detergent lignin, Lignin
ADF : Acid Detergent Fibre, Säure-Detergenz-Faser, ein Teil der Gerüstsubstanzen von Futtermitteln
NDF : Neutral Detergent Fibre, Neutral-Detergenz-Faser, fasrige Anteile von Futtermitteln 

more ADF/ NDF --> less nutritional value because less easy to digest

according to forage quality estimation based on alfalfa (medicago sativa) values. [source](https://openprairie.sdstate.edu/cgi/viewcontent.cgi?referer=https://scholar.google.ch/&httpsredir=1&article=1351&context=extension_extra)

[% concentration in aboveground grassland plant community biomass without litter]
```{r, eval=F}
forage.quality <- data.table::fread(info_data$path_unix[grep("forage.quality", info_data$ColumnName)], header=T, sep='\t')
# 2009
forage.quality.2009 <- forage.quality[year == "2009"]
forage.quality.2009 <- forage.quality.2009[, c("Plot", "N", "ADF", "NDF")]
data.table::setnames(forage.quality.2009, old=c("N", "ADF", "NDF"), new = c("N.2009", "ADF.2009", "NDF.2009"))
# 2010
forage.quality.2010 <- forage.quality[year == "2010"]
forage.quality.2010 <- forage.quality.2010[, c("Plot", "N", "ADF", "NDF")]
data.table::setnames(forage.quality.2010, old=c("N", "ADF", "NDF"), new = c("N.2010", "ADF.2010", "NDF.2010"))
# 2011
forage.quality.2011 <- forage.quality[year == "2011"]
forage.quality.2011 <- forage.quality.2011[, c("Plot", "N", "ADF", "NDF")]
data.table::setnames(forage.quality.2011, old=c("N", "ADF", "NDF"), new = c("N.2011", "ADF.2011", "NDF.2011"))
# 2012
forage.quality.2012 <- forage.quality[year == "2012"]
forage.quality.2012 <- forage.quality.2012[, c("Plot", "N", "ADF", "NDF")]
data.table::setnames(forage.quality.2012, old=c("N", "ADF", "NDF"), new = c("N.2012", "ADF.2012", "NDF.2012"))
# 2013
forage.quality.2013 <- forage.quality[year == "2013"]
forage.quality.2013 <- forage.quality.2013[, c("Plot", "N", "ADF", "NDF")]
data.table::setnames(forage.quality.2013, old=c("N", "ADF", "NDF"), new = c("N.2013", "ADF.2013", "NDF.2013"))
# merge all years with raw_functions
raw_functions <- merge(raw_functions, forage.quality.2009, by="Plot", all=T)
raw_functions <- merge(raw_functions, forage.quality.2010, by="Plot", all=T)
raw_functions <- merge(raw_functions, forage.quality.2011, by="Plot", all=T)
raw_functions <- merge(raw_functions, forage.quality.2012, by="Plot", all=T)
raw_functions <- merge(raw_functions, forage.quality.2013, by="Plot", all=T)
```



# Phosphorous (loss) 2011
measures phoshporous that is not stored in plants or microbes.

Resin refers to the method of phosphorous measurement in soil.
```{r}
P_loss2011 <- data.table::fread(info_data$path_unix[grep("P_loss2011", info_data$ColumnName)], header=T, sep='\t')
# filter out grassland plots
P_loss2011 <- P_loss2011[grep("G", P_loss2011$EP)]
# get new plot names
data.table::setnames(P_loss2011, old=c("EP","Resin_P"), new=c("Plot","P_loss2011"))
raw_functions <- merge(raw_functions, P_loss2011, by="Plot", all=T); rm(P_loss2011)
```

## 2015
Resin P : small bags with filled with resin: a granular material onto which nutrients are adsorbed that comes into contact with the material. These bags are incubated at a certain depth in order that all nutrients that are in solution around the bags is caught in the bag. After incubation, nitrate is detached from the resin and the concentrations are measured.

[µg/g Resin] 	
```{r}
P_loss2015 <- data.table::fread(info_data$path_unix[grep("P_loss2015", info_data$ColumnName)], header=T, sep='\t')
P_loss2015 <- P_loss2015[,c("Plot", "Mean_P")]
data.table::setnames(P_loss2015, old=c("Mean_P"), new=c("P_loss2015"))
raw_functions <- merge(raw_functions, P_loss2015, by="Plot", all=T); rm(P_loss2015)
```


## Phosphorous 2014
Olsen method, Pi is inorganic Phosphorous
```{r}
OlsenPi2014 <- data.table::fread(info_data$path_unix[grep("OlsenPi2014", info_data$ColumnName)], header=T, sep='\t')
# grassland plots are abbreviated with a "G", e.g. "AEG", forest plots with "W", e.g. "AEW"
OlsenPi2014 <- OlsenPi2014[grep("G",OlsenPi2014$Plot_ID),c("Plot_ID","OlsenPi")]
data.table::setnames(OlsenPi2014, old=c("Plot_ID", "OlsenPi"), new=c("Plotn", "OlsenPi2014"))
raw_functions <- merge(raw_functions, OlsenPi2014, by="Plotn", all=T); rm(OlsenPi2014)
```



## Microbial Enzyme activities
from dataset of study which characterised soil microbial communities at three levels of resolution (gene abundance, enzyme activity, and community level).

beginning of May 2011 as a mixed sample from the top horizon (0-10 cm)

beta_Glucosidase [nmol MUF g-1 h-1], Xylosidase [nmol MUF g-1 h-1], N_Acetyl_beta_Glucosaminidase [nmol MUF g-1 h-1], Urease [µg N g-1 DM 2h-1], DEA = DEA_with_C2H2 [µg N2O-N+N2 g-1 soil dw h-1], Phosphatase [nmol MUF g-1 h-1]

all concentrations given g-1.

```{r}
microbial_enzyme_activities <- data.table::fread(info_data$path_unix[grep("beta_Glucosidase", info_data$ColumnName)], header=T, sep='\t', drop=c("Exploratory", "Year"))
data.table::setnames(microbial_enzyme_activities, old=c("EP_Plot_ID","DEA_with_C2H2"), new=c("Plot", "DEA"))
raw_functions <- merge(raw_functions, microbial_enzyme_activities, by="Plot", all=T); rm(microbial_enzyme_activities)
```



# Potential nitrification
Abundance of nitrogen fixing microbes and ammonia oxidizers. abundance and activity of nitrogen fixing microbes and ammonia oxidizing bacteria and archaea.

potential nitrification is determined according to Hoffmann et al, 2007. [ngN / g DW soil / h]

year is not clear, but probably 2011. start date is 1.9.2011, end date is 30.01.2013.

```{r}
Potential.nitrification <- data.table::fread(info_data$path_unix[grep("Potential.nitrification", info_data$ColumnName)])
Potential.nitrification <- Potential.nitrification[,c("PlotID", "pot_nitrification")]
data.table::setnames(Potential.nitrification, old=c("PlotID", "pot_nitrification"), new=c("Plot", "Potential.nitrification"))
raw_functions <- merge(raw_functions, Potential.nitrification, by="Plot", all=T); rm(Potential.nitrification)
```

# Nmic
Within the soil sampling campaign 2011 all samples are taken in the beginning of May 2011 as a mixed sample from the top horizon (0-10 cm) of the all 50 grassland experimental plots of all three exploratories.

For our study, we only use Microbial nitrogen.
```{r}
Nmic.2011 <- data.table::fread(info_data$path_unix[grep("Nmic.2011", info_data$ColumnName)])
Nmic.2011 <- Nmic.2011[,c("EP_Plot_ID", "Microbial_nitrogen")]
data.table::setnames(Nmic.2011, old=c("EP_Plot_ID", "Microbial_nitrogen"), new=c("Plot", "Nmic.2011"))
raw_functions <- merge(raw_functions, Nmic.2011, by="Plot", all=T); rm(Nmic.2011)
```


# NH4 and NO3 2014 - Potential nitrification 2
potential nitrification is measured and NH4 as well as NO3 content is determined.

bdl : below detection limit

11 Plots are measured twice : "SEG40" "SEG41" "SEG42" "SEG43" "SEG44" "SEG45" "SEG46" "SEG47" "SEG48" "SEG49" "SEG50". 
Like in synthesis dataset, the first measure only is taken.

NH4 : µg / g DW soil, Ammonium

NO3 : mg / g DW soil, Nitrate
```{r}
NH4.2014 <- data.table::fread(info_data$path_unix[grep("NH4.2014", info_data$ColumnName)])
NH4.2014 <- NH4.2014[grep("G", NH4.2014$Plot_ID), c("Plot_ID", "NH4", "NO3")]
# show duplicates
# NH4.2014[NH4.2014$Plot_ID %in% NH4.2014[duplicated(NH4.2014$Plot_ID), Plot_ID]]
# filter out second duplicate.
NH4.2014 <- NH4.2014[!which(duplicated(NH4.2014$Plot_ID)),]
# translate "bdl" to "0.000000" as in synthesis dataset
NH4.2014$NH4 <- as.numeric(sub("bdl", "0.000000000", NH4.2014$NH4))
NH4.2014$NO3 <- as.numeric(sub("bdl", "0.000000000", NH4.2014$NO3))
data.table::setnames(NH4.2014, old=c("Plot_ID", "NH4", "NO3"), new=c("Plot", "NH4.2014", "NO3.2014"))
raw_functions <- merge(raw_functions, NH4.2014, by="Plot", all=T); rm(NH4.2014)
```

# mAMFhyphae
on 93 grassland EPs, 2011. 17 from AEG, 22 from HEG and 18 from SEG.

extraradical fungal hyphal length, measured in hyphae per gram soil
```{r}
mAMFhyphae <- data.table::fread(info_data$path_unix[grep("mAMFhyphae", info_data$ColumnName)])
mAMFhyphae <- mAMFhyphae[,c("Plot", "mAMFhyphaepergramsoil")]
# plots measured, number of plots per exploratory
# pm <- raw_functions$Plot[!raw_functions$Plot %in% mAMFhyphae$Plot]
# length(grep("AEG", pm))
# length(grep("HEG", pm))
# length(grep("SEG", pm))
data.table::setnames(mAMFhyphae, old="mAMFhyphaepergramsoil", new="mAMFhyphae")
raw_functions <- merge(raw_functions, mAMFhyphae, by="Plot", all=T); rm(mAMFhyphae)
```

# Groundwater recharge
contains values from 2010 to 2016, each year will be added to a separate column

Annual actual evapotranspiration (ETa), downward water flux (DF), upward water flux (UF), and net flux (NF) were calculated in 14-day resolution for the 0-0.15 m soil layer with a soil water balance model and aggregated to annual values. A complete description of the methods and especially of the differences in the methods used to derive dataset 22486 (water fluxes grassland VIPs; Leimer et al. 2018) can be provided on request. 

NF1 [mm]: net flux betw. 0 - 0.15 m soil layer and deeper soil

Groundwater recharge and discharge areas differ considerably in terms of soil pH and plant species richness, with both increasing towards discharge areas (Giesler et al., 1998; Zinko et al., 2005).

```{r}
Groundwater.recharge <- data.table::fread(info_data$path_unix[grep("Groundwater.recharge", info_data$ColumnName)][1])
# note (why do we need the [1] in the above line?): the info_data file contains entries for all years fo Groundwater recharge separately. 
# But each year uses the same dataset, therefore only the first entry is used to load the dataset for all years.
Groundwater.recharge <- Groundwater.recharge[,c("EP_Plotid", "year", "NF1")]
Gtot <- data.table::copy(PlotExplo[, "Plot"])
# 2010
G <- Groundwater.recharge[year == 2010,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2010"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2010")], by="Plot")
# 2011
G <- Groundwater.recharge[year == 2011,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2011"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2011")], by="Plot")
# 2012
G <- Groundwater.recharge[year == 2012,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2012"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2012")], by="Plot")
# 2013
G <- Groundwater.recharge[year == 2013,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2013"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2013")], by="Plot")
# 2014
G <- Groundwater.recharge[year == 2014,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2014"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2014")], by="Plot")
# 2015
G <- Groundwater.recharge[year == 2015,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2015"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2015")], by="Plot")
# 2016
G <- Groundwater.recharge[year == 2016,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2016"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2016")], by="Plot")

list_raw_functions[["Groundwater.recharge"]] <- Gtot ; rm(Gtot); rm(G); rm(Groundwater.recharge)
# also include years in raw_functions
# raw_functions <- merge(raw_functions, Gtot, by="Plot", all=T)
```


# dung decomposition NOT INCLUDED YET
assessing work of Dung beetles (Scarabaeidae)

quantitative surveys of dung beetles and their dung removal rates. The measure is removal in g dry weight after 48 hours.

percentage of dung dry mass removed after 48 h

some only summer 2014, others the whole years (9 grassland plots per year)

fresh dung was placed in bags accessible to beatles and weighted before and after. after 48h, the samples were removed.
dung with a fresh weight of approx. 220.7 (± 19.9) g of cow, 34.4 (± 3.8) g of horse, 50.5 (± 3.6) g of sheep, 32.6 (± 1.6) g of deer, 14.5 (± 1.4) g of fox and 47.6 (± 2.4) g of wild boar. 

see [PhD thesis](http://tuprints.ulb.tu-darmstadt.de/7402/1/20180502%20Frank.pdf)

```{r, eval=F}
dw <- c("Cow", "Horse", "Sheep", "Deer", "Fox", "Boar")
# "Cow"       "Horse"     "Deer"      "Sheep"     "Wildboar"  "Fox"       "Brownbear" "Elephant"  "Elk"       "Wisent"    "Wolf"
dung.decomposition <- data.table::fread(info_data$path_unix[grep("dung.decomposition", info_data$ColumnName)])
dung.decomposition[EP == "AEG2" & dungtype == "Cow", ]
# 0.458213653
#TODO : ask if there is a script around. 
```

# Litter decomposition
litter bag approach, Mass loss in the litter bags, after two months, and four months. 1.5g of dry plant material. January 2013 put 10 bags each, 2 months later retrieval of 5 bags, 2 more months : retrieval of other 5 bags. calculated daily decomposition rates.

will be assembled from input data.
```{r}
Litter.decomposition <- data.table::fread(info_data$path_unix[grep("Litter.decomposition", info_data$ColumnName)])
list_raw_functions[["Litter.decomposition"]] <- Litter.decomposition ; rm(Litter.decomposition)
```

# Parasitoid traps
will be assembled from input data.
```{r}
AEGpt <- data.table::fread(strsplit(info_data$path_unix[grep("Parasitoid.traps", info_data$ColumnName)], split=", ")[[1]][1])
HEGpt <- data.table::fread(strsplit(info_data$path_unix[grep("Parasitoid.traps", info_data$ColumnName)], split=", ")[[1]][2])
SEGpt <- data.table::fread(strsplit(info_data$path_unix[grep("Parasitoid.traps", info_data$ColumnName)], split=", ")[[1]][3])
Parasitoid.traps <- rbind(AEGpt, HEGpt, SEGpt)
# store in list of raw datasets
list_raw_functions[["Parasitoid.traps"]] <- Parasitoid.traps ; rm(Parasitoid.traps); rm(AEGpt); rm(HEGpt); rm(SEGpt)
```


# Pathogen infection
mean_prevalence X mean_severity, combined with 2011 plant cover data (19686) to calculate total area infected by pathogens per plot

**pathogen data**

4 transects of 25x1 m were searched for foliar fungal pathogens on all grassland EP's. The prevalence and severity of each pathogen was estimated.

species_richness : Richness of foliar fungal pathogen species richness on the plot

mean_prevalence : mean disease prevalence on plot averaged across all pathogens. The percentage of plants with visible infection symptoms.

mean_severity : mean disease severity on plot averaged across all pathogens. The percentage of aboveground plant surface affected by pathogen

% infizierte pflanzen
% pflanzenoberfläche betroffen (über alle pflanzen?!)

**plant cover data** 

```{r, eval=F}
# infections
pathogen.infection <- data.table::fread(info_data$path_unix[grep("pathogen.infection", info_data$ColumnName)][[1]])
pathogen.infection <- pathogen.infection[, c("EP_plot_ID", "mean_prevalence","mean_severity")]
data.table::setnames(pathogen.infection, old=c("EP_plot_ID", "mean_prevalence","mean_severity"), new=c("Plot", "mean_prevalence","mean_severity"))

# plant cover
pathogen.infection2 <- data.table::fread(info_data$path_unix[grep("pathogen.infection2", info_data$ColumnName)])
pathogen.infection2 <- pathogen.infection2[Year == "2011", c("EP_PlotID", "Species", "Cover")]
data.table::setnames(pathogen.infection2, old = "EP_PlotID", new= "Plot")
pathogen.infection2[, Cover := as.numeric(Cover)]

list_raw_functions[["pathogen.infection"]] <- pathogen.infection; rm(pathogen.infection)
list_raw_functions[["pathogen.infection2"]] <- pathogen.infection2; rm(pathogen.infection2)
```

# Total pollinators
The total amount of visits per plot will be taken as total pollinators function.
```{r}
Total.pollinators <- data.table::fread(info_data$path_unix[grep("Total.pollinators", info_data$ColumnName)])
Total.pollinators <- Total.pollinators[,c("EP_ID", "Total")]
data.table::setnames(Total.pollinators, old=c("EP_ID", "Total"), new=c("Plot", "Total.pollinators"))
list_raw_functions[["Total.pollinators"]] <- Total.pollinators; rm(Total.pollinators)
```



# Charismatic butterflys
Diversity and abundance of day active butterflies and day-active moths along a land use intensity gradient (2008)

```{r, eval=F}
butterflys <- data.table::fread(info_data$path_unix[grep("Charism.butterfly.Abund", info_data$ColumnName)])
```

# amoA_AOA and AOB
Nitrifiers Abundances at Grassland, The abundance of ammonia oxidizing bacteria and archaea was estimated using respective amoA gene.

AOA : ammonia oxidizing archaea

AOB : ammonia oxidizing bacteria

```{r, eval=F}
amo <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/21547_Nitrifiers Abundances at Grassland EPs_1.6.8/21547.txt")
data.table::setnames(amo, old=c("Plot_ID"), new=c("Plot"))
raw_functions <- merge(raw_functions, amo[,c("Plot","amoA_AOA", "amoA_AOB")], by="Plot", all = T); rm(amo)
```

# soil depth
soil.depth

<!-- Soils were described according to the German soil classification system (KA 5) and the FAO-Guidelines for Soil Description (2006). -->

<!-- A multitude of soil measures were taken. -->

<!-- theory : plants share the belowground space. A diverse set of plants can vary in their root length - different plants sometimes differ in the depth where they are able to get nutrients. The "potential rooting depth" is the soil depth in which plants can potentially have roots. Soil horizons O, A and B can harbour plant roots. -->

<!-- Mächtigkeit der obsersten bodenschicht herausfinden. -->
<!-- ```{r, eval=F} -->
<!-- z <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/5340_Soil_ Inventory organic soils grid plots_1.287.23_PublicDataset/5340_Block0.txt") -->
<!-- o <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/5340_Soil_ Inventory organic soils grid plots_1.287.23_PublicDataset/5340_Block1.txt") -->
<!-- t <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/5340_Soil_ Inventory organic soils grid plots_1.287.23_PublicDataset/5340_Block2.txt") -->

<!-- data.table::fread(info_data$path_unix[grep("Nshoot.2009", info_data$ColumnName)]) -->
<!-- ``` -->

using soil depth from Eric, who assembled it from dataset 5340 (as indicated in synthesis dataset 21688)

```{r}
soil.depth <- data.table::fread(info_data$path_unix[grep("soil.depth", info_data$ColumnName)])
data.table::setnames(soil.depth, old="Plot", new="Plotn")
raw_functions <- merge(raw_functions, soil.depth[,c("Plotn","soil.depth")], by="Plotn", all = T); rm(soil.depth)
```

# pH
soil PH

