---
title: "prepare functions"
author: "noelle"
date: "November 14, 2018"
output: html_document
  toc: true
  toc_depth: 2
---

# Loading raw data
An overview of the data is given in "info_data.csv". From there, the path of data files can be read out. 

In order to read in data correctly, the path to the folder which contains the data needs to be given (`pathtodata`). The data means the downloaded and unzipped data files as they are downloaded from Bexis. The names of the BExIS files are given in "info_data.csv". The path in `pathtodata` is pasted together with the path given in "info_data.csv" (column "subpath").

Example : 

`pathtodata <- "C:/Documents/Projects/Synthesis/Functions_dataset/BExIS_raw_data/"`

will be merged with "12706_Measurement of biomass (2010, all grassland EPs)_1.1.8_PublicDataset/12706.txt" from "info_data.csv" to load Biomass data from 2010.

Reading pathtodata and the file "info_data.csv"
```{r}
#TODO : insert needed path.
# pathtodata <- "C:/Documents/Projects/Synthesis/Functions_dataset/BExIS_raw_data/"

#TODO : insert path to info_data.csv
# info_data <- data.table::fread("C:/Documents/Projects/Synthesis/Functions_dataset/info_data.csv")
```

## Loading raw data, University of Bern
In case you are a member of the University of Bern and have access to the shared network drive from the University of Bern, you can use the file `nonpublic.R` to load the required paths. In this case, no `pathtodata` variable needs to be given (because this is done in `nonpublic.R`).

*note* : In order to read in data, you need to be connected to server. 

```{r}
# only show lines which contain path info
info_data <- info_data[grep("/", info_data$subpath),c("ColumnName", "subpath")]
#TODO : for windows : change "/" to "\" in info_data
```


# Plot, exploratories
```{r}
PlotExplo <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2011", info_data$ColumnName)], sep=""), header=T, sep='\t')
PlotExplo <- PlotExplo[,c("EpPlotID", "explo")]
data.table::setnames(PlotExplo, old=c("EpPlotID", "explo"), new=c("Plot", "Explo"))
# generate old plot names
Plot <- c(paste("AEG", seq(1,50), sep=""), paste("HEG", seq(1,50), sep=""), paste("SEG", seq(1,50), sep=""))
# generate new plot names
Plotn <- c(paste("AEG", sprintf("%02d", 1:50), sep=""), paste("HEG", sprintf("%02d", 1:50), sep=""), paste("SEG", sprintf("%02d", 1:50), sep=""))
PlotExplo <- merge(PlotExplo, cbind(Plot, Plotn), by = "Plot")

rm(Plot); rm(Plotn)
```

# raw dataset
Generate a dataset which stores all raw data. Will be assembled with the script `2calc_raw_dataset.R`. Datasets that can not be easily included as raw datasets into `raw_functions` are assembled (calculations are done) in `2calc_raw_dataset.R` and then integrated into raw_functions as well (and assembled functions variable, of corse).

*Covariates* : are also included in the functions dataset and taken out later.
```{r}
# subset for raw dataset construction
raw_functions <- data.table::copy(PlotExplo); rm(PlotExplo)
# generate list for datasets that will be constructed in 2calc_raw_dataset.R script directly.
list_raw_functions <- list()
```



# Biomass
years 2008 to 2017. **caution**, 2008 might not be reliable.

all biomasses in g/m^2

2009 : Biomass was clipped in each two subplots per plot. In each subplot we clipped four 50 x 50 cm plots 2-3 cm above-ground.
two treatments, K1 and K2, K2 is taken in synthesis. 149 plots only.

*difference of 2009 to other years* : 
- clipped mid april to mid may, instead of mid may to mid june
- same area, more replicates per subplot
solution : either correct all samples to mid may to mid june or take average
```{r}

# 2008
Biomass2008 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2008", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2008 <- Biomass2008[, c("PlotID", "biomass_g_m_2")]
data.table::setnames(Biomass2008, old= c("PlotID", "biomass_g_m_2"), new = c("Plot", "Biomass2008"))
raw_functions <- merge(raw_functions, Biomass2008, by="Plot", all=T)
rm(Biomass2008)

# 2009
Biomass2009 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2009", info_data$ColumnName)], sep=""), header=T, sep='\t')
# Biomass2009 <- Biomass2009[Treatment == "K2",]
# Biomass2009 <- Biomass2009[,c("GpPlotID", "Biomass")]
Biomass2009 <- Biomass2009[, c("GpPlotID", "Day_of_year", "Treatment", "Biomass")]
data.table::setnames(Biomass2009, old=c("GpPlotID", "Biomass"), new=c("Plot", "Biomass2009"))
# raw_functions <- merge(raw_functions, Biomass2009, by="Plot", all.x=T)
list_raw_functions[["Biomass2009"]] <- Biomass2009; rm(Biomass2009)

# 2010
Biomass2010 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2010", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2010 <- Biomass2010[, c("PlotID", "Biomasse")]
data.table::setnames(Biomass2010, old= c("PlotID", "Biomasse"), new = c("Plot", "Biomass2010"))
raw_functions <- merge(raw_functions, Biomass2010, by="Plot", all=T)
rm(Biomass2010)

# 2011
Biomass2011 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2011", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2011 <- Biomass2011[,c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2011, old = c("EpPlotID", "biomass_g"), new = c("Plot", "Biomass2011"))
raw_functions <- merge(raw_functions, Biomass2011, by="Plot", all=T)
rm(Biomass2011)

# 2012
Biomass2012 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2012", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2012 <- Biomass2012[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2012, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2012"))
raw_functions <- merge(raw_functions, Biomass2012, by="Plot", all=T)
rm(Biomass2012)

# 2013
Biomass2013 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2013", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2013 <- Biomass2013[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2013, old= c("EpPlotID", "biomass_g"), new= c("Plot", "Biomass2013"))
raw_functions <- merge(raw_functions, Biomass2013, by="Plot", all=T); rm(Biomass2013)

# 2014
Biomass2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2014", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2014 <- Biomass2014[,c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2014, old = c("EpPlotID", "biomass_g"), new = c("Plot", "Biomass2014"))
raw_functions <- merge(raw_functions, Biomass2014, by="Plot", all=T); rm(Biomass2014)

# 2015
Biomass2015 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2015", info_data$ColumnName)], sep=""), header=T, sep='\t', quote="", drop=19)
Biomass2015 <- Biomass2015[,c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2015, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2015"))
raw_functions <- merge(raw_functions, Biomass2015, by="Plot", all=T); rm(Biomass2015)

# 2016
Biomass2016 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2016", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2016 <- Biomass2016[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2016, old= c("EpPlotID", "biomass_g"), new= c("Plot", "Biomass2016"))
raw_functions <- merge(raw_functions, Biomass2016, by="Plot", all=T); rm(Biomass2016)

# 2017
Biomass2017 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Biomass2017", info_data$ColumnName)], sep=""), header=T, sep='\t')
Biomass2017 <- Biomass2017[, c("EpPlotID", "biomass_g")]
data.table::setnames(Biomass2017, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2017"))
raw_functions <- merge(raw_functions, Biomass2017, by="Plot", all=T); rm(Biomass2017)
```



# Soil Organic C
2011
```{r}
SoilOrganicC <- data.table::fread(paste(pathtodata, info_data$subpath[grep("SoilOrganicC", info_data$ColumnName)], sep=""), header=T, sep='\t')
SoilOrganicC <- SoilOrganicC[which(SoilOrganicC$Type == "G"), c("EP_Plotid","Organic_C")]
data.table::setnames(SoilOrganicC, old=c("EP_Plotid", "Organic_C"), new=c("Plot", "SoilOrganicC"))
raw_functions <- merge(raw_functions, SoilOrganicC, by="Plot", all=T); rm(SoilOrganicC)
```


# Root biomass
2011, 
Includes fine and coarse biomass, will add them together.
```{r}
Root.biomass <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Root.biomass", info_data$ColumnName)], sep=""), header=T, sep='\t')
Root.biomass <- Root.biomass[Type == "G"]
Root.biomass <- Root.biomass[,c("EP_Plotid", "Fine_Roots_Biomass", "Coarse_Roots_Biomass")]
data.table::setnames(Root.biomass, old="EP_Plotid", new="Plot")
list_raw_functions[["Root.biomass"]] <- Root.biomass ; rm(Root.biomass)
```

# Flower cover
**leave out**
We can go from "total_influorescence_area_in_cm" of each species to a total flower cover. The total area measured was : 600m² per grassland plot.
TODO : left open as not sure if I include it
```{r, eval=F}
Total.flower.cover <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Total.flower.cover", info_data$ColumnName)], sep=""), header=T, sep='\t')
```

# PRI - Phosphorous retention index
## Microbial phosphorous in soil
2011 and 2014

(mg/kg)

```{r}
Pmic.2011 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Pmic.2011", info_data$ColumnName)], sep=""), header=T, sep='\t')
Pmic.2011 <- Pmic.2011[grep("G", Pmic.2011$EP)]
data.table::setnames(Pmic.2011, old=c("EP", "Pmic"), new=c("Plot", "Pmic.2011"))
raw_functions <- merge(raw_functions, Pmic.2011, by="Plot", all=T); rm(Pmic.2011)
```

```{r}
Pmic.2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Pmic.2014", info_data$ColumnName)], sep=""), header=T, sep='\t')
Pmic.2014 <- Pmic.2014[,c("EP", "Pmic")]
data.table::setnames(Pmic.2014, old=c("EP", "Pmic"), new=c("Plot", "Pmic.2014"))
raw_functions <- merge(raw_functions, Pmic.2014, by="Plot", all=T) ; rm(Pmic.2014)
```

## Soil Phosphorous fractions
used to calculate PRI

years : 2008, 2009

better use Olsen Pi than NaHCO3_Pi, because olsen measurese the plant available P
```{r}
np <- data.table::fread(paste(pathtodata, info_data$subpath[grep("NaHCO3.Pi", info_data$ColumnName)], sep=""), header=T, sep='\t') 
np <- np[grep("G", np$EP)]
np <- np[, c("EP", "NaHCO3_Pi")]
data.table::setnames(np, old="EP", new="Plot")
raw_functions <- merge(raw_functions, np, by="Plot", all=T) ; rm(np)
```


## Plant Nutrients (Pshoot)
2009, 2010, 2011, 2012, 2013

contains: Nshoot and Pshoot of all years

near infrared spectroscopy (NIRS), 

*2014* data comes from different dataset.

note : there are **values for 2008** in dataset 5401 (used for Biomass2008)
```{r}
nutrients <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Nshoot.2009", info_data$ColumnName)], sep=""), header=T, sep='\t')
# 2009
nutrients2009 <- nutrients[year == "2009", c("Plot", "N", "P")]
data.table::setnames(nutrients2009, old=c("N", "P"), new=c("Nshoot.2009", "Pshoot.2009"))
# 2010
nutrients2010 <- nutrients[year == "2010", c("Plot", "N", "P")]
data.table::setnames(nutrients2010, old=c("N", "P"), new=c("Nshoot.2010", "Pshoot.2010"))
# 2011
nutrients2011 <- nutrients[year == "2011", c("Plot", "N", "P")]
data.table::setnames(nutrients2011, old=c("N", "P"), new=c("Nshoot.2011", "Pshoot.2011"))
# 2012
nutrients2012 <- nutrients[year == "2012", c("Plot", "N", "P")]
data.table::setnames(nutrients2012, old=c("N", "P"), new=c("Nshoot.2012", "Pshoot.2012"))
# 2013
nutrients2013 <- nutrients[year == "2013", c("Plot", "N", "P")]
data.table::setnames(nutrients2013, old=c("N", "P"), new=c("Nshoot.2013", "Pshoot.2013"))

# read in 2014 data
nutrients2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Nshoot.2014", info_data$ColumnName)], sep=""), header=T, sep='\t')
nutrients2014 <- nutrients2014[Year == "2014", c("Plot_ID", "N", "P")]
data.table::setnames(nutrients2014, old = c("Plot_ID", "N", "P"), new=c("Plot", "Nshoot.2014", "Pshoot.2014"))

# take nutrients together for list_raw_functions
nutrients <- Reduce(function(x, y) merge(x, y, all=TRUE), list(nutrients2009, nutrients2010, nutrients2011, nutrients2012, nutrients2013, nutrients2014))
list_raw_functions[["nutrients"]] <- nutrients
raw_functions <- merge(raw_functions, nutrients, by="Plot", all=T)
rm(nutrients); rm(nutrients2009); rm(nutrients2010); rm(nutrients2011); rm(nutrients2012); rm(nutrients2013); rm(nutrients2014)
```


# NRI - Nitrogen retention index
requires : Bulk.density2014, Nshoot.2014, NH4.2014, NO3.2014, Biomass2014, Nmic.2014

Most of these datasets are already read in in this script. Nmic.2014 needs to be read in.

```{r, eval=F}
#TODO : wait for acces to nmic dataset
Nmic.2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Nmic.2014", info_data$ColumnName)], sep=""), header=T, sep='\t')

raw_functions <- merge(raw_functions, Nmic.2014, by="Plot", all=T); rm(Root.decomposition)
```



# Root decomposition
Decomposition measurements of fine roots (<2mm) within the upper 10 cm of the mineral soil. Determination of mass loss after 6 months. Litterbag method. Standardized control material: beech roots for forest sites and herbaceous roots for grassland sites. Numbers in [%]

mass "loss" because decomposed roots are lost.
```{r}
Root.decomposition <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Root.decomposition", info_data$ColumnName)], sep=""), header=T, sep='\t')
Root.decomposition <- Root.decomposition[Type == "G"]
Root.decomposition <- data.table::setnames(Root.decomposition[,c("EP_Plotid", "Mass_loss_October_2012")], old=c("EP_Plotid", "Mass_loss_October_2012"), new=c("Plot", "Root.decomposition"))
raw_functions <- merge(raw_functions, Root.decomposition, by="Plot", all=T); rm(Root.decomposition)
```


# Bulk density and soil C stock
**Bulk density** : Increased use of machine compact the soil. At a given point, the root growth is impaired.

**soil C stock** : C in soil

Determination bulk density and calculated CN stocks for the upper 10 cm of the mineral soil. Dry combustion of soil at a temperature of 1100°C and subsequent determination of evolving CO2 and N2 with a Thermal Conductivity Detector (TCD).

[g/cm^3]

**data is from 2011, not 2014!**

```{r}
Bulk.density.soil.C.stock.2011 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Soil.C.stock", info_data$ColumnName)], sep=""), header=T, sep='\t')
Bulk.density.soil.C.stock.2011 <- Bulk.density.soil.C.stock.2011[Type == "G"]
Bulk.density.soil.C.stock.2011 <- data.table::setnames(Bulk.density.soil.C.stock.2011[,c("EP_Plotid", "BD", "OC_stock")], old=c("EP_Plotid", "BD", "OC_stock"), new=c("Plot", "Bulk.density2011", "Soil.C.stock"))
raw_functions <- merge(raw_functions, Bulk.density.soil.C.stock.2011, by="Plot", all=T); rm(Bulk.density.soil.C.stock.2011)

# 2014
Bulk.density2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Bulk.density2014", info_data$ColumnName)], sep=""), header=T, sep='\t')
Bulk.density2014 <- Bulk.density2014[, c("EP_Plot_ID", "Bulk_density")]
data.table::setnames(Bulk.density2014, old = c("EP_Plot_ID", "Bulk_density"), new = c("Plot", "Bulk.density2014"))
raw_functions <- merge(raw_functions, Bulk.density2014, by="Plot", all=T); rm(Bulk.density2014)
```

# Herbivory

## 2013
invertebrate herbivory in 146 managed temperate grasslands. Year: 2013

increased land-use - less invertebrates - less herbivory

Visually estimation of herbivory by eye using templates and measurement of leaf area left after feeding of the herbivores using LI-COR area meter

2 samples per plot in 145 cases, 1 sample for :  "AEG6"  "HEG16" "HEG43" "SEG19" "SEG33" and 3 samples for "SEG45". 2 Columns are included in the raw_functions dataset, 1 for each sample.

SEG45 : The mean of all 3 samples was already taken and stored as sample2. Taking the mean of samples2 first and then with sample1 would lead to different results than in the synthesis dataset.
```{r}
herb <- data.table::fread(paste(pathtodata, info_data$subpath[grep("herbivory.2013", info_data$ColumnName)], sep=""), header=T, sep='\t')
data.table::setorder(herb, EP_Plotid)
list_raw_functions[["Herbivory.2013"]] <- herb ; rm(herb)
```

## 2017 and 2018
plant material was collected before the first mowing (May 2017/2018). For each plant, we visually estimated the area damaged by herbivorous insects and gall mites on 12 - 200 leaves (depending on leaf size) and measured total leaf area using a leaf area meter. The deduced herbivory rates (% damaged area) per plant species were than summarised to community-level herbivory rates based on the respective plant cover values in the vegetation records of the survey years (2017 for AEG, AEW, HEW, SEW; 2018 for HEG, SEG from data set 24247 (pathogen infection)).

*Noelle :* 
prop_a_herb : proportion of leaf area damaged of the 10 (8) most abundant species. Serves as estimate for the whole plot and can therefore be taken directly.

AEG plots were measured in 2017, SEG/HEG in 2018. 49 AEG plots, 48 HEG plots and 50 SEG plots.

```{r}
herb <- data.table::fread(paste(pathtodata, info_data$subpath[grep("herbivory.20172018", info_data$ColumnName)], sep=""), header=T, sep='\t')
data.table::setnames(herb, old=c("plotid_withzero", "plotid"), new=c("Plotn", "Plot"))
herb <- herb[grep("G", herb$Plotn)]
data.table::setorder(herb, Plotn)
data.table::setnames(herb, old="prop_a_herb", new="herbivory.20172018")
raw_functions <- merge(raw_functions, herb[,c("Plot", "herbivory.20172018")], by="Plot", all=T); rm(herb)
```



# Forage quality - NOT READY YET
"Futter qualität", year : 2009
forage quality (index based on crude protein concentration and relative forage value) - Eric Ecology Letters LUI

RFV : Relative Feed Value
RFQ : Relative Forage Quality
ADL : acid detergent lignin, Lignin
ADF : Acid Detergent Fibre, Säure-Detergenz-Faser, ein Teil der Gerüstsubstanzen von Futtermitteln
NDF : Neutral Detergent Fibre, Neutral-Detergenz-Faser, fasrige Anteile von Futtermitteln 

more ADF/ NDF --> less nutritional value because less easy to digest

according to forage quality estimation based on alfalfa (medicago sativa) values. [source](https://openprairie.sdstate.edu/cgi/viewcontent.cgi?referer=https://scholar.google.ch/&httpsredir=1&article=1351&context=extension_extra)

[% concentration in aboveground grassland plant community biomass without litter]
```{r}
forage.quality <- data.table::fread(paste(pathtodata, info_data$subpath[grep("forage.quality", info_data$ColumnName)], sep=""), header=T, sep='\t')
list_raw_functions[["forage.quality"]] <- forage.quality; rm(forage.quality)

# # 2009
# forage.quality.2009 <- forage.quality[year == "2009"]
# forage.quality.2009 <- forage.quality.2009[, c("Plot", "N", "ADF", "NDF")]
# data.table::setnames(forage.quality.2009, old=c("N", "ADF", "NDF"), new = c("N.2009", "ADF.2009", "NDF.2009"))
# # 2010
# forage.quality.2010 <- forage.quality[year == "2010"]
# forage.quality.2010 <- forage.quality.2010[, c("Plot", "N", "ADF", "NDF")]
# data.table::setnames(forage.quality.2010, old=c("N", "ADF", "NDF"), new = c("N.2010", "ADF.2010", "NDF.2010"))
# # 2011
# forage.quality.2011 <- forage.quality[year == "2011"]
# forage.quality.2011 <- forage.quality.2011[, c("Plot", "N", "ADF", "NDF")]
# data.table::setnames(forage.quality.2011, old=c("N", "ADF", "NDF"), new = c("N.2011", "ADF.2011", "NDF.2011"))
# # 2012
# forage.quality.2012 <- forage.quality[year == "2012"]
# forage.quality.2012 <- forage.quality.2012[, c("Plot", "N", "ADF", "NDF")]
# data.table::setnames(forage.quality.2012, old=c("N", "ADF", "NDF"), new = c("N.2012", "ADF.2012", "NDF.2012"))
# # 2013
# forage.quality.2013 <- forage.quality[year == "2013"]
# forage.quality.2013 <- forage.quality.2013[, c("Plot", "N", "ADF", "NDF")]
# data.table::setnames(forage.quality.2013, old=c("N", "ADF", "NDF"), new = c("N.2013", "ADF.2013", "NDF.2013"))
# # merge all years with raw_functions
# raw_functions <- merge(raw_functions, forage.quality.2009, by="Plot", all=T)
# raw_functions <- merge(raw_functions, forage.quality.2010, by="Plot", all=T)
# raw_functions <- merge(raw_functions, forage.quality.2011, by="Plot", all=T)
# raw_functions <- merge(raw_functions, forage.quality.2012, by="Plot", all=T)
# raw_functions <- merge(raw_functions, forage.quality.2013, by="Plot", all=T)
```



# Phosphorous (loss) 2011
measures phoshporous that is not stored in plants or microbes.

Resin refers to the method of phosphorous measurement in soil.
```{r}
P_loss2011 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("P_loss2011", info_data$ColumnName)], sep=""), header=T, sep='\t')
# filter out grassland plots
P_loss2011 <- P_loss2011[grep("G", P_loss2011$EP)]
# get new plot names
data.table::setnames(P_loss2011, old=c("EP","Resin_P"), new=c("Plot","P_loss2011"))
# take reverse values so high values represent low loss
P_loss2011[, P_loss2011 := (max(P_loss2011, na.rm=T) - P_loss2011)]
raw_functions <- merge(raw_functions, P_loss2011, by="Plot", all=T); rm(P_loss2011)
```

## 2015
Resin P : small bags with filled with resin: a granular material onto which nutrients are adsorbed that comes into contact with the material. These bags are incubated at a certain depth in order that all nutrients that are in solution around the bags is caught in the bag. After incubation, nitrate is detached from the resin and the concentrations are measured.

[µg/g Resin] 	
```{r}
P_loss2015 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("P_loss2015", info_data$ColumnName)], sep=""), header=T, sep='\t')
P_loss2015 <- P_loss2015[,c("Plot", "Mean_P")]
data.table::setnames(P_loss2015, old=c("Mean_P"), new=c("P_loss2015"))
# take reverse values so high values represent low loss
P_loss2015[, P_loss2015 := (max(P_loss2015, na.rm=T) - P_loss2015)]
raw_functions <- merge(raw_functions, P_loss2015, by="Plot", all=T); rm(P_loss2015)
```


## Phosphorous 2014
Olsen method, Pi is inorganic Phosphorous
```{r}
OlsenPi2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("OlsenPi2014", info_data$ColumnName)], sep=""), header=T, sep='\t')
# grassland plots are abbreviated with a "G", e.g. "AEG", forest plots with "W", e.g. "AEW"
OlsenPi2014 <- OlsenPi2014[grep("G",OlsenPi2014$Plot_ID),c("Plot_ID","OlsenPi")]
data.table::setnames(OlsenPi2014, old=c("Plot_ID", "OlsenPi"), new=c("Plotn", "OlsenPi2014"))
raw_functions <- merge(raw_functions, OlsenPi2014, by="Plotn", all=T); rm(OlsenPi2014)
```



## Microbial Enzyme activities
from dataset of study which characterised soil microbial communities at three levels of resolution (gene abundance, enzyme activity, and community level).

beginning of May 2011 as a mixed sample from the top horizon (0-10 cm)

beta_Glucosidase [nmol MUF g-1 h-1], Xylosidase [nmol MUF g-1 h-1], N_Acetyl_beta_Glucosaminidase [nmol MUF g-1 h-1], Urease [µg N g-1 DM 2h-1], DEA = DEA_with_C2H2 [µg N2O-N+N2 g-1 soil dw h-1], Phosphatase [nmol MUF g-1 h-1]

all concentrations given g-1.

```{r}
microbial_enzyme_activities <- data.table::fread(paste(pathtodata, info_data$subpath[grep("beta_Glucosidase", info_data$ColumnName)], sep=""), header=T, sep='\t', drop=c("Exploratory", "Year"))
data.table::setnames(microbial_enzyme_activities, old=c("EP_Plot_ID","DEA_with_C2H2"), new=c("Plot", "DEA"))
raw_functions <- merge(raw_functions, microbial_enzyme_activities, by="Plot", all=T); rm(microbial_enzyme_activities)
```


# Nmic
Within the soil sampling campaign 2011 all samples are taken in the beginning of May 2011 as a mixed sample from the top horizon (0-10 cm) of the all 50 grassland experimental plots of all three exploratories.

For our study, we only use Microbial nitrogen.
```{r}
Nmic.2011 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Nmic.2011", info_data$ColumnName)], sep=""), header=T, sep='\t')
Nmic.2011 <- Nmic.2011[,c("EP_Plot_ID", "Microbial_nitrogen")]
data.table::setnames(Nmic.2011, old=c("EP_Plot_ID", "Microbial_nitrogen"), new=c("Plot", "Nmic.2011"))
raw_functions <- merge(raw_functions, Nmic.2011, by="Plot", all=T); rm(Nmic.2011)
```

# Potential nitrification
Abundance of nitrogen fixing microbes and ammonia oxidizers. abundance and activity of nitrogen fixing microbes and ammonia oxidizing bacteria and archaea.

potential nitrification is determined according to Hoffmann et al, 2007. [ngN / g DW soil / h]

year is not clear, but probably 2011. start date is 1.9.2011, end date is 30.01.2013.

```{r}
Potential.nitrification <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Potential.nitrification2011", info_data$ColumnName)], sep=""), header=T, sep='\t')
Potential.nitrification <- Potential.nitrification[,c("PlotID", "pot_nitrification")]
data.table::setnames(Potential.nitrification, old=c("PlotID", "pot_nitrification"), new=c("Plot", "Potential.nitrification2011"))
raw_functions <- merge(raw_functions, Potential.nitrification, by="Plot", all=T); rm(Potential.nitrification)
```


# NH4 and NO3 2014 - Potential nitrification 2
potential nitrification is measured and NH4 as well as NO3 content is determined.

bdl : below detection limit

11 Plots are measured twice : "SEG40" "SEG41" "SEG42" "SEG43" "SEG44" "SEG45" "SEG46" "SEG47" "SEG48" "SEG49" "SEG50". 
Like in synthesis dataset, the first measure only is taken.

NH4 : µg / g DW soil, Ammonium

NO3 : mg / g DW soil, Nitrate
```{r}
NH4.2014 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("NH4.2014", info_data$ColumnName)], sep=""), header=T, sep='\t')
NH4.2014 <- NH4.2014[grep("G", NH4.2014$Plot_ID), c("Plot_ID", "NH4", "NO3", "PNR")]
# show duplicates
# NH4.2014[NH4.2014$Plot_ID %in% NH4.2014[duplicated(NH4.2014$Plot_ID), Plot_ID]]
# filter out second duplicate.
NH4.2014 <- NH4.2014[!which(duplicated(NH4.2014$Plot_ID)),]
# translate "bdl" (below detection limit) to "0.000000" as in synthesis dataset. 
# "nd" (not detectable) is automatically translated to NA
NH4.2014$NH4 <- as.numeric(sub("bdl", "0.000000", NH4.2014$NH4))
NH4.2014$NO3 <- as.numeric(sub("bdl", "0.000000", NH4.2014$NO3))
NH4.2014$PNR <- as.numeric(sub("bdl", "0.000000", NH4.2014$PNR))
data.table::setnames(NH4.2014, old=c("Plot_ID", "NH4", "NO3", "PNR"), new=c("Plot", "NH4.2014", "NO3.2014", "Potential.nitrification2014"))
raw_functions <- merge(raw_functions, NH4.2014, by="Plot", all=T); rm(NH4.2014)
```

# mAMFhyphae and Aggreggation
on 93 grassland EPs, 2011. 17 from AEG, 22 from HEG and 18 from SEG.

explore how land use was affecting the ability of soil microbes to aggregate soil.

extraradical fungal hyphal length, measured in hyphae per gram soil
```{r}
mAMFhyphae <- data.table::fread(paste(pathtodata, info_data$subpath[grep("mAMFhyphae", info_data$ColumnName)], sep=""), header=T, sep='\t')
mAMFhyphae <- mAMFhyphae[,c("Plot", "percentAggregation", "mAMFhyphaepergramsoil")]
# plots measured, number of plots per exploratory
# pm <- raw_functions$Plot[!raw_functions$Plot %in% mAMFhyphae$Plot]
# length(grep("AEG", pm))
# length(grep("HEG", pm))
# length(grep("SEG", pm))
data.table::setnames(mAMFhyphae, old=c("percentAggregation", "mAMFhyphaepergramsoil"), new=c("Aggregation", "mAMFhyphae"))
raw_functions <- merge(raw_functions, mAMFhyphae, by="Plot", all=T); rm(mAMFhyphae)
```



# Groundwater recharge
contains values from 2010 to 2016, each year will be added to a separate column

Annual actual evapotranspiration (ETa), downward water flux (DF), upward water flux (UF), and net flux (NF) were calculated in 14-day resolution for the 0-0.15 m soil layer with a soil water balance model and aggregated to annual values. A complete description of the methods and especially of the differences in the methods used to derive dataset 22486 (water fluxes grassland VIPs; Leimer et al. 2018) can be provided on request. 

NF1 [mm]: net flux betw. 0 - 0.15 m soil layer and deeper soil

Groundwater recharge and discharge areas differ considerably in terms of soil pH and plant species richness, with both increasing towards discharge areas (Giesler et al., 1998; Zinko et al., 2005).


```{r}
Groundwater.recharge <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Groundwater.recharge", info_data$ColumnName)][1], sep=""), header=T, sep='\t')
Groundwater.recharge <- Groundwater.recharge[,c("EP_Plotid", "year", "NF1")]
Gtot <- data.table::copy(raw_functions[, "Plot"])
# 2010
G <- Groundwater.recharge[year == 2010,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2010"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2010")], by="Plot")
# 2011
G <- Groundwater.recharge[year == 2011,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2011"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2011")], by="Plot")
# 2012
G <- Groundwater.recharge[year == 2012,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2012"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2012")], by="Plot")
# 2013
G <- Groundwater.recharge[year == 2013,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2013"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2013")], by="Plot")
# 2014
G <- Groundwater.recharge[year == 2014,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2014"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2014")], by="Plot")
# 2015
G <- Groundwater.recharge[year == 2015,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2015"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2015")], by="Plot")
# 2016
G <- Groundwater.recharge[year == 2016,]
data.table::setnames(G, old=c("EP_Plotid", "NF1"), new=c("Plot", "Groundwater.recharge2016"))
Gtot <- merge(Gtot, G[,c("Plot", "Groundwater.recharge2016")], by="Plot")

list_raw_functions[["Groundwater.recharge"]] <- Gtot ; rm(G); rm(Groundwater.recharge)
# also include years in raw_functions
raw_functions <- merge(raw_functions, Gtot, by="Plot", all=T) ; rm(Gtot)
```


# dung decomposition
assessing work of Dung beetles (Scarabaeidae)

quantitative surveys of dung beetles and their dung removal rates. The measure is removal in g dry weight after 48 hours. percentage of dung dry mass removed after 48 h. fresh dung was placed in bags accessible to beatles and weighted before and after. after 48h, the samples were removed.
dung with a fresh weight of approx. 220.7 (± 19.9) g of cow, 34.4 (± 3.8) g of horse, 50.5 (± 3.6) g of sheep, 32.6 (± 1.6) g of deer, 14.5 (± 1.4) g of fox and 47.6 (± 2.4) g of wild boar.

see [PhD thesis](http://tuprints.ulb.tu-darmstadt.de/7402/1/20180502%20Frank.pdf)

some measurements only summer 2014, others the whole years (9 grassland plots per year)

Only measurements from summer 2014 taken, where 5 replicates were measured on each plot. 

```{r}
dung.decomposition <- data.table::fread(paste(pathtodata, info_data$subpath[grep("dung.decomposition", info_data$ColumnName)], sep=""), header=T, sep='\t')
# only take Grassland plots
dung.decomposition <- dung.decomposition[grep("G", dung.decomposition$EP)]
dung.decomposition <- dung.decomposition[grep("2014", dung.decomposition$month)]
dung.decomposition <- dung.decomposition[month %in% c("June_2014", "July_2014", "August_2014")]
# p <- ggplot(dung.decomposition, aes(x=EP, color=dungtype)) +
#   geom_histogram(stat="count", alpha=1)
list_raw_functions[["dung.decomposition"]] <- dung.decomposition[,c("EP", "dungtype", "removal_g")] ; rm(dung.decomposition)
```

# Litter decomposition
litter bag approach, Mass loss in the litter bags, after two months, and four months. 1.5g of dry plant material. January 2013 put 10 bags each, 2 months later retrieval of 5 bags, 2 more months : retrieval of other 5 bags. calculated daily decomposition rates.

will be assembled from input data.
```{r}
Litter.decomposition <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Litter.decomposition", info_data$ColumnName)], sep=""), header=T, sep='\t')
list_raw_functions[["Litter.decomposition"]] <- Litter.decomposition ; rm(Litter.decomposition)
```

# Parasitoid traps
will be assembled from input data.
```{r}
AEGpt <- data.table::fread(strsplit(info_data$subpath[grep("Parasitoid.traps", info_data$ColumnName)], split=", ")[[1]][1])
HEGpt <- data.table::fread(strsplit(info_data$subpath[grep("Parasitoid.traps", info_data$ColumnName)], split=", ")[[1]][2])
SEGpt <- data.table::fread(strsplit(info_data$subpath[grep("Parasitoid.traps", info_data$ColumnName)], split=", ")[[1]][3])
Parasitoid.traps <- rbind(AEGpt, HEGpt, SEGpt)
# store in list of raw datasets
list_raw_functions[["Parasitoid.traps"]] <- Parasitoid.traps ; rm(Parasitoid.traps); rm(AEGpt); rm(HEGpt); rm(SEGpt)
```


# Pathogen infection
note : is created from **assembled data**, available locally only.

**pathogen data** : 4 transects of 25x1 m were searched for foliar fungal pathogens on all grassland EP's. The prevalence and severity of each pathogen was estimated.

**plant cover data** : sampled all species in an area of 4m x 4m and estimated the percentage cover of each species.

The datasets are stored at the same place as other raw_datasets. 

note : code adapted from Eric Allan
```{r}
path <- paste(pathtodata, info_data$subpath[grep("pathogen.infection", info_data$ColumnName)], sep="")

pat <- read.table(paste(path, "BE.foliar pathogen grass.txt", sep=""), header =T)
# data is in relative abundances.
npla <- data.table::fread(paste(pathtodata, info_data$subpath[grep("plant.cover", info_data$ColumnName)][[1]], sep=""), header=T, sep="\t")
npla <- npla[Year == "2011", c("Useful_EP_PlotID", "Species", "Cover")]
data.table::setnames(npla, old = "Useful_EP_PlotID", new= "Plot")
npla[, Cover := as.numeric(Cover)]

list_raw_functions[["pathogen.infection"]] <- pat; rm(pat)
list_raw_functions[["plant.cover"]] <- npla; rm(npla)
rm(path)
```


# Total pollinators
The total amount of visits per plot will be taken as total pollinators function.
```{r}
Total.pollinators <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Total.pollinators", info_data$ColumnName)], sep=""), header=T, sep='\t')
Total.pollinators <- Total.pollinators[,c("EP_ID", "Total")]
data.table::setnames(Total.pollinators, old=c("EP_ID", "Total"), new=c("Plot", "Total.pollinators"))
list_raw_functions[["Total.pollinators"]] <- Total.pollinators; rm(Total.pollinators)
```



# Charismatic butterflys
Diversity and abundance of day active butterflies and day-active moths along a land use intensity gradient (2008)

```{r, eval=F}
butterflys <- data.table::fread(paste(pathtodata, info_data$subpath[grep("Charism.butterfly.Abund", info_data$ColumnName)], sep=""), header=T, sep='\t')
```

# amoA_AOA, AOB and nifH
Nitrifiers Abundances at Grassland, The abundance of ammonia oxidizing bacteria and archaea was estimated using respective amoA gene.

AOA : ammonia oxidizing archaea

AOB : ammonia oxidizing bacteria

nifH : (only 2011) nitrogen fixing locus

nxrA_NS and 16S_NB are present in nitrite-oxidizing bacteria (NOB)

2011 and 2016 datasets.
```{r}
# 2016
amo <- data.table::fread(paste(pathtodata, info_data$subpath[grep("amoA_AOB.2016", info_data$ColumnName)], sep=""), header=T, sep='\t')
data.table::setnames(amo, old=c("Plot_ID", "amoA_AOA", "amoA_AOB"), new=c("Plot", "amoA_AOA.2016", "amoA_AOB.2016"))
raw_functions <- merge(raw_functions, amo[,c("Plot","amoA_AOA.2016", "amoA_AOB.2016", "nxrA_NS", "16S_NB")], by="Plot", all = T); rm(amo)

# 2011
nit2011 <- data.table::fread(paste(pathtodata, info_data$subpath[grep("nifH", info_data$ColumnName)], sep=""), header=T, sep='\t')
data.table::setnames(nit2011, old=c("PlotID", "amoA_AOA", "amoA_AOB"), new=c("Plot", "amoA_AOA.2011", "amoA_AOB.2011"))
raw_functions <- merge(raw_functions, nit2011[, c("Plot", "amoA_AOA.2011", "amoA_AOB.2011", "nifH")], by="Plot", all=T) ; rm(nit2011)
```

# soil depth
soil.depth

<!-- Soils were described according to the German soil classification system (KA 5) and the FAO-Guidelines for Soil Description (2006). -->

<!-- A multitude of soil measures were taken. -->

<!-- theory : plants share the belowground space. A diverse set of plants can vary in their root length - different plants sometimes differ in the depth where they are able to get nutrients. The "potential rooting depth" is the soil depth in which plants can potentially have roots. Soil horizons O, A and B can harbour plant roots. -->

<!-- Mächtigkeit der obsersten bodenschicht herausfinden. -->
<!-- ```{r, eval=F} -->
<!-- z <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/5340_Soil_ Inventory organic soils grid plots_1.287.23_PublicDataset/5340_Block0.txt") -->
<!-- o <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/5340_Soil_ Inventory organic soils grid plots_1.287.23_PublicDataset/5340_Block1.txt") -->
<!-- t <- data.table::fread("/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/raw_data/5340_Soil_ Inventory organic soils grid plots_1.287.23_PublicDataset/5340_Block2.txt") -->

<!-- data.table::fread(paste(pathtodata, info_data$subpath[grep("Nshoot.2009", info_data$ColumnName)], sep=""), header=T, sep='\t') -->
<!-- ``` -->

using soil depth from Eric, who assembled it from dataset 5340 (as indicated in synthesis dataset 21688)

```{r}
soil.depth <- data.table::fread(paste(pathtodata, info_data$subpath[grep("soil.depth", info_data$ColumnName)], sep=""), header=T, sep=',')
data.table::setnames(soil.depth, old="Plot", new="Plotn")
raw_functions <- merge(raw_functions, soil.depth[,c("Plotn","soil.depth")], by="Plotn", all = T); rm(soil.depth)
```

# pH
soil PH

2 measures per site, take mean of them.
 
```{r}
ph <- data.table::fread(paste(pathtodata, info_data$subpath[grep("PH", info_data$ColumnName)], sep=""), header=T, sep='\t')
ph <- ph[Type == "G", c("EP_Plotid", "pH_1", "pH_2")]
ph[,"PH"] <- apply(ph[, c("pH_1", "pH_2")], 1, mean, na.rm=T)
data.table::setnames(ph, old="EP_Plotid", new="Plot")
raw_functions <- merge(raw_functions, ph[,c("Plot","PH")], by="Plot", all=T); rm(ph)
```


