---
title: "Select and Combine functions"
output:
  pdf_document: default
  html_notebook: default
---
Here, ecosystem functions which were read in in `1read_raw_dataset.Rmd` and calculated in `2calc_raw_dataset` are further processed. The functions needed for the project *BetaDivMultifun* are selected and assembled functions are created.


# Create assembled functions
Assembled functions were constructed according to E. Allan (cite project?).

## soilCflxs
Enzymes related to soil C. All measured in the year 2011.
```{r}
sce <- raw_functions[, c("beta_Glucosidase", "N_Acetyl_beta_Glucosaminidase", "Xylosidase")]
raw_functions[, "soilCflxs" := multidiv(sce, sc="sd", cent=TRUE)[,1]]
rm(sce)
```

## soilNflxs
Enzymes related to soil N. A measure of soil N cycling with enzyme activities and functional gene abundances

All from 2011
```{r}
nce <- raw_functions[,c("Urease", "DEA","Potential.nitrification","nifH","amoA_AOB.2011","amoA_AOA.2011")]
raw_functions[, "soilNflxs" := multidiv(nce, sc="sd", cent=TRUE)[,1]]
rm(nce)
```

## P_loss_comb


## N_loss_comb

# Select functions for BetaDivMultifun project