---
title: "assembled dataset construction"
author: "noelle"
date: "December 5, 2018"
output: html_document
---

# Biomass
2009 has 2 measures per plot
```{r}
#TODO find problem and solve (see compare to synthesis script)
```


# Root biomass
rb = fine root + coarse root biomass
```{r}
Root.biomass <- list_raw_functions$Root.biomass
Root.biomass[, "Root.biomass" := rowSums(.SD,na.rm=T), .SDcols=c("Fine_Roots_Biomass", "Coarse_Roots_Biomass")]
Root.biomass <- Root.biomass[, c("Plot", "Root.biomass")]
raw_functions <- merge(raw_functions, Root.biomass, by="Plot", all=T)
```

# Groundwater recharge
```{r}
Gtot <- list_raw_functions$Groundwater.recharge
raw_functions[, Groundwater.recharge := rowMeans(Gtot[,c(paste("Groundwater.recharge", seq(2010,2016), sep=""))])]
```


# forage quality
RFV : Relative Feed Value
RFQ : Relative Forage Quality
ADL : acid detergent lignin, Lignin
ADF : Acid Detergent Fibre, S채ure-Detergenz-Faser, ein Teil der Ger체stsubstanzen von Futtermitteln
NDF : Neutral Detergent Fibre, Neutral-Detergenz-Faser, fasrige Anteile von Futtermitteln 

according to forage quality estimation based on alfalfa (medicago sativa) values. [source](https://openprairie.sdstate.edu/cgi/viewcontent.cgi?referer=https://scholar.google.ch/&httpsredir=1&article=1351&context=extension_extra)

[% concentration in aboveground grassland plant community biomass without litter]

```{r}
# calculating the columns needed for forage quality calc
forage.quality[, Crude.protein := 6.25*N]
forage.quality[, DDM := 88.9-(0.779* ADF)]
forage.quality[, DMI := 120/NDF]
forage.quality[, RFV := (DDM*DMI)/1.29]
forage.quality[, forage.quality := apply(scale(forage.quality[,c("RFV", "Crude.protein")]), 1, mean)]
```
See in compare_raw_function_to_synthesis for yearly calculation!

# Nshoot and Pshoot
same dataset as forage.quality
```{r}
nutrients <- list_raw_functions$nutrients
#TODO : decide if this values are taken or not. BUT : there is something wrong with Biomass 2009
# multiply with biomass?
# nutrients <- list_raw_functions$nutrients
# d09 <- merge(Biomass2009, nutrients, by="Plot", all=T)
# d09[, Nshoot.2009 := Biomass2009 * (Nshoot2009percent / 100)]
# bal <- merge(d09, synthesis[,c("Plot", "Nshoot.2009")])
# # maximum deviation of 6
# max(bal$Nshoot.2009.x - bal$Nshoot.2009.y, na.rm=T)
# # the larger the total values, the larger the deviation. Could be acceptable?
# plot(bal$Nshoot.2009.x, bal$Nshoot.2009.x - bal$Nshoot.2009.y)

# calculate values, but use with caution.

# 2010
assemblednutrients <- merge(Biomass2010, nutrients, by="Plot", all=T)
assemblednutrients[, Nshoot.2010 := Biomass2010 * (Nshoot2010percent / 100)]
assemblednutrients[, Pshoot.2010 := Biomass2010 * (Pshoot2010percent / 100)]
# add to raw_functions
raw_functions <- merge(raw_functions, assemblednutrients[,c("Plot", "Nshoot.2010", "Pshoot.2010")], by="Plot")


# 2011
assemblednutrients <- merge(Biomass2011, nutrients, by="Plot", all=T)
assemblednutrients[, Nshoot.2011 := Biomass2011 * (Nshoot2011percent / 100)]
assemblednutrients[, Pshoot.2011 := Biomass2011 * (Pshoot2011percent / 100)]
# add to raw_functions
raw_functions <- merge(raw_functions, assemblednutrients[,c("Plot", "Nshoot.2011", "Pshoot.2011")], by="Plot")

# 2012
# TODO
```


# Herbivory
## 2013
invertebrate herbivory in 146 managed temperate grasslands. Year: 2013

increased land-use - less invertebrates - less herbivory

Visually estimation of herbivory by eye using templates and measurement of leaf area left after feeding of the herbivores using LI-COR area meter

2 samples per plot in 145 cases, 1 sample for :  "AEG6"  "HEG16" "HEG43" "SEG19" "SEG33" and 3 samples for "SEG45". 2 Columns are included in the raw_functions dataset, 1 for each sample.

SEG45 : The mean of all 3 samples was already taken and stored as sample2. Taking the mean of samples2 first and then with sample1 would lead to different results than in the synthesis dataset.
```{r}
Herbivory.2013 <- list_raw_functions$Herbivory.2013

Herbivory.2013 <- aggregate(PercHerbivory~EP_Plotid, Herbivory.2013, mean)
data.table::setnames(Herbivory.2013, old=c("EP_Plotid", "PercHerbivory"), new=c("Plot", "herbivory"))
raw_functions <- merge(raw_functions, Herbivory.2013, by="Plot", all.x=T)
```

# Litter.decomposition
litter bag approach, Mass loss in the litter bags, after two months, and four months. 1.5g of dry plant material. January 2013 put 10 bags each, 2 months later retrieval of 5 bags, 2 more months : retrieval of other 5 bags. calculated daily decomposition rates.
```{r}
Litter.decomposition <- list_raw_functions$Litter.decomposition
# take the mean daily decomposition rate over all measures, the ones with 2 months and the ones with 4 months decomposition.
Litter.decomposition <- aggregate(DailyDecompositionRate~EPPlotID,Litter.decomposition,mean)
data.table::setnames(Litter.decomposition, old=c("EPPlotID", "DailyDecompositionRate"), new=c("Plot", "Litter.decomposition"))
raw_functions <- merge(raw_functions, Litter.decomposition, by="Plot", all.x = T)
```


# Parasitoid traps
If there are more parasitoid wasps, they predate caterpillars which eat the plants.

In spring 2008, a total of 95 grassland plots were selected in Hainich-D체n, Schorfheide-Chorin and Schw채bische Alb. After collecting the traps at the end of September, nests of trap-nesting bees, wasps and their natural enemies were collected and recorded. The experimental grassland plots were chosen according to their different land use types and their distance from each other.

Cat : Replace by a better measure (caterpillars predation not found in bexis)

3 datasets (13350, 13347, 13349) for 3 regions: 34 from AEG, 28 from HEG and 34 from SEG - 96 plots measured

recorded total broodcell number : Total_BC_No
```{r}
# get dataset from raw functions list
pt <- list_raw_functions$Parasitoid.traps
pt <- pt[Function == "Parasitoid", c("PlotID", "Total_BC_No")]
pt <- aggregate(Total_BC_No~PlotID, pt, sum)
data.table::setnames(pt, old=c("PlotID", "Total_BC_No"), new=c("Plot", "Parasitoid.traps"))
# append to raw functions list
raw_functions <- merge(raw_functions, pt, by="Plot", all.x = T)
#TODO maybe directly append to assembled functions list?
```

# Total pollinators
Plot number AEG24 has 939 total pollinators in synthesis dataset, but only 339 total pollinators in the downloaded dataset. All other numbers are perfectly identical.

The number from the downloaded dataset is taken, but the change is recorded.
```{r}
Total.pollinators <- list_raw_functions$Total.pollinators
Total.pollinators <- aggregate(Total.pollinators~Plot, Total.pollinators, sum)
raw_functions <- merge(raw_functions, Total.pollinators, by="Plot", all=T)
```





# store assembled dataset
store as .csv document
```{r}
write.csv(raw_functions, file="~/Desktop/december2018_assembled_functions_dataset.csv")
```

