---
title: "assembled dataset construction"
author: "noelle"
date: "December 5, 2018"
output: html_document
---


# Root biomass
rb = fine root + coarse root biomass
```{r}

Root.biomass[, "Root.biomass" := rowSums(.SD,na.rm=T), .SDcols=c("Fine_Roots_Biomass", "Coarse_Roots_Biomass")]
Root.biomass <- data.table::setnames(Root.biomass[,c("EP_Plotid", "Root.biomass")], old=c("EP_Plotid", "Root.biomass"), new=c("Plot", "Root.biomass"))
```

# forage quality
RFV : Relative Feed Value
RFQ : Relative Forage Quality
ADL : acid detergent lignin, Lignin
ADF : Acid Detergent Fibre, Säure-Detergenz-Faser, ein Teil der Gerüstsubstanzen von Futtermitteln
NDF : Neutral Detergent Fibre, Neutral-Detergenz-Faser, fasrige Anteile von Futtermitteln 

according to forage quality estimation based on alfalfa (medicago sativa) values. [source](https://openprairie.sdstate.edu/cgi/viewcontent.cgi?referer=https://scholar.google.ch/&httpsredir=1&article=1351&context=extension_extra)

[% concentration in aboveground grassland plant community biomass without litter]

```{r}
# calculating the columns needed for forage quality calc
forage.quality[, Crude.protein := 6.25*N]
forage.quality[, DDM := 88.9-(0.779* ADF)]
forage.quality[, DMI := 120/NDF]
forage.quality[, RFV := (DDM*DMI)/1.29]
forage.quality[, forage.quality := apply(scale(forage.quality[,c("RFV", "Crude.protein")]), 1, mean)]
```
See in compare_raw_function_to_synthesis for yearly calculation!
