---
title: "prepare_covariates"
author: "noelleschenk"
date: "April 16, 2019"
output: html_document
---

- PCA of soils
- prepare covariates for analysis 

Depends on `clean_and_load_soil_covariates.Rmd` and `analysis_nonpublic.R`.

# Soil
```{r}
# copy to keep glsoil variable
soilcov <- data.table::copy(glsoil)
```


## Normalisation
scale the columns which are neither binary nor factors. Scaled by centering around 0 (substracting the mean) and dividing by standard deviation.
```{r}
# take out soil type as this is converted to dummy variables
soilcov[, Soil.type := NULL]
# create scaling vector T/F of which columns to scale in pcr
scalecols <- c("Clay", "Fine.Silt", "Medium.Silt", "Coarse.Silt", "Fine.Sand", "Medium.Sand", "Coarse.Sand", "pH", "Soil.depth", "TWI")
soilcov[, (scalecols) := lapply(.SD, function(x) scale(x, center = T)), .SDcols = scalecols]
```

Could be done within PCA computation with parameter `scale = TRUE`.
```{r}
mysubset <- names(soilcov)[!names(soilcov) %in% c("Plotn", "Soil.type")]
```
*note* : The binary categoric columns are not affected by this.
```{r, eval=F}
plot(soilcov$Rendzina)
points(soilcov$Braunerde)
points(soilcov$Pelosol)
points(soilcov$Pseudogley)
points(soilcov$Erdniedermoor)
points(soilcov$Gley)
points(soilcov$Mulmniedermoor)
points(soilcov$Parabraunerde)
points(soilcov$Fahlerde)
```


## soil PCA
```{r}
# Plot information is taken as rownames of a data.frame (instead of data.table).
pc_glsoil <- as.data.frame(soilcov[, !"Plotn", with=FALSE])
rownames(pc_glsoil) <- soilcov$Plotn
rm(soilcov)

# run the pca
pca_from_glsoil <- prcomp(pc_glsoil[, !(names(pc_glsoil) %in% "Soil.type")], scale = F)
plot(pca_from_glsoil) 
```
The first 4 axis would make sense if the dimensionality would need to be reduced a lot. 

However, here it is not that much about reducing dimensionality, but taking out unnecessary variables. Therefore, Taking PC1 to 14 is suitable, as PC 15 to 19 do not add any more information.
```{r}
# first 4 axis would make sense
summary(pca_from_glsoil)
```

### PCA insights
```{r, eval=F}
# color by soil type
ggplot2::autoplot(pca_from_glsoil, data = pc_glsoil, colour = 'Soil.type')
ggplot2::autoplot(pca_from_glsoil, data = pc_glsoil, colour = 'Soil.depth')

ggplot2::autoplot(pca_from_glsoil, data = pc_glsoil, loadings = TRUE, loadings.colour = 'darkgray', loadings.label = TRUE, loadings.label.size = 2)
```

```{r, eval=F}
ggplot2::autoplot(pca_from_glsoil, x=3, y=4, data=pc_glsoil, colour="Soil.type")
ggplot2::autoplot(pca_from_glsoil, x=3, y=4, data=pc_glsoil, colour="pH")
```

### Component selection
Components 1 to 14 (as argued above).
```{r}
pca_from_glsoil <- pca_from_glsoil$x[, 1:14]
```

## Computation of Dissimilarity Matrices
With euclidian distance : `edis_glsoil`

note : betadiversity is not calculated, but in case : 
With betadiversity : `beta_glsoil`.
*note* beta.pair.abund can not handle negative abundance values. Therefore, data needs to be centered around 0 instead of centered around 0. However, max is `r max(pca_from_glsoil)` and min is `r min(pca_from_glsoil)`. Would need an "abundance" value which makes sense.

```{r}
# select plot set from matrix
pca_from_glsoil <- data.frame(pca_from_glsoil)
pca_from_glsoil <- pca_from_glsoil[which(rownames(pca_from_glsoil) %in% plotNAset), ]
# pca_from_glsoil <- pca_from_glsoil[which(rownames(pca_from_glsoil) != "HEG31"), ] # not needed if plotNA set is in the first version. only run this line in case this is the first time you are running that script.

# compute dissimilarities for the first 14 principal components (euclidian distance)
edis_glsoil <- vegan::vegdist(pca_from_glsoil, method = "euclid")

# # In case I would want to calc beta, I would need to find good solution for handling negative values. Don't think so
# beta_glsoil <- betapart::beta.pair.abund(pca_from_glsoil)
rm(pc_glsoil); rm(pca_from_glsoil)
```

## Cleaning for GDM
```{r}
# prepare for gdm (the desired format)
# edis_glsoil <- as.matrix(edis_glsoil)
# edis_glsoil <- cbind(row.names(edis_glsoil), edis_glsoil)

# bring to long format for GDM and add to edis_glsoil table
edis_glsoil <- as.matrix(edis_glsoil)
edis_glsoil[!lower.tri(edis_glsoil)] <- NA
edis_glsoil <- reshape2::melt(edis_glsoil, value.name = "edis_soil")
edis_glsoil <- edis_glsoil[!is.na(edis_glsoil[, 3]),]
nrow(edis_glsoil) == choose(127, 2)

# store soil dissimilarity matrix separately, as storing diss object in a list as rds is not possible
saveRDS(edis_glsoil, "predictors_soil.rds")
```

# Plot isolation
```{r}
plt.sur <- unique(plt.sur) #for some areas each row has two values
plt.sur <- plt.sur[,c("EP_PLOTID","G500")]
data.table::setnames(plt.sur, "EP_PLOTID", "Plot")
plt.sur <- merge(plt.sur, usefulplotids, by="Plot")
plt.sur[, PLAND_G_500 := 100 - (G500 *100)] # change so more isolated plots have higher values
plt.sur[, G500 := NULL] ; plt.sur[, Plot := NULL] # remove unused columns
# select plot set
plt.sur <- plt.sur[Plotn %in% plotNAset,]
plt.sur <- as.data.frame(plt.sur)# convert to datatable for analysis
rownames(plt.sur) <- plt.sur[["Plotn"]]
plt.sur <- plt.sur["PLAND_G_500"]
colnames(plt.sur) <- "plot_isolation"
# in old code, plt.sur is now as column in a variable called "predictors for GDM preparation"
```

# geographic distance
```{r}
plot.names <- names_gl[Landuse == "G", .(EP_PlotID, PlotID)]
data.table::setnames(plot.names, old = "EP_PlotID", new = "Plot")
plot.names <- merge(plot.names, usefulplotids, by = "Plot", all = T)

geodist <- geodist[Landuse == "G", .(Plot_ID, Longitude_Dec_Plotcenter, Latitude_Dec_Plotcenter)]
data.table::setnames(geodist, "Plot_ID", "PlotID")
geodist <- merge(geodist, plot.names, by = "PlotID")
geodist[, c("PlotID", "Plot") := NULL]
# remove two duplicated rows
# geodist[which(duplicated(geodist$Plotn)), ]
# geodist[Plotn %in% c("AEG29", "AEG10"), ] # there is only very little difference - remove one of them
geodist <- geodist <- geodist[!which(duplicated(geodist$Plotn)),]
# select plot set
geodist <- geodist[Plotn %in% plotNAset]
data.table::setkey(geodist,Plotn)
geodist <- data.frame(geodist)
rownames(geodist) <- geodist$Plotn
geodist <- geodist[, c("Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")]
rm(plot.names)
```

# combine to predictors
combine covariates to a predictor dataset
```{r, eval=F}
predictors <- merge(plt.sur, geodist, by = "row.names")

saveRDS(predictors, file = "predictors.rds")

rm(edis_glsoil); rm(plt.sur); rm(geodist)
```
