---
title: "prepare_covariates"
author: "noelleschenk"
date: "April 16, 2019"
output: html_document
---

- PCA of soils
- prepare covariates for analysis 

Depends on `clean_and_load_soil_covariates.Rmd` and `analysis_nonpublic.R`.
```{r}
# run clean_and_load_soil_covariates.Rmd
sections_to_be_loaded <- c("assemble_covariates")
source("vignettes/analysis_nonpublic.R")
```


# Soil
```{r}
# copy to keep glsoil variable
soilcov <- data.table::copy(glsoil)
```


## Normalisation

scale the columns which are neither binary nor factors. Scaled by centering around 0 (substracting the mean) and dividing by standard deviation.
```{r}
# take out soil type as this is converted to dummy variables
soilcov[, Soil.type := NULL]
# create scaling vector T/F of which columns to scale in pcr
scalecols <- c("Clay", "Fine.Silt", "Medium.Silt", "Coarse.Silt", "Fine.Sand", "Medium.Sand", "Coarse.Sand", "pH", "Soil.depth", "TWI")
soilcov[, (scalecols) := lapply(.SD, function(x) scale(x, center = T)), .SDcols = scalecols]
# note : Could be done within PCA computation with parameter `scale = TRUE`.

mysubset <- names(soilcov)[!names(soilcov) %in% c("Plotn", "Soil.type")]
```
*note* : The binary categoric columns are not affected by this.
```{r, eval=F}
plot(soilcov$Rendzina)
points(soilcov$Braunerde)
points(soilcov$Pelosol)
points(soilcov$Pseudogley)
points(soilcov$Erdniedermoor)
points(soilcov$Gley)
points(soilcov$Mulmniedermoor)
points(soilcov$Parabraunerde)
points(soilcov$Fahlerde)
```


## soil PCA

From 11 variables, 1 of those (soil type) is split into 9 variables --> 19 columns.
```{r}
# Plot information is taken as rownames of a data.frame (instead of data.table).
pc_glsoil <- as.data.frame(soilcov[, !"Plotn", with=FALSE])
rownames(pc_glsoil) <- soilcov$Plotn
rm(soilcov)

# run the pca
pca_from_glsoil <- prcomp(pc_glsoil[, !(names(pc_glsoil) %in% "Soil.type")], scale = F)
plot(pca_from_glsoil)

temp <- cbind(pc_glsoil, "region" = sub("[0-9][0-9]", "", rownames(pc_glsoil)))
pcaplots <- list(autoplot(pca_from_glsoil, data = temp, colour = "region"),
                 autoplot(pca_from_glsoil, data = temp, colour = "region", x = 3, y = 4),
                 autoplot(pca_from_glsoil, data = temp, colour = "region", x = 5, y = 6),
                 autoplot(pca_from_glsoil, data = temp, colour = "region", x = 7, y = 8))

pcaplots2 <- list(autoplot(pca_from_glsoil, data = temp, colour = "pH"),
                  autoplot(pca_from_glsoil, data = temp, colour = "pH", label = T, label.size = 2, shape = F, loadings = T),
                 autoplot(pca_from_glsoil, data = temp, colour = "pH", loadings = T, loadings.label = T),
                 autoplot(pca_from_glsoil, data = temp, colour = "pH", x =3, y = 4, loadings = T, loadings.label = T))
plot_grid(plotlist = pcaplots, nrow = 2)
plot_grid(plotlist = pcaplots2, nrow = 2)
# save plots as : "<date>_calc_covariates_soil_pca1.pdf"
rm(temp); rm(pcaplots); rm(pcaplots2)
```
Take all axis until 90% cumulative variance explained.

We follow the philosophy that a PCaxis is useful, when it explains more variance than original variables. This can be done by picking all PC axes that explain more variance than original functions, i.e. all those axes which explain more than 1/n % `1/11*100` % of variance, in which n is the number of input variables.

```{r}
summary(pca_from_glsoil) # take axes 1 - 7
```

### PCA insights
```{r, eval=F}
# color by soil type
ggplot2::autoplot(pca_from_glsoil, data = pc_glsoil, colour = 'TWI')
ggplot2::autoplot(pca_from_glsoil, data = pc_glsoil, colour = 'Soil.depth')
ggplot2::autoplot(pca_from_glsoil, data = pc_glsoil, loadings = TRUE, loadings.colour = 'darkgray', loadings.label = TRUE, loadings.label.size = 2)
ggplot2::autoplot(pca_from_glsoil, x=3, y=4, data=pc_glsoil, colour="TWI")
ggplot2::autoplot(pca_from_glsoil, x=3, y=4, data=pc_glsoil, colour="pH")
```

### Component selection
Components 1 to 7 (as argued above).
```{r}
pca_from_glsoil <- pca_from_glsoil$x[, 1:7]
```

## Computation of Dissimilarity Matrices
With euclidian distance : `edis_glsoil`

```{r}
# select plot set from matrix
pca_from_glsoil <- data.frame(pca_from_glsoil)
pca_from_glsoil <- pca_from_glsoil[which(rownames(pca_from_glsoil) %in% plotNAset), ]
# pca_from_glsoil <- pca_from_glsoil[which(rownames(pca_from_glsoil) != "HEG31"), ] # not needed if plotNA set is in the first version. only run this line in case this is the first time you are running that script.

# compute dissimilarities for the first 14 principal components (euclidian distance)
edis_glsoil <- vegan::vegdist(pca_from_glsoil, method = "euclid")

rm(pc_glsoil); rm(pca_from_glsoil)
```

## Cleaning for GDM
```{r}
# prepare for gdm (the desired format)
# edis_glsoil <- as.matrix(edis_glsoil)
# edis_glsoil <- cbind(row.names(edis_glsoil), edis_glsoil)

# bring to long format for GDM and add to edis_glsoil table
edis_glsoil <- as.matrix(edis_glsoil)
edis_glsoil[!lower.tri(edis_glsoil)] <- NA
edis_glsoil <- reshape2::melt(edis_glsoil, value.name = "edis_soil")
edis_glsoil <- edis_glsoil[!is.na(edis_glsoil[, 3]),]
nrow(edis_glsoil) == choose(127, 2)

# store soil dissimilarity matrix separately if needed
saveRDS(edis_glsoil, "predictors_soil.rds")
```

# Non-soil predictors

## DEPRECATED Plot isolation
```{r, eval = F}
plt.sur <- unique(plt.sur) #for some areas each row has two values
plt.sur <- plt.sur[,c("EP_PLOTID","G500")]
data.table::setnames(plt.sur, "EP_PLOTID", "Plot")
plt.sur <- merge(plt.sur, usefulplotids, by="Plot")
plt.sur[, PLAND_G_500 := 100 - (G500 *100)] # change so more isolated plots have higher values
plt.sur[, G500 := NULL] ; plt.sur[, Plot := NULL] # remove unused columns
# select plot set
plt.sur <- plt.sur[Plotn %in% plotNAset,]
plt.sur <- as.data.frame(plt.sur)# convert to dataframe for analysis
rownames(plt.sur) <- plt.sur[["Plotn"]]
plt.sur <- plt.sur["PLAND_G_500"]
colnames(plt.sur) <- "plot_isolation"
# in old code, plt.sur is now as column in a variable called "predictors for GDM preparation"
```

## historic/ landscape LUI and landscape covariates

- historic LUI : Grassland permanency. Measure of historic land-use in 1 km radius around the plot. Is the ratio of the mean grassland cover from 1820/50 to 2008 to the standard deviation over that period. $\frac{\text{mean grl cover y}_i}{\text{stdev grl cover y}_i} $ High permanency : high grassland cover over time which did not fluctuate. Means that the plot has been used since a long time. (Maybe extinction debt already paid)
- landscape LUI : % arable land in 1 km radius around the plot. (is more suitable than plot isolation which was used in earlier versions). High values indicate more of the surrounding area is arable land --> hypothetically less suitable habitat around.
- landscape diversity : number of different types and abundances of land-use types. High values indicate a lot of different land-uses around the plots.
- Landscape fragmentation : total edges. High values indicate high fragmentation.

```{r}
landscape <- data.table(BEplotZeros(landscape, column = "EP_PlotID", plotnam = "EP"))
setnames(landscape, old = c("EP_PlotID", "EP"), new = c("oldEP", "Plotn"))
landscape <- landscape[, .(Plotn, Grassland.perm.1000, Arable.1000, Shannon.1000)]
landscape <- landscape[Plotn %in% plotNAset, ]

landscape2 <- data.table(BEplotZeros(landscape2, column = "plot_id", plotnam = "EP"))
landscape2[, plot_id := NULL]
setnames(landscape2, old = c("EP"), new = c("Plotn"))
landscape2 <- landscape2[Year == 2012 & scale == 1000, .(Plotn, Land.te)]
setnames(landscape2, old = "Land.te", new = "Land.te.1000")

landscape <- merge(landscape, landscape2, by = "Plotn")
rm(landscape2)

# convert to data.frame
data.table::setkey(landscape, Plotn)
landscape <- data.frame(landscape)
rownames(landscape) <- landscape$Plotn
landscape <- landscape[, c("Grassland.perm.1000", "Arable.1000", "Shannon.1000", "Land.te.1000")]
```


## geographic distance
```{r}
plot.names <- names_gl[Landuse == "G", .(EP_PlotID, PlotID)]
data.table::setnames(plot.names, old = "EP_PlotID", new = "Plot")
plot.names <- merge(plot.names, usefulplotids, by = "Plot", all = T)

geodist <- geodist[Landuse == "G", .(Plot_ID, Longitude_Dec_Plotcenter, Latitude_Dec_Plotcenter)]
data.table::setnames(geodist, "Plot_ID", "PlotID")
geodist <- merge(geodist, plot.names, by = "PlotID")
geodist[, c("PlotID", "Plot") := NULL]
# remove two duplicated rows
# geodist[which(duplicated(geodist$Plotn)), ]
# geodist[Plotn %in% c("AEG29", "AEG10"), ] # there is only very little difference - remove one of them
# note : in the newest dataset BExIS ID 1000, plot coordinates are given as selected here. 
#    (48.38/9.21 for AEG10 and 48.42/9.36 for AEG29)
geodist <- geodist <- geodist[!which(duplicated(geodist$Plotn)),]
# select plot set
geodist <- geodist[Plotn %in% plotNAset]
data.table::setkey(geodist,Plotn)
geodist <- data.frame(geodist)
rownames(geodist) <- geodist$Plotn
geodist <- geodist[, c("Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")]
rm(plot.names)
```


## save as predictors

combine covariates to a predictor dataset
```{r, eval=T}
predictors <- merge(landscape, geodist, by = "row.names")
colnames(predictors)[1] <- "Plotn"

saveRDS(predictors, file = "predictors.rds")

rm(edis_glsoil); rm(landscape); rm(geodist)
```
