---
title: "mini-multifunctionalities from imputed functions"
author: "Noelle Schenk"
date: "June 5, 2019"
output: html_document
---

# Calculate mini-multifunctionalities
The mini-multifunctionalities are calculated from the imputed values.

## soilCflxs
Enzymes related to soil C. All measured in the year 2011.
```{r}
sce <- imputed_grlfuns[, c("beta_Glucosidase", "N_Acetyl_beta_Glucosaminidase", "Xylosidase")]
# take z-scores of the functions
imputed_grlfuns[, "soilCflxs" := multidiv(sce, sc="sd", cent=TRUE)[,1]]
rm(sce)
```

## soilNflxs
Enzymes related to soil N. A measure of soil N cycling with enzyme activities and functional gene abundances

amoA_AOB and A are from both 2011 and 2016, took the mean of both years.
```{r}
nce <- imputed_grlfuns[,c("Urease", "DEA","Potential.nitrification2011", "Potential.nitrification2014","nifH","amoA_AOB.2011","amoA_AOA.2011", "amoA_AOB.2016", "amoA_AOA.2016", "nxrA_NS", "16S_NB")]
# take mean of multi-year-measurements
nce[, "amoA_AOB" := apply(nce[,c("amoA_AOB.2011", "amoA_AOB.2016")],1, mean)]
nce[, "amoA_AOA" := apply(nce[,c("amoA_AOA.2011", "amoA_AOA.2016")],1, mean)]
nce[, "Potential.nitrification" := apply(nce[, c("Potential.nitrification2014", "Potential.nitrification2011")], 1, mean)]
nce[, c("amoA_AOB.2011", "amoA_AOB.2016", "amoA_AOA.2011", "amoA_AOA.2016", "Potential.nitrification2014", "Potential.nitrification2011") := NULL]

imputed_grlfuns[, "soilNflxs" := multidiv(nce, sc="sd", cent=TRUE)[,1]]
rm(nce)
```


# Select functions
selecting the functions which are used in analysis.
```{r}
# select variables which are included in betadiv multifun analysis
usedforBetadivMultifun <- usedforBetadivMultifun[!usedforBetadivMultifun == "Plot"]
imputed_grlfuns <- imputed_grlfuns[, ..usedforBetadivMultifun]
#TODO
```







# Correlations
Of the non-factor columns only
```{r}
# the dataset is complete now, use "complete.obs" for the cor() function now
M <- cor(imputed_grlfuns[, !colnames(imputed_grlfuns) %in% "Plotn", with=F], use="complete.obs")

# imputed_grlfun_corrplot1
corrplot::corrplot(M,type="lower",addCoef.col = "black",method="color",diag=F, tl.srt=1, tl.col="black", mar=c(0,0,0,0), number.cex=0.6)

# imputed_grlfun_corrplot2
corrplot::corrplot(M, type = "upper", tl.col="black", tl.srt=40, diag = F)

# only show high correlations
tres <- 0.7
M[which(M < tres & M > -tres)] <- 0
corrplot::corrplot(M, type = "upper", tl.col="black", tl.srt=40, diag = F)
```
- PRI and NRI: constructed partly with same data
- PRI and P_loss_comb : the latter contains the first

- Groundwater recharche and soil depth: ? the deeper the soil, the less groundwater recharche? soil depth defines the soil type. 

- SoilOrganicC and Soil.C.stock : 

old stuff
```{r}


####################
# 5. IMPUTATION
# Impute
imputeall<-missForest(befun[,-1])
imp_error <- imputeall$OOBerror

#extract imputed data and retransform everything
impdat <- data.table(Plot0=befun$Plot, exp(imputeall$ximp[,-35])-1, Exploratory=imputeall$ximp[,35])
impdat$pathogen.infection <- impdat$pathogen.infection-abs(minpath)
impdat$forage.quality <- impdat$forage.quality-abs(minfor)

# exclude all columns that are not functions : see script "calculate_multifunctionality_dissimilarity_DEMO.r

#################
# 6. CLEANING
#################
# deleting all stored variables which I will not use any more in the analysis.
rm(befun) ; rm(M) ; rm(imputeall)  ; rm(funs) ; rm(minpath) ; rm(minfor)



# ##########################################
# # 7. DIFFERENT WAY OF ERROR ESTIMATION
# # Check if missForest is doing a good job
# # read functions table
# # befun is the table I use.
# 
# # calculate  % of missing data in each function (i.e. column)
# NA_per_column <- apply(befun,2,function(c) sum(is.na(c))/length(c))
# # remove NA's from the table (rows) - so you have a dataset with no NAs
# 
# navals <- apply(befun, 1, function(r) any(is.na(r)))
# noNA_befun <- befun[!navals,] ; rm(navals)
# original_noNA_befun <- noNA_befun
# 
# # randomly remove the same % in each column (as we had in the real dataset)
# settoNA <- round(NA_per_column*nrow(noNA_befun)) ; rm(NA_per_column) # the amount of values to set to NA
# whichrowsNA <- list()
# for(i in 1:length(settoNA)){
#   rows <- sample(1:nrow(noNA_befun), settoNA[[i]], replace=F) # get random rows to set to NA
#   noNA_befun[rows,i] <- NA
#   # save which rows were set to NA where
#   whichrowsNA[[i]] <- rows
# }
# rm(i) ; rm(rows) ; rm(settoNA)
# 
# # impute with missForest
# imputeall<-missForest(noNA_befun[,-1])
# imputeall$OOBerror # 12% of estimated error
# 
# check_imp <- as.data.frame(imputeall$ximp)
# 
# # compare imputed with real data (calculate NRMSE=normalized root mean square error value)
# nrmse(imputeall$ximp, noNA_befun[,-1], original_noNA_befun[,-1])
#
# rm(noNA_befun)
# ##########
# # visualization
# original_noNA_befun <- as.data.frame(original_noNA_befun)
# imp_no <- c(check_imp[whichrowsNA[[22]],22],check_imp[whichrowsNA[[18]],18], check_imp[whichrowsNA[[17]],17],
#             check_imp[whichrowsNA[[16]],16] , check_imp[whichrowsNA[[19]],19])
# rm(check_imp)
# true_no <- c(original_noNA_befun[whichrowsNA[[22]],22],original_noNA_befun[whichrowsNA[[18]],18],original_noNA_befun[whichrowsNA[[17]],17],
#              original_noNA_befun[whichrowsNA[[16]],16], original_noNA_befun[whichrowsNA[[19]],19])
# rm(original_noNA_befun) ; rm(whichrowsNA)
# imp_true_no <- cbind(imp_no,true_no, c(rep(22,11),rep(18,8),rep(17,8),rep(16,6), rep(19,7))) ; rm(imp_no) ; rm(true_no)
# plot(imp_true_no[,1],imp_true_no[,2] ,main="imputed versus true", col=imp_true_no[,3],pch=19,
#      xlab="imputed", ylab="true")
# rm(imp_true_no)
# abline(a=0, b=1,col="grey")
# # without standardization, there is a clustering (because of the different range the values are in)
# # with standardization, the clusters disappear.

```

