---
title: "run GDM"
author: "Noelle Schenk"
date: "June 21, 2019"
output: html_document
---
Requirements : 
- script is runned right after `prepare_and_run_GDM.Rmd`. Requires its output `gdminput`.
    - note that `gdminput` is also saved as .csv and .Rds from the above mentioned script.
    - `nicenames` for nice correlation plot

The script calculates euclidean distances abs(simple substraction) of the covariates and LUI as well as the geographic distance based on coordinates and forms a table with all distances. A correlation plot is produced and a *linear model* fitted and both are saved. Variance inflation of the linear model is calculated and saved.

DEPRECIATED notice : linear model not done any more : assumption of independent observations (i.i.d. assumption) not met and therefore p-values not trustable. Also : nonlinearities are exptected.
GAM used to check, but it also violates the i.i.d. assumption.
Individual associations (correlation plot) does not correct for other predictors.
Need to check associations among variables and GDM assumption of monotonic increase.
--> all methods violate important assumptions --> compare the 3 methods and make the best of it.

Output : 
- correlation plot `vignettes/out/<modelname>_corrplot_before_gdm.pdf`
- DEPRECIATED linear model output `vignettes/out/<modelname>_linear_model_before_gdm.Rds`
- DEPRECIATED variance inflation `out/<modelname>_vif_before_gdm.txt`
- DEPRECIATED linear model plot `vignettes/out/<modelname>_linear_model_directions.pdf`

# creating dataset
```{r}
# first run prepare_and_run_GDM
lui <- c("LUI", "deltaLUI")
checkgdminput <- data.table::copy(gdminput)
# checkgdminput$rown <- rownames(checkgdminput)
checkgdminput <- data.table(checkgdminput)

# calc euclidean distance between each predictor pair of checkgdminput
# diversity : take out s2. because they are all 0
dists <- colnames(checkgdminput)[grep("beta", colnames(checkgdminput))]
dists <- grep("s2.", dists, value = T)
checkgdminput[, (dists) := NULL]
checkgdminput[, weights := NULL]

# LUI
luicols <- grep(paste(lui, collapse="|"), colnames(checkgdminput), value=TRUE)
if(length(grep("Gstd", lui)) > 1){
  print("lui components")
  checkgdminput[, Gstd := abs(s1.Gstd - s2.Gstd)]
  checkgdminput[, Mstd := abs(s1.Mstd - s2.Mstd)]
  checkgdminput[, Fstd := abs(s1.Fstd - s2.Fstd)]
  checkgdminput[, deltaGstd := abs(s1.deltaGstd - s2.deltaGstd)]
  checkgdminput[, deltaMstd := abs(s1.deltaMstd - s2.deltaMstd)]
  checkgdminput[, deltaFstd := abs(s1.deltaFstd - s2.deltaFstd)]
  checkgdminput[, (luicols) := NULL]
} else if(length(grep("LUI", lui)) > 1) {
  print("LUI")
  checkgdminput[, LUI := abs(s1.LUI - s2.LUI)]
  checkgdminput[, deltaLUI := abs(s1.deltaLUI - s2.deltaLUI)]
  checkgdminput[, meanLUI := rowMeans(.SD), .SDcols = c("s1.LUI", "s2.LUI")]
  checkgdminput[, sdLUI := apply(.SD, 1, sd), .SDcols = c("s1.LUI", "s2.LUI")]
  checkgdminput[, (luicols) := NULL]
}

checkgdminput[, geosphere_distgeo := geosphere::distGeo(
  p1 = as.matrix(checkgdminput[, .(s1.yCoord, s1.xCoord)]),
  p2 = as.matrix(checkgdminput[, .(s2.yCoord, s2.xCoord)])
  )]
geocol <- grep("Coord", colnames(checkgdminput), value = T)
checkgdminput[, (geocol) := NULL]
# checkgdminput[, s2.geosphere_distgeo := NULL]

# predictors
checkgdminput[, edis_soil := abs(s1.edis_soil - s2.edis_soil)]
checkgdminput[, plot_isolation := abs(s1.plot_isolation - s2.plot_isolation)]
del <- c("s1.edis_soil", "s2.edis_soil", "s1.plot_isolation", "s2.plot_isolation", "s1.geosphere_distgeo", "s2.geosphere_distgeo")
checkgdminput[, (del) := NULL]

# rename s1 from diversity columns
colnames(checkgdminput) <- gsub("s1.", "", colnames(checkgdminput))

# take out meanLUI and sdLUI if necessary
delcols <- c("meanLUI", "sdLUI")
checkgdminput[, (delcols) := NULL]
# # OR take out LUI and deltaLUI
# checkgdminput[, (lui) := NULL]

#TODO save dataset checkgdminput?
# saveRDS(checkgdminput, file = paste(pathtodata, "/analysis/output_datasets/gdm_input_TEST.Rds", sep = ""))
```

# correlation plot
```{r}
# read in prepared set of variables for full overview
# checkgdminput <- readRDS(paste(pathtodata, "/analysis/output_datasets/gdm_input_distances_main_models.Rds", sep = ""))

## correlation plot
# use nice names for plotting, only those which are available
c <- nicenames$nicenames
names(c) <- nicenames$names
c <- c[names(checkgdminput)]
names(checkgdminput)[!is.na(c)] <- c[!is.na(c)]
# names(checkgdminput)[is.na(c)] <- c("EFdist", "geographic distance")
names(checkgdminput) <- gsub("\\\\narthropodsoillarvae", " a. larvae", names(checkgdminput))

# name : modelname "_corrplot_before_gdm.pdf"
m <- cor(checkgdminput, method = "spearman")
cairo_pdf(paste("vignettes/out/", modelname, "_corrplot_before_gdm_main_models.pdf", sep = ""), width = 11.7, height = 8.3)
corrplot::corrplot(m, type = "lower", addCoef.col = "black", method = "color", diag = F, tl.srt = 50, tl.col = "black", order = "hclust", tl.cex = 0.5, cl.cex = 0.3, number.cex = 0.25)
dev.off()

m[m < 0.5 & m > -0.5] <- 0
corrplot::corrplot(m, type = "lower", addCoef.col = "black", method = "color", diag = F, tl.srt = 10, tl.col = "black", order = "hclust", tl.cex = 0.5, cl.cex = 0.3, number.cex = 0.3)

# output high correlations as a table
h <- as.matrix(m)
h[!lower.tri(h)] <- 1000
h <- reshape2::melt(h, value.name = "correlation")
h <- data.table(h)
h <- h[!correlation == 1000, ]
h[correlation > 0.7 | correlation < -0.7, ][order(correlation)]
```


```{r}
# ## variance inflation in linear model
# # note : p-values are not trustable, because the assumption of independent observations is  not met! (matrix correlation)
# #    DEPRECIATED --> new method needed
# # generate model
# mod <- paste("distance ~ " ,paste(colnames(checkgdminput)[!colnames(checkgdminput) %in% c("distance")], collapse = " + "))
# m1 <- lm(formula = formula(mod), data = checkgdminput)
# # save linear model
# saveRDS(m1, file = paste("vignettes/out/", modelname, "_linear_model_before_gdm.Rds", sep = ""))
# # save linear model output
# capture.output(summary(m1),
#                file = paste("vignettes/out/", modelname, "_lm_summary.txt"))
# 
# vif_output <- car::vif(m1)
# vif_output[which(vif_output > 5)]
# # vif_output[which(vif_output[, "GVIF"] > 5),]
# capture.output(c(mod,
#                vif_output[which(vif_output > 5)]),
#                file = paste("vignettes/out/", modelname, "_vif_before_gdm.txt", sep = ""))
# 
# # plot linear model
# # m1$coefficients
# test <- summary(m1)
# coefs <- test$coefficients
# coefs <- data.table(coefs)
# coefs[, names := rownames(test$coefficients)]
# coefs[`Pr(>|t|)` < 0.05, pval_symbol := "*"]
# coefs[`Pr(>|t|)` > 0.05, pval_symbol := ""]
# 
# linplot <- ggplot(coefs, aes(x = names, y = Estimate)) +
#   geom_bar(stat = "identity") +
#   coord_flip() +
#   geom_text(aes(label=pval_symbol), y=0.5, color="black", size=3.5) +
#   geom_errorbar(aes(ymin=Estimate-`Std. Error`, ymax=Estimate+`Std. Error`), width=.2, color = "red") +
#   background_grid() +
#   ggtitle(modelname) +
#   ylab("linear model estimates") +
#   xlab("") +
#   ylim(-0.4, 0.5) +
#   labs(caption = "error bars correspond to standard errors")
# 
# ggsave(linplot, filename = paste("vignettes/out/", modelname, "_linear_model_directions.pdf", sep = ""), dpi = 400,
#        units = "cm", width = 29.7, height = 21)
```





Next script is `plot_GDM.Rmd`.