---
title: "clean and load soil covariates"
author: "Noelle Schenk"
date: "June 18, 2019"
output: html_document
---
requires `analysis_nonpublic.R` to load soil covariates and raw functions for plot IDs

# Clean and load soil covariates
This scripts loads and cleans soil covariates (also selects the plots for analysis). The output is then used (by other scripts) for 
- preparation for soil covariates for analysis


# Soil
## Cleaning soil covariates
```{r}
#USER : change paths and uncomment wanted
# main_glsoil <- fread("../data/Grassland EP soil descriptors.txt")
# small_glsoil <- fread("../data/BE env covariates.txt") # from here, I only take soil type and TWI
small_glsoil <- small_glsoil[, c("Plot", "Soil.type", "TWI")]
# take out empty row at the end
small_glsoil <- small_glsoil[Plot != "",]
```

clean and merge covariate dataset
```{r}
plot.names <- names_gl[Landuse == "G",.(EP_PlotID,PlotID)]
data.table::setnames(plot.names, old="EP_PlotID", new="Plot")
# add useful plot IDs
plot.names <- merge(plot.names, raw_grlfuns[, c("Plot", "Plotn")], by="Plot", all=T)
# merge main_glsoil with plot.names to get the wanted names
data.table::setnames(main_glsoil, "plotid","PlotID")
main_glsoil <- merge(main_glsoil, plot.names, by="PlotID", all=T)
# main_glsoil[Plotn %in% c("AEG29", "AEG10"),] # there are two duplicated rows without data
main_glsoil <- main_glsoil[!PlotID %in% c("A32886", "A8338"),]

# merge the 2 datasets
data.table::setnames(small_glsoil, "Plot", "Plotn")
glsoil <- merge(main_glsoil, small_glsoil, by="Plotn", all=T)
rm(main_glsoil) ; rm(plot.names) ; rm(small_glsoil)

# remove unwanted columns
glsoil[, c("PlotID", "Exploratory", "Plot") := NULL]
```

## Plot selection
```{r, eval=T}
# take out the plots that I exclude from analysis
glsoil <- glsoil[Plotn %in% plotNAset ,]
```

## Create binary data from Soil.type
Because PCA can not fit categorical data which have > 2 categories. 

This probably adds some overlaping information with the amount of sand, clay etc. But this does not matter, as selecting the PCA components will take this extra information out again.

TODO : *This is a "human" classification, not a "pure physical". Based on which informations? Do we add some "false" information with that, which is then left in the data as only the overlapping information is "removed" in the PCA?*

```{r}
stype <- glsoil[, .(Plotn, Soil.type)]
stype[, Rendzina := ifelse(Soil.type == "Rendzina", 1, 0)]
stype[, Braunerde := ifelse(Soil.type == "Braunerde", 1, 0)]
stype[, Pelosol := ifelse(Soil.type == "Pelosol", 1, 0)]
stype[, Pseudogley := ifelse(Soil.type == "Pseudogley", 1, 0)]
stype[, Erdniedermoor := ifelse(Soil.type == "Erdniedermoor", 1, 0)]
stype[, Gley := ifelse(Soil.type == "Gley", 1, 0)]
stype[, Mulmniedermoor := ifelse(Soil.type == "Mulmniedermoor", 1, 0)]
stype[, Parabraunerde := ifelse(Soil.type == "Parabraunerde", 1, 0)]
stype[, Fahlerde := ifelse(Soil.type == "Fahlerde", 1, 0)]
stype[, Soil.type := NULL]
# merge to glsoil
glsoil <- merge(glsoil, stype, by="Plotn", all=T)
rm(stype)
```
