---
title: "prepare functions"
author: "noelle"
date: "November 14, 2018"
output: html_document
---


# Loading file paths
The paths of file locations are stored externally in "info_data.csv" where all information is stored.
Load this file externally and store in variable "info_data". More information is given in "nonpublic.R", 
at server where the whole analysis is stored.

*note* : to read in data, you need to be connected to server. 

```{r}
info_data[,c("ColumnName", "path_unix")]
# only show lines which contain path info
info_data <- info_data[grep("txt", info_data$path_unix),c("ColumnName", "path_unix")]
```


```{r}
require(data.table)
```


# Ecosystem functions dataset

sample dataset (from Eric, some Columns will be included)
```{r}
efsample <- fread(paste(pathtodata, "../assembled_data/EP grass functions & services.txt", sep=""), header=T)
```

## Plot, exploratories
TODO : find nice Plot names to work with them and not HEG1. Use this dataset to convert "bad" names.
```{r}
PlotExplo <- fread(info_data$path_unix[grep("Biomass2011", info_data$ColumnName)], header=T, sep='\t')
PlotExplo <- PlotExplo[,c("EpPlotID", "explo")]
setnames(PlotExplo, old=c("EpPlotID", "explo"), new=c("Plot", "Explo"))
```



# Biomass
years 2008 to 2017. **caution**, 2008 might not be reliable.

all biomasses in g/m^2
```{r}
Biomass2011 <- fread(info_data$path_unix[grep("Biomass2011", info_data$ColumnName)], header=T, sep='\t')
Biomass2011 <- Biomass2011[,c("EpPlotID", "biomass_g")]
setnames(Biomass2011, old = c("EpPlotID", "biomass_g"), new = c("Plot", "Biomass2011"))

Biomass2014 <- fread(info_data$path_unix[grep("Biomass2014", info_data$ColumnName)], header=T)
Biomass2014 <- Biomass2014[,c("EpPlotID", "biomass_g")]
setnames(Biomass2014, old = c("EpPlotID", "biomass_g"), new = c("Plot", "Biomass2014"))

Biomass2015 <- fread(info_data$path_unix[grep("Biomass2015", info_data$ColumnName)], header=T, quote="", drop=19)
Biomass2015 <- Biomass2015[,c("EpPlotID", "biomass_g")]
setnames(Biomass2015, old=c("EpPlotID", "biomass_g"), new=c("Plot", "Biomass2015"))
```



# Soil Organic C
```{r}
SoilOrganicC <- fread(info_data$path_unix[grep("SoilOrganicC", info_data$ColumnName)], header=T, sep='\t')
SoilOrganicC <- SoilOrganicC[which(SoilOrganicC$Type == "G"), c("EP_Plotid","Organic_C")]
```


# Root biomass
Includes fine and coarse biomass, added them together.
```{r}
Root.biomass <- fread(info_data$path_unix[grep("Root.biomass", info_data$ColumnName)], header=T, sep='\t')
Root.biomass <- Root.biomass[Type == "G"]
Root.biomass[, "Root.biomass" := rowSums(.SD,na.rm=T), .SDcols=c("Fine_Roots_Biomass", "Coarse_Roots_Biomass")]
Root.biomass <- setnames(Root.biomass[,c("EP_Plotid", "Root.biomass")], old=c("EP_Plotid", "Root.biomass"), new=c("Plot", "Root.biomass"))
```

# Flower cover
We can go from "total_influorescence_area_in_cm" of each species to a total flower cover. The total area measured was : 600m² per grassland plot.
TODO : left open as not sure if I include it
```{r}
Total.flower.cover <- fread(info_data$path_unix[grep("Total.flower.cover", info_data$ColumnName)], header=T, sep='\t')
```


# Microbial phosphorous in soil
(mg/kg)
TODO : why is it a function/ important?
```{r}
Pmic.2011 <- fread(info_data$path_unix[grep("Pmic.2011", info_data$ColumnName)], header=T, sep='\t')
```


# Root decomposition
Decomposition measurements of fine roots (<2mm) within the upper 10 cm of the mineral soil. Determination of mass loss after 6 months. Litterbag method. Standardized control material: beech roots for forest sites and herbaceous roots for grassland sites. Numbers in [%]

we use Mass_loss_Octover_2012 TODO : why?
```{r}
Root.decomposition <- fread(info_data$path_unix[grep("Root.decomposition", info_data$ColumnName)], header=T, sep='\t')
Root.decomposition <- Root.decomposition[Type == "G"]
Root.decomposition <- setnames(Root.decomposition[,c("EP_Plotid", "Mass_loss_October_2012")], old=c("EP_Plotid", "Mass_loss_October_2012"), new=c("Plot", "Root.decomposition"))
```


# Bulk density and soil C stock
**Bulk density** : Increased use of machine compact the soil. At a given point, the root growth is impaired.

**soil C stock** : C in soil

Determination bulk density and calculated CN stocks for the upper 10 cm of the mineral soil. Dry combustion of soil at a temperature of 1100°C and subsequent determination of evolving CO2 and N2 with a Thermal Conductivity Detector (TCD).

[g/cm^3]

**data is from 2011, not 2014!**

```{r}
Bulk.density.soil.C.stock.2011 <- fread(info_data$path_unix[grep("Soil.C.stock", info_data$ColumnName)], header=T, sep='\t')
Bulk.density.soil.C.stock.2011 <- Bulk.density.soil.C.stock.2011[Type == "G"]
Bulk.density.soil.C.stock.2011 <-setnames(Bulk.density.soil.C.stock.2011[,c("EP_Plotid", "BD", "OC_stock")], old=c("EP_Plotid", "BD", "OC_stock"), new=c("Plot", "Bulk.density.2014", "Soil.C.stock"))
```


# Herbivory

## 2013
invertebrate herbivory in 146 managed temperate grasslands. Year: 2013

increased land-use - less invertebrates - less herbivory

Visually estimation of herbivory by eye using templates and measurement of leaf area left after feeding of the herbivores using LI-COR area meter

2 samples per plot. The mean of both measures is taken.
```{r}
herb <- fread(info_data$path_unix[grep("herbivory.2013", info_data$ColumnName)], header=T, sep='\t')
setorder(herb, EP_Plotid)
# prepare empty dt and names vector for forloop
ns <- unique(herb[,EP_Plotid])
herbivory = data.table(Plot = ns, herbivory = NA)
# fill in herbivory values in forloop
for(i in 1:length(ns)){
  herbivory$herbivory[i] <- mean(herb[EP_Plotid == herbivory$Plot[i], PercHerbivory], na.rm = T)
}
rm(herb)
```


# Forage quality
"Futter qualität", year : 2009
forage quality (index based on crude protein concentration and relative forage value) - Eric Ecology Letters LUI

RFV : Relative Feed Value
RFQ : Relative Forage Quality
ADL : acid detergent lignin, Lignin
ADF : Acid Detergent Fibre, Säure-Detergenz-Faser, ein Teil der Gerüstsubstanzen von Futtermitteln
NDF : Neutral Detergent Fibre, Neutral-Detergenz-Faser, fasrige Anteile von Futtermitteln 

more ADF/ NDF --> less nutritional value because less easy to digest

according to forage quality estimation based on alfalfa (medicago sativa) values. [source](https://openprairie.sdstate.edu/cgi/viewcontent.cgi?referer=https://scholar.google.ch/&httpsredir=1&article=1351&context=extension_extra)

[% concentration in aboveground grassland plant community biomass without litter]
```{r}
forage.quality <- fread(info_data$path_unix[grep("forage.quality", info_data$ColumnName)], header=T, sep='\t')
forage.quality <- forage.quality[year == "2009"]
# calculating the columns needed for forage quality calc
forage.quality[, Crude.protein := 6.25*N]
forage.quality[, DDM := 88.9-(0.779* ADF)]
forage.quality[, DMI := 120/NDF]
forage.quality[, RFV := (DDM*DMI)/1.29]

forage.quality[, forage.quality := apply(scale(forage.quality[,c("RFV", "Crude.protein")]), 1, mean)]
forage.quality <- setnames(forage.quality[,c("Plot_name", "forage.quality")], old=c("Plot_name", "forage.quality"), new= c("Plot", "forage.quality"))
```
TODO : **different values than in efsample!!** Really calculated with 2009? Or is it the mean among all years (2009 - 2013). If yes - take mean among years AFTER forage.quality calculation?! Note : the column names given in assembled dataset do not correspond to columns in "my" dataset - e.g. ShootN.conc in Erics dataset is "N" in mine.



# Phosphorous (loss) 2011
Resin refers to the method of phosphorous measurement in soil.
```{r}
P.loss2011 <- fread(info_data$path_unix[grep("P.loss2011", info_data$ColumnName)], header=T, sep='\t')
```

TODO : how to quantify LOSS? I have the phosphorous measured...


## Phosphorous 2014
Olsen method, Pi is inorganic Phosphorous
```{r}
OlsenPi2014 <- fread(info_data$path_unix[grep("OlsenPi2014", info_data$ColumnName)], header=T, sep='\t')
# grassland plots are abbreviated with a "G", e.g. "AEG", forest plots with "W", e.g. "AEW"
OlsenPi2014 <- OlsenPi2014[grep("G",OlsenPi2014$Plot_ID),c("Plot_ID","OlsenPi")]
```








