---
title: "detect GDM effect shape"
author: "Noelle Schenk"
date: "1/15/2020"
output: html_document
---
this script detects the shape of gdm model lines
required dataset : 

```{r}
require(MASS)
require(pgirmess)
require(nlme)
require(nls2)
```

```{r}
## define model names either automatically or manually
model_name <- model_names$name[1]
# model_name <- "gdm_EFdistance_LUI"

gdm_output <- readRDS(paste_gdm_input_path_together(pathtoout = pathtoout, name = model_name))
exSplines <- gdm::isplineExtract(gdm_output)
```
prepare data (copied from `plot_GDM.Rmd`)
```{r}
# extract I splines y axes and get LUI x axis
lineplot_data <- data.table(cbind(exSplines$x[, "Geographic"], exSplines$y))
setnames(lineplot_data, old = c("V1", "Geographic"), new = c("xaxis", "LUI"))
lineplot_data <- melt(lineplot_data, measure.vars = colnames(lineplot_data)[-1], id.vars = "xaxis", 
                      variable.name = "names")
# adding plot information from nicenames
lineplot_data <- merge(lineplot_data, nicenames, by = "names", all.x = T)
lineplot_data <- lineplot_data[, .(names, xaxis, value)]
```

# simulations
## power law
```{r}
cols <- RColorBrewer::brewer.pal(9, "Set1")
# power law
par(mfrow = c(1, 2))
x <- seq(0, 3, 0.001)
plot(x, 0.1 * x ^ 0.1, type = "l", ylim = c(0, 0.12), main = "y ~ b * x ^ z , increasing z with constant b")
lines(x, 0.1 * x ^ 0.3, col = cols[1])
lines(x, 0.1 * x ^ 0.5, col = cols[2])
lines(x, 0.1 * x ^ 0.7, col = cols[3])
lines(x, 0.1 * x ^ 0.9, col = cols[4])
lines(x, 0.1 * x ^ 1, col = cols[5])
lines(x, 0.1 * x ^ 1.5, col = cols[6])
lines(x, 0.1 * x ^ 2, col = cols[7])
lines(x, 0.1 * x ^ 3, col = cols[8])
lines(x, 0.1 * x ^ 4, col = cols[9])

plot(x, 1 * x ^ 0.1, type = "l", ylim = c(0, 1.1), main = "y ~ b * x ^ z , decreasing b with 2 constant z (<1 and >1)")
lines(x, 0.9 * x ^ 0.1, col = cols[1])
lines(x, 0.5 * x ^ 0.1, col = cols[2])
lines(x, 0.1 * x ^ 0.1, col = cols[3])
lines(x, 0.3 * x ^ 0.1, col = cols[4])
lines(x, 0.7 * x ^ 0.1, col = cols[5])
lines(x, 2 * x ^ 3, col = cols[6])
lines(x, 3 * x ^ 3, col = cols[7])
lines(x, 4 * x ^ 3, col = cols[8])
lines(x, 5 * x ^ 3, col = cols[9])
par(mfrow = c(1, 1))
```
# exponential
find appropriate starting values for exponential model
```{r}
par(mfrow = c(1, 1))
x <- seq(0, 3, 0.001)

plot(x, 1*exp(-1*x) + 0, type = "l", main = "y ~ k*exp(-b1*xaxis) + b0", ylim = c(0, 1))
lines(x, 2*exp(-1*x) + 0, col = cols[1])
lines(x, 0.5*exp(-1*x) + 0, col = cols[2])
lines(x, 2*exp(-1*x) + 0, col = cols[3])

lines(x, 1*exp(0.5*x) -1, col = cols[4])
lines(x, 1*exp(0.1*x) -1, col = cols[5])

lines(x, 1*exp(-1*x) -0.1, col = cols[6])
```


```{r}
par(mfrow = c(2, 2))
test <-lineplot_data[names == "autotroph.beta.sne",]
# test with other names as well (e.g. LUI, geo.dist, ...)

nlc <- nls.control(maxiter = 10000)

# fit a linear model
lin_mod <- lm(value ~ xaxis, data = test)
AIC(lin_mod)
par(mfrow = c(2, 2))
# plot(lin_mod)
plot(test$xaxis, test$value, main = "modelled values and fitted with power law model", sub = round(AIC(lin_mod), 2))
abline(lin_mod, col = cols[1], lwd = 3)

# fit an asymptotic model
asym_mod <- gnls(value ~ SSlogis(xaxis, Asym, xmid, scal), data = test,
            weights = varPower())
plot(test$xaxis, test$value, main = "modelled values and fitted with asymptotic model", sub = round(AIC(asym_mod), 2))
yy <- coef(asym_mod)[1] / (1 + exp((coef(asym_mod)[2] - test$xaxis)/coef(asym_mod)[3] )) # == SSlogis(xx, *):
lines(test$xaxis, yy, col = cols[1], lwd = 3)

# fit exponential model
exp_mod <- tryCatch(nls(value ~ k*exp(-b1 * xaxis) + b0, start = list(k = 1, b0 = -0.5, b1 = -1), data = test), silent = T) # growth
exp_mod <- tryCatch(nls(value ~ k*exp(-b1 * xaxis) + b0, start = list(k = 1, b0 = 1, b1 = 0), data = test), silent = T) # decay
plot(test$xaxis, test$value, main = "modelled values and fitted with exponential model", sub = round(AIC(exp_mod), 2))
yy <- coef(exp_mod)[1] * (exp(-(coef(exp_mod)[3]) * test$xaxis)) + coef(exp_mod)[2]
lines(test$xaxis, yy, col = cols[1], lwd = 3)

# power law
pow_mod <- tryCatch(nls(value ~ b*xaxis^z, start = list(b = 1, z = 0.1), data = test))
summary_pow_mod <- summary(pow_mod)
plot(test$xaxis, test$value, main = "modelled values and fitted with power law model", , sub = round(AIC(pow_mod), 2))
lines(test$xaxis, summary_pow_mod$coefficients[1,1] * test$xaxis ^ summary_pow_mod$coefficients[2, 1], col = cols[1], lwd = 3)
```


- linear : lm() , y ~ a + b
- asymptotic : model <- drm(Y ~ X, fct = DRC.asymReg())
gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym + lrc ~ LUI.sd, R0 ~ Exploratory), start = c(p[1],0,p[3],0,p[3],1,1), DF,control=nlc)
- exponential : 
try(gnls(response ~ a*exp(-b*LUI), params=list(a ~ Exploratory + LUI.sd, b ~ LUI.sd), start=c(1,1,1,1,1,1),DF)),
- power law
```{r}
set.seed(1234)
X <- c(1, 3, 5, 7, 9, 11, 13, 20)
a <- 20; b <- 5; c <- 0.3
Ye <- asymReg.fun(X, a, b, c)
epsilon <- rnorm(8, 0, 0.5)
Y <- Ye + epsilon
model <- drm(Y ~ X, fct = DRC.asymReg())
plot(model, log="", main = "Asymptotic regression")
```



```{r}
evaluate.all <- function(response, DF, p = NULL, reduced = FALSE, models = "All"){

DF <- cbind(response = DF[[response]], DF)

DF <- DF[complete.cases(DF$response),]

nlc <- nls.control(maxiter = 10000)

if(length(p)==0){

p <- try(getInitial(response ~ SSasymp(LUI, Asym, R0, lrc), data = DF))

if(inherits(p, "try-error")){

print("Please supply starting parameters for Asymptotic exponential")
}}

spars<-cbind(seq(0,1,length.out=1000), seq(0,2,length.out = 1000), seq(-3,1, length.out = 1000))
spars <- data.frame(apply(spars,2,sample))
names(spars)<-letters[1:3]
invisible(capture.output(m <- try(nls2(response ~ a+b*(LUI^c), start=spars, algorithm="brute-force",DF))))
p2 <- coef(m)

L=list(
#### No LUI.sd
try(lm(response ~  Exploratory, DF)),
try(lm(response ~  Exploratory + Fstd, DF)),
try(lm(response ~  Exploratory + Mstd, DF)),
try(lm(response ~  Exploratory + Gstd, DF)),
try(lm(response ~  Exploratory + Fstd + Gstd, DF)),
try(lm(response ~  Exploratory + Fstd + Mstd, DF)),
try(lm(response ~  Exploratory + Mstd + Gstd, DF)),
try(lm(response ~  Exploratory + Fstd + Mstd + Gstd, DF)),

try(lm(response ~  Exploratory + poly(Fstd,2), DF)),
try(lm(response ~  Exploratory + poly(Mstd,2), DF)),
try(lm(response ~  Exploratory + poly(Gstd,2), DF)),

try(lm(response ~  Exploratory + poly(Fstd,2) + Mstd, DF)),
try(lm(response ~  Exploratory + poly(Fstd,2) + Mstd + Gstd, DF)),
try(lm(response ~  Exploratory + poly(Fstd,2) + Gstd, DF)),

try(lm(response ~  Exploratory + poly(Mstd,2) + Fstd, DF)),
try(lm(response ~  Exploratory + poly(Mstd,2) + Fstd + Gstd, DF)),
try(lm(response ~  Exploratory + poly(Mstd,2) + Gstd, DF)),

try(lm(response ~  Exploratory + poly(Gstd,2) + Fstd, DF)),
try(lm(response ~  Exploratory + poly(Gstd,2) + Fstd + Mstd, DF)),
try(lm(response ~  Exploratory + poly(Gstd,2) + Mstd, DF)),

try(lm(response ~  Exploratory + poly(Fstd,2) + poly(Mstd,2), DF)),
try(lm(response ~  Exploratory + poly(Fstd,2) + poly(Gstd,2), DF)),
try(lm(response ~  Exploratory + poly(Mstd,2) + poly(Gstd,2), DF)),

try(lm(response ~  Exploratory + poly(Fstd,2) + poly(Mstd,2) + Gstd, DF)),
try(lm(response ~  Exploratory + poly(Fstd,2) + poly(Gstd,2) + Mstd, DF)),
try(lm(response ~  Exploratory + poly(Mstd,2) + poly(Gstd,2) + Fstd, DF)),

try(lm(response ~  Exploratory + poly(Fstd,2) + poly(Mstd,2) + poly(Gstd,2), DF)),

try(lm(response ~  Exploratory + LUI, DF)),
try(lm(response ~  Exploratory + poly(LUI,2), DF)),
try(lm(response ~  Exploratory + poly(LUI,3), DF)),
try(gnls(response ~ a*exp(-b*LUI), params = list(a~Exploratory, b~1), start = c(1,1,1,1), DF)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym + lrc ~ 1, R0 ~ Exploratory), start = c(p[c(1,3)],p[2],1,1), DF,control=nlc)),
try(gnls(response ~ a+b*(LUI^c), params=list(a ~Exploratory, b + c ~ 1), start=c(p2[1],0,0,p2[2],p2[3]),DF)),

### LUI.sd main effect
try(lm(response ~ Exploratory + LUI.sd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Fstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Mstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Gstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Fstd + Gstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Fstd + Mstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Mstd + Gstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + Fstd + Mstd + Gstd, DF)),

try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2), DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Mstd,2), DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Gstd,2), DF)),

try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + Mstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + Mstd + Gstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + Gstd, DF)),

try(lm(response ~  Exploratory + LUI.sd + poly(Mstd,2) + Fstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Mstd,2) + Fstd + Gstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Mstd,2) + Gstd, DF)),

try(lm(response ~  Exploratory + LUI.sd + poly(Gstd,2) + Fstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Gstd,2) + Fstd + Mstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Gstd,2) + Mstd, DF)),

try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + poly(Mstd,2), DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + poly(Gstd,2), DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Mstd,2) + poly(Gstd,2), DF)),

try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + poly(Mstd,2) + Gstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Fstd,2) + poly(Gstd,2) + Mstd, DF)),
try(lm(response ~  Exploratory + LUI.sd + poly(Mstd,2) + poly(Gstd,2) + Fstd, DF)),

try(lm(response ~  Exploratory + LUI.sd +  poly(Fstd,2) + poly(Mstd,2) + poly(Gstd,2), DF)),

try(lm(response ~ Exploratory + LUI + LUI.sd, DF)),
try(lm(response ~ Exploratory + poly(LUI,2) + LUI.sd, DF)),
try(lm(response ~ Exploratory + poly(LUI,3) + LUI.sd, DF)),
try(gnls(response ~ a*exp(-b*LUI), params=list(a ~ Exploratory + LUI.sd, b ~ 1), start=c(1,1,1,1,1),DF)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym + lrc ~ 1, R0 ~ Exploratory + LUI.sd), start = c(p[c(1,3)],p[2],1,1,0), DF,control=nlc)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym ~ LUI.sd,  lrc ~ 1, R0 ~ Exploratory), start = c(p[1],0,p[3],p[2],1,1), DF,control=nlc)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym ~ 1,  lrc ~ LUI.sd, R0 ~ Exploratory), start = c(p[c(1,3)],0,p[2],1,1), DF,control=nlc)),
try(gnls(response ~ a+b*(LUI^c), params=list(a ~Exploratory + LUI.sd, b + c ~ 1), start=c(p2[1],0,0,0,p2[2],p2[3]),DF)),

### LUI.sd interaction
try(lm(response ~ Exploratory + LUI * LUI.sd, DF)),
try(lm(response ~ Exploratory + poly(LUI,2) * LUI.sd, DF)),
try(lm(response ~ Exploratory + poly(LUI,3) * LUI.sd, DF)),
try(gnls(response ~ a*exp(-b*LUI), params=list(a ~ Exploratory + LUI.sd, b ~ LUI.sd), start=c(1,1,1,1,1,1),DF)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym + lrc ~ LUI.sd, R0 ~ Exploratory), start = c(p[1],0,p[3],0,p[3],1,1), DF,control=nlc)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym~ LUI.sd, lrc ~ 1, R0 ~ Exploratory + LUI.sd), start = c(p[1],0,p[3],p[3],1,1,0), DF,control=nlc)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym~ 1, lrc ~ LUI.sd, R0 ~ Exploratory + LUI.sd), start = c(p[1],p[3],0,p[3],1,1,0), DF,control=nlc)),
try(gnls(response ~ Asym+(R0-Asym)*exp(-exp(lrc)*LUI), params=list(Asym + lrc ~ LUI.sd, R0 ~ Exploratory + LUI.sd), start = c(p[1],0,p[3],0,p[3],1,1,0), DF,control=nlc)),
try(gnls(response ~ a+b*(LUI^c), params=list(a ~Exploratory+ LUI.sd, b + c ~ LUI.sd), start=c(p2[1],0,0,0,p2[2],0,p2[3],0),DF))
)

names(L) <- paste("model", 1:length(L), sep ="")

### select reduced set without lui.sd interactions
if (reduced == TRUE){
L <- L[1:68]
}
else{}

#### select set with only LUI or without LUI

cfs <- lapply(L,"[",1)
cfs <- lapply(cfs, function(x)(names(unlist(x))))

if(models == "LUI.only"){
L <- L[-grep("std",cfs)]
}
if(models == "no.LUI"){
L <- L[grep("std",cfs)]
}
else{}

#to select only those models that converged without error

L2 <- Filter(function(x) !inherits(x, "try-error"), L)

## pseudo r2

fit <- lapply(L2, fitted)

pr2 <- lapply(fit, function(x)(cor(x, DF$response, method = "pearson", use = "complete.obs")^2))

pr2 <- unlist(pr2)

# actual model selection based on AICc:

df <- data.frame(selMod(L2))

nn <- as.numeric(row.names(df))

pr2 <- pr2[nn]

df <- cbind(df, "pseudo.r2" = pr2)

# return the models accounting for 95% of AIC weights

if(sum(df[,"w_ic"])<0.95){

called <- L2[nn]
print("all models in 95% set")
}
else{
selected <- nn[1:which(cumsum(df[,"w_ic"])>=0.95)[1]]

# return the model formula (this has the structure response~...,data=DF)
called <- L2[selected]
}

### rename rownames of df so that the numbers correspond to model positions in the full list not in the reduced list (L2)

names(L2) <- which(lapply(L, inherits,"try-error")==FALSE)
rownames(df) <- names(L2[nn])
 
# Finally, return the selected models

returnlist <- list(response=response, models=df, selected.model=called, all.models=L2)

return(returnlist)
}
```

