---
title: "Functions imputation"
output:
  pdf_document: default
  html_document: default
---

This script : 
- uses raw functions from the synthesis dataset (loaded with `analysis_nonpublic.R`)
- performs gapfilling
- saves the gapfilled dataset

# Data
The dataset used for gap filling is larger than the dataset used for the analysis. More variables enhance the performance of gapfilling.

Note : Only use variables which are not used in the analysis. Otherwise, data can be circular.
```{r, eval=F}
#USER : insert path to grassland function file and read in as grlfuns (uncomment line below)
# raw_grlfuns <- data.table::fread("<pathto_grl_functions.csv>", header=T)
source("analysis_nonpublic.R")
```
Grassland functions data used for gapfilling is stored in `raw_grlfuns`, selected variables according to info data. `raw_grlfuns` is loaded by `analysis_nonpublic.R`.

Requires the plot set defined by diversity dataset with the final set of included plots for analysis. *Note* that this script removes an additional plot from the plot set. However, this does not matter for the performance of this imputation script.


# Gap filling
A Gap-filling of the functions dataset is performed, because many values are missing (NA).

Two variables are produced, `impdat` which contains the gapfilled dataset which will be used in further analysis as well as `imp_err`, which contains the imputation error automatically estimated by `missForest`. The imputed dataset is stored as .rds file.

## Data preparation

```{r}
befun <- data.table::copy(raw_grlfuns)
```

chosing the plot set
```{r}
befun <- befun[Plotn %in% plotNAset]
```

cleaning out variables with too many missing values
```{r}
visdat::vis_miss(befun, sort_miss = T)
```
Values with over 21% missing data could be excluded here (mAMF hyphae, Parasitoid traps, Aggregation, PRI.2011 and NaHCO3.Pi).

Here, values with too many missings are not taken out yet. Instead, the imputation is done and the variables are only removed in case their imputation error is too high.
```{r, eval=F}
#USER : set treshold for missing values
treshold <- 0.21
t <- apply(befun, 2, function(x) sum(is.na(x)))
exclude <- names(which(t > 150 * treshold))
befun <- befun[, !colnames(befun) %in% exclude, with=F]
rm(t); rm(exclude); rm(treshold)
```

If the missing plot from 16S, amoA_AOA.2016 & Co is taken out, 3 more functions can be included in the functions set.
```{r}
# befun[is.na(`16S_NB`), Plotn]
# befun[is.na(amoA_AOA.2016), Plotn]
# befun[is.na(amoA_AOB.2016), Plotn]
# The missing plot is "HEG31"
befun <- befun[Plotn != "HEG31"]
```
update the helper vector `plotNAset.rds` and take out the plot "HEG31"
```{r, eval=F}
# readRDS("plotNAset.rds")
plotNAset <- plotNAset[-which(plotNAset == "HEG31")]
saveRDS(plotNAset, "plotNAset.rds")
```

## Check Correlations
Of the non-factor columns only
```{r}
corrbefun <- befun[, !colnames(befun) %in% c("Explo", "Soil.type", "Plotn", "Plot"), with=F]
M <- cor(corrbefun, use="pairwise.complete.obs")

# raw_grlfun_corrplot1
corrplot::corrplot(M,type="lower",addCoef.col = "black",method="color",diag=F, tl.srt=1, tl.col="black", mar=c(0,0,0,0), number.cex=0.25, tl.cex = 0.3)

# raw_grlfun_corrplot2
corrplot::corrplot(M, type = "upper", tl.col="black", tl.srt=40, diag = F, tl.cex = 0.3)

# only show high correlations
tres <- 0.7
M[which(M < tres & M > -tres)] <- 0
corrplot::corrplot(M, type = "upper", tl.col="black", tl.srt=40, diag = F, tl.cex = 0.3)
```
Network
```{r}
# show network - is there a variable "alone"? bzw. isolated?
network <- igraph::graph_from_adjacency_matrix(M, weighted=T, mode="undirected", diag=F)
plot(network)
```
There is no variable which is not connected with at least one line to another variable. Lines represent correlations > threshold.
```{r}
rm(M); rm(network)
```


## Log transformation
before changing values, store the original dataset
```{r}
befunbackup <- data.table::copy(befun)
```
remove negative values, remember to add again after imputation!

find negative values
```{r}
par(mar = c(10, 1, 3, 1))
boxplot(corrbefun, las=2)
# minimum of the numeric columns
negative <- colnames(corrbefun)[which(corrbefun[, lapply(.SD, function(x) min(x, na.rm=T))] < 0)]
```
remove negative values - shift whole vector up.
```{r}
# DEA.inverted
mindea <- min(befun$DEA.inverted, na.rm=T)
befun$DEA.inverted <- befun$DEA.inverted + abs(mindea)
# NO3.2014
minno <- min(befun$NO3.2014, na.rm=T)
befun$NO3.2014 <- befun$NO3.2014 + abs(minno)
# dung.removal
mindung <- min(befun$dung.removal, na.rm=T)
befun$dung.removal <- befun$dung.removal + abs(mindung)

rm(negative)
```
Log transformation of the numeric columns. The natural logartithm of values smaller than 1 is negative. Tehrefore, all values are shifted by 1 again to avoid any negative values in the log transformed dataset.
```{r}
numcols <- colnames(befun)[!colnames(befun) %in% c("Explo", "Soil.type", "Plotn", "Plot")]
befun[, (numcols) := lapply(.SD, as.numeric), .SDcols = numcols]
# log transform the values + 1, later shift them back.
befun[, (numcols) := lapply(.SD, function(x) log(x+1)), .SDcols = numcols]
# record NA value positions for visualistation
naloc <- is.na(befun)
```

## Imputation

The imputation is repeated 50 times, and the mean of the imputed values is taken. 

```{r}
# 50 imputations
impvals <- list()
imperr <- list()
# based on one-time imputation, take out : "amoA_AOA.2016" ,"amoA_AOB.2016",  "nxrA_NS", "16S_NB"
numcols <- numcols[which(!numcols %in% c("Plotn", "Plot"))]
dataset <- befun[, !colnames(befun) %in% c("Plotn", "Plot", "Explo"), with=F]
```

```{r imputation, results="hide", eval=F}
for(i in 1:50){
  print(i)
  current <- missForest::missForest(dataset, variablewise = T)
  imperr[[i]] <- data.table::data.table("columns" = colnames(dataset), "errors" =current$OOBerror)
  current <- data.table::as.data.table(current$ximp)
  # re-transform the numeric variables
  current[, (numcols) := lapply(.SD, function(x) (exp(x)-1)), .SDcols = numcols]
  # re-transform the negative values
  current$DEA.inverted <- current$DEA.inverted - abs(mindea)
  current$NO3.2014 <- current$NO3.2014 - abs(minno)
  current$dung.removal <- current$dung.removal - abs(mindung)
  # add back the "Plotn" column which was taken out for imputation
  current[, Plotn := befun$Plotn]
  # convert imputed data.table to matrix, as more handy for imputed values handling
  impvals[[i]] <- current
}
rm(current); rm(i); rm(mindea); rm(minno); rm(mindung)
# USER : may change filename
saveRDS(impvals, file="raw_imputed_function_values_complete.rds")
saveRDS(imperr, file = "raw_imputed_function_errors_complete.rds")
```

```{r}
impvals <- readRDS("raw_imputed_function_values_complete.rds")
imperr <- readRDS("raw_imputed_function_errors_complete.rds")
```


Check imputation error
```{r}
imperr <- do.call(rbind, imperr)
mimperr <- data.table::as.data.table(aggregate(errors ~ columns, imperr, mean))
mean(mimperr[,errors]) # 0.071

# barplot(height=mimperr[,errors], names.arg = mimperr[, columns])
library(cowplot)
p<-ggplot2::ggplot(data=mimperr, ggplot2::aes(x=columns, y=errors)) +
  ggplot2::geom_bar(stat="identity") +
  theme(axis.text.x=element_text(color = "black", size=6, angle=30, vjust=.8, hjust=0.8)) +
  geom_hline(yintercept=0.25, linetype="dashed", color = "gray")
cowplot::plot_grid(naniar::vis_miss(dataset, sort_miss = T), p, nrow = 2)
```
The two variables Total pollinators and Parasitoid traps are removed, because the imputation error is too high and because they rely on diversity data (which is hard / not possible to impute).
The other variables with imputation errors over 25% are left, because only 1-2 values are missing in those datasets.
```{r}
dataset[, c("Total.pollinators", "Parasitoid.traps") := NULL]
numcols <- numcols[-which(numcols %in% c("Total.pollinators", "Parasitoid.traps"))]
```

```{r, eval=F}
pdf("imputation_errors.pdf", paper = "a4")
cowplot::plot_grid(naniar::vis_miss(dataset, sort_miss = T), p, nrow = 2)
dev.off()
```


Get together imputed dataset from 50 imputations.
```{r}
# backup <- impvals
X <- do.call(rbind, impvals)
X <- data.table::as.data.table(X)
X[, (numcols) := lapply(.SD, as.numeric), .SDcols=numcols]
# backup2 <- data.table::copy(X)

# two ways of aggregating the 50 imputed values together
# Y <- aggregate(X[, ..numcols], list(Plotn=X$Plotn), mean)
Y <- X[, lapply(.SD, mean, na.rm=T), .SDcols=numcols, by=Plotn]
Y <- data.table::data.table(Y)
```

Compare imputed values with real values. Creates the pdf `imputed_values.pdf`
```{r, eval=F}
pdf("imputed_values.pdf", paper="a4r")
par(mfrow = c(2,7))
for(i in 1:length(numcols)){
  x <- rep(1, length(plotNAset))
  colname <- numcols[i]
  # length of the imputed value vector
  rowlocations <- naloc[, colname]
  y <- Y[rowlocations, get(colname)]
  plot(x, befunbackup[, get(colname)], xlab="", ylab="range of values", main=colname, sub=paste(length(y), "NAs"))
  x <- x[0:length(y)]
  points(x, y, col="red",pch=19)
}
dev.off()
```

Store imputed grassland function dataset.
```{r}
saveRDS(Y, file="imputed_function_values.rds")
# clean a bit
imputed_grlfuns <- data.table::copy(Y)
rm(befun); rm(befunbackup); rm(corrbefun); rm(dataset); rm(imperr); rm(impvals)
rm(X); rm(Y); rm(p); rm(naloc); rm(mimperr)
rm(numcols); rm(tres)
```

The multifunctionalities of imputed functions are calculated in another script. Current name `calc_multifun_from_imputed.Rmd`.
