<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{ISMB Twitter Analysis}
-->

*note:*

This is where the analysis is performed and described.
All necessary functions will be generated in separate R files.
As described [here](http://rmflight.github.io/post/vignette-analysis/).


# Reproducibility
This vignette and package were built as recommended by the blogpost of [Hilary Parker](https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/), [the book "R packages"](http://r-pkgs.had.co.nz/) and the idea of using `R packages` for an analysis is coming from Robert M. Flight's blogposts, where he explains [why](http://rmflight.github.io/post/analyses-as-packages/) and [how](http://rmflight.github.io/post/vignette-analysis/) to use `R package` for an analysis.

## packrat
In order to be clear on R package versions and dependencies, [packrat](http://rstudio.github.io/packrat/) was used. 
```{r}
# devtools::install_github("rstudio/packrat")
require(packrat)
```
Follow the [walktrough](http://rstudio.github.io/packrat/walkthrough.html) from packrat, which they highly recommend.


# Dataset
## obtain dataset
With account on [BExIS](https://www.bexis.uni-jena.de). Name and content of datasets : 

¦ name             ¦ content                   ¦
¦ ---------------- ¦ ------------------------- ¦
¦ bla              ¦ bla                       ¦


## clean datasets
Extract necessary data from datasets.
```{r}
trophic_groups <- c()
years <- c()
# cleanDataset1(trophic_groups, years)
```

test
