---
title: "GDM Threshold considerations"
author: 'Noëlle Schenk'
date: "5/27/2021"
output: html_document
---

requirements : 
- TODO user : connect to planteco and nsch to get input files
- GDM output as .rds, e.g. gdm_EFnestedness_LUI_input.Rds
    - find at planteco/.../BetaDivMultifun/analysis/output_datasets
- cluster output as .rds, e.g. gdm_EFnestedness_LUI_permutation.Rds
- run `results_nonpublic.R` : load pathtodata, nicenames and model_names
  - pathtodata : path to location where gdm input of model is stored. points to a folder, where the folder cluster witht the model input is. *note : was pathtoout before.*
  - nicenames : helper file to get nice names of functions and trophic levels
  - model_names : helper file for automation. Contains 3 columns : `name`  contains the all model names for `model_name`, `luiinput` contains either "lui" or "components" as LUI input for `luiinput` and `permut` is either TRUE or FALSE.
- `analysis_nonpublic.R` loads `EFmaster_all_thresholds.Rds`


# create sensitivity comparison: deviance explained, sdev
```{r, eval = F}
#############################
#
# treshold selection
#
#############################

# models with function calculated with threshold : 
# thresholds between 0.1 and 0.9.
# need selection : by (new) : sdev of input. old : percent deviance explained.

# VARIABILITY OF RESPONSE VARIABLE
# require : input data : EFmaster_all_thresholds.Rds : load from script "analysis_nonpublic.R"
EFmaster_sd <- apply(EFmaster[, -c("Var1", "Var2")], 2, function(x) sd(x, na.rm = F))
EFmaster_var <- apply(EFmaster[, -c("Var1", "Var2")], 2, function(x) var(x, na.rm = F))
EFmaster <- data.table(name = names(EFmaster), sd = EFmaster_sd, var = EFmaster_var)
EFmaster[, type := "EFdistance"]
EFmaster[grep("nestedness", EFmaster$name), type := "nestedness"]
EFmaster[grep("turnover", EFmaster$name), type := "turnover"]
EFmaster[grep("beta", EFmaster$name), type := "beta"]

ggplot(EFmaster, aes(x = name, y = sd, fill = type)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


# PERCENT DEVIANCE EXPLAINED
# require : 
# model_results : the model results saved manually from the models
setorder(model_results, cols = "name")
model_results[, EF := factor(EF, ordered = T, levels = c("distance", "singleEF", "abund", "beta", "turnover", "nestedness"))]
str(model_results)
model_results <- model_results[!is.na(dev_expl_by_GDM)]

devexpl <- ggplot(model_results, aes(x = name, y = dev_expl_by_GDM, fill = EF)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(filename = paste(pathtoout, "/plot_GDM_EFbeta_tresholds_var_expl.pdf", sep = ""), 
       plot = devexpl, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)

#TODO can add nice names at some point
# STANDARD DEVIATION AND PERCENT DEVIANCE EXPLAINED
model_results[, name := gsub("_LUI", "", model_results$name)]
model_results[, name := gsub("gdm_", "", model_results$name)]
model_results <- merge(model_results, EFmaster, by = "name")

sdplot <- ggplot(model_results, aes(x = name, y = sd, fill = EF)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
threshold_selection <- plot_grid(sdplot, devexpl, nrow = 2)

ggsave(filename = paste(pathtoout, "/plot_GDM_EFbeta_tresholds_var_expl_and_sd.pdf", sep = ""), 
       plot = threshold_selection, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```



# thresholds EF heatmap

```{r, eval = F}
# dataset is stored, next chunk reads it in
mod_names <- model_results[EF %in% c("distance", "nestedness", "turnover") & luiinput == "lui", name]
mod_names <- mod_names[-grep("0.75", mod_names)]

i <- mod_names[1]
a <- paste(pathtodata, "/analysis/output_datasets/", i, "_output.Rds", sep = "")
gdm_output <- readRDS(a)
# get maxsplines
exSplines <- gdm::isplineExtract(gdm_output)
# take maximum of each variable for barplot
maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
restab <- data.table(model = i, data.frame(as.list(maxsplines)))

for(i in mod_names[-1]){
  a <- paste(pathtodata, "/analysis/output_datasets/", i, "_output.Rds", sep = "")
  gdm_output <- readRDS(a)
  # get maxsplines
  exSplines <- gdm::isplineExtract(gdm_output)
  # take maximum of each variable for barplot
  maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
  restab_newline <- data.table(model = i, data.frame(as.list(maxsplines)))
  # append to restab
  restab <- rbindlist(list(restab, restab_newline), use.names = T)
  rm(restab_newline); rm(gdm_output)
}
rm(i); rm(exSplines); rm(a); rm(maxsplines)

# make restab column names nice
setnames(restab, old = as.character(nicenames$names),  new = as.character(nicenames$nicenames), skip_absent = T)
# transpose data.table
restab <- dcast(melt(restab, id.vars = "model"), variable ~ model)
setnames(restab, old = "variable", new = "nicenames")
restab <- merge(nicenames, restab, by = "nicenames")

# save
saveRDS(restab, file = "vignettes/out/plot_GDM_results_EFthresholds.rds")
```
create heatmap
```{r}
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))

include <- c("nicenames", names(restab)[grep("gdm", names(restab))])
# pal <- rev(RColorBrewer::brewer.pal(9, "RdYlBu"))
pal <- RColorBrewer::brewer.pal(9, "YlOrRd")
# pal <- c("white", RColorBrewer::brewer.pal(9, "YlOrRd"))[-2]
# colgrad <- melt(restab2[, ..include], id.vars = c("nicenames", "n_affected_fun"))
# colgrad <- colgrad$value

df <- restab[ground == "a", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA] # 0 values are set to NA to make them appear grey in the plot

a <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 9, hjust = 1),
    plot.margin = margin(l = 50, r = 20),
    axis.line.x=element_blank(),
    axis.text.y = element_text(size=9, angle = 0),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.title = element_blank(),
    legend.position = "top") +
  scale_x_discrete(position = "top")

df <- restab[ground == "b", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
b <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


df <- restab[ground == "x", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
c <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


heatmap <- plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, 
                     nrow = 3, rel_heights = c(0.52, 0.39, 0.09), align = "v")

ggsave(filename = paste(pathtoout, "/plot_GDM_thresholds_heatmap.pdf", sep = ""), plot = heatmap, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```

clustered heatmap
```{r, eval = F}
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))

include <- c("nicenames", names(restab)[grep("gdm", names(restab))])
include <- include[-which(include %in% c("gdm_EFdistance_LUI", "gdm_EFturnover_median_LUI", "gdm_EFnestedness_median_LUI"))]

restab1 <- restab[, ..include]
restab1 <- as.matrix(restab1[, -1])
rownames(restab1) <- restab$nicenames
colnames(restab1) <- gsub("EFnestedness_", "NES ", gsub("EFturnover_", "TO ", 
                                                        gsub("gdm_", "", gsub("_LUI", "", colnames(restab1)))))

heatmap(restab1, hclustfun = hclust)
# manually saved as "plot_GDM_thresholds_heatmap_cluster.pdf
```



# average effects over all thresholds
```{r, eval = F}
# created dataset which is read in in next chunk
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))
include <- c("nicenames", names(restab)[grep("gdm", names(restab))])
include <- include[-which(include %in% c("gdm_EFdistance_LUI", "gdm_EFturnover_median_LUI", "gdm_EFnestedness_median_LUI"))]
restab <- restab[, ..include]
# long format for aggregate
restab1 <- melt(restab, id.vars = "nicenames")
restab1[grep("turnover", variable), component := "EFturnover"]
restab1[grep("nestedness", variable), component := "EFnestedness"]

# calculate average and sd
EFmean <- data.table(aggregate(value ~ component + nicenames, restab1, function(x) mean(x, na.rm = T)))
setnames(EFmean, old = "value", new = "mean")
# EFsd <- data.table(aggregate(value ~ component + nicenames, restab1, sd))
EFsd <- data.table(aggregate(value ~ component + nicenames, restab1, function(x) sd(x)/sqrt(length(x))))
setnames(EFsd, old = "value", new = "sd")

# combine to new data table
EFmean <- merge(EFmean, EFsd, by = c("component", "nicenames"))
rm(EFsd)
t1 <- EFmean[component == "EFturnover", -"component"]
setnames(t1, old = c("mean", "sd"), new = c("EFturnover_mean", "EFturnover_sd"))
t2 <- EFmean[component == "EFnestedness", -"component"]
setnames(t2, old = c("mean", "sd"), new = c("EFnestedness_mean", "EFnestedness_sd"))
EFmean <- merge(t1, t2, by = "nicenames")
rm(t1); rm(t2)

# add to threshold dataset
restab <- merge(restab, EFmean, by = "nicenames")
# save dataset
saveRDS(restab, file = "vignettes/out/plot_GDM_results_EFthresholds_mean_se.rds")
# note : there is also "vignettes/out/plot_GDM_results_EFthresholds_mean_sd.rds" with standard deviation
```
plot
```{r}
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds_mean_se.rds", sep = ""))
restab <- merge(nicenames[, .(nicenames, ground)], restab, by = "nicenames")
# OPTION : select the variables to include in heatmap
#    OPTION 2 : take out "white" from color pal
include <- colnames(restab)[-which(colnames(restab) == "ground")]
# include <- c("nicenames", "EFturnover_mean", "EFturnover_sd", "EFnestedness_mean", "EFnestedness_sd")

# pal <- c("white", RColorBrewer::brewer.pal(9, "YlOrRd"))
pal <- c(RColorBrewer::brewer.pal(9, "YlOrRd"))

# below : code copied from heatmap above -----------------------------
df <- restab[ground == "a", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA] # 0 values are set to NA to make them appear grey in the plot

a <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 9, hjust = 1),
    plot.margin = margin(l = 50, r = 20),
    axis.line.x=element_blank(),
    axis.text.y = element_text(size=9, angle = 0),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.title = element_blank(),
    legend.position = "top") +
  scale_x_discrete(position = "top")

df <- restab[ground == "b", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
b <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


df <- restab[ground == "x", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
c <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))
# above : code copied from heatmap above -----------------------------

heatmap <- plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, 
                     nrow = 3, rel_heights = c(0.48, 0.42, 0.1), align = "v")

# manually saved small heatmap with only mean and sd (no other thesholds) 
#    as : GDM_thresholds_heatmap_mean_sd_only
ggsave(filename = paste(pathtoout, "/plot_GDM_thresholds_heatmap_mean_se.pdf", sep = ""), plot = heatmap, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```


## average effects over all thresholds weighted by model sd


```{r}
# from Gossner 2016
# We also calculated a bootstrapped P value
# for each term in the full GDM, using the gdm.varImp function in the gdm library
# (Supplementary Table 5-2). Additionally we estimated uncertainty for the GDM
# plots by using 100 bootstraps for each model, each time removing 30% of the plot
# pairs and then fitting a GDM and extracting the predictions. We then calculated the
# s.d. of the predictions and added this (±-) to the fitted line (Extended Data Fig. 6).


pdf("test_gdm_plotuncertainty.pdf", paper = "a4r")
gdm::plotUncertainty(gdminput, sampleSites = 0.3, bsIters = 5, geo=T, splineCol = "black")
dev.off()

test <- plotUncertainty_getsd(gdminput, sampleSites = 0.3, bsIters = 5, geo = T, splineCol = "black")
# do for each threshold


#calculate average positive/negative weighted effects 
#a. multiply weights*estimates 
modres$weightEstimate <- modres$Estimate/modres$Std.Error 
modres.abs$weightEstimate<- abs(modres.abs$Estimate/modres.abs$Std.Error) 
#b. sum weight
# Estimate by group and divide by sum of weights, separately for positive and negative 
modres[,w.avg:=sum(weightEstimate)/sum(1/Std.Error),by=list(groups,response,Estimate>0)] 
modres.abs[,w.avg:=sum(weightEstimate)/sum(1/Std.Error),by=list(groups,response)] 
modres.abs <- unique(modres.abs, by=c("groups","response")) 
```



## barplot average effects over all thresholds
```{r}
# Note : need to select the model below

restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds_mean_se.rds", sep = "")) # Note : there is se.rds and sd.rds : standard error or standard deviation
include <- c("nicenames", "EFturnover_mean", "EFturnover_sd", "EFnestedness_mean", "EFnestedness_sd")
restab <- restab[, ..include]
restab <- merge(nicenames, restab, by = "nicenames")

####
# SELECT MODEL
####
# chose variable to plot
# # TURNOVER
# setnames(restab, old = c("EFturnover_mean", "EFturnover_sd"), new = c("maxsplines", "sd"))
# model_name <- "EFturnover_mean_across_thresholds"
# NESTEDNESS
setnames(restab, old = c("EFnestedness_mean", "EFnestedness_sd"), new = c("maxsplines", "sd"))
model_name <- "EFnestedness_mean_across_thresholds"

###########
# produce barplots
###########
p <- create_bio_aboveground_barplot()
b <- create_bio_belowground_barplot()
q <- create_abio_barplot()

###########
# add sd
###########
p <- p + geom_errorbar(aes(ymin = maxsplines - sd, ymax = maxsplines + sd), 
                  linetype = "solid") 
  # ylim(-0.01, 0.45)
b <- b + geom_errorbar(aes(ymin = maxsplines - sd/2, ymax = maxsplines + sd/2), 
                  linetype = "solid") #+
  # ylim(-0.01, 0.45)
q <- q + geom_errorbar(aes(ymin = maxsplines - sd/2, ymax = maxsplines + sd/2), 
                  linetype = "solid") #+
  # ylim(-0.01, 0.45)


# copied from barplot section above --------------------------------------------
###########
# produce overview bars
###########
    # ABOVE - BELOW - ABIOTIC
restab2 <- create_restab2()
restab2 <- restab2[type %in% c("unscaled", "averaged_scaled")]
df <- data.frame(restab2)
ov1 <- create_overview_above_below_abiotic_barplot()
ov1L <- create_overview_above_below_abiotic_barplot(legend = T) # legend only, plot with plot_grid()
    # TURNOVER - NESTEDNESS - ABIOTIC
restab3 <- create_restab_3()
restab3 <- restab3[type %in% c("unscaled", "averaged_scaled")]
df <- data.frame(restab3)
ov2 <- create_overview_turnover_nestedness_abiotic_barplot()
ov2L <- create_overview_turnover_nestedness_abiotic_barplot(legend = T) # only legend, plot with plot_grid()
ov_legends <- plot_grid(ov1L, ov2L, nrow = 2)

###########
# add plots together
###########
multipanel_with_overview_bars <- plot_grid(p, b, q, ov1, ov2, labels = c('A', '', 'B', 'C', ''), label_size = 12, nrow = 5,
                                             rel_heights = c(0.25, 0.4, 0.16, 0.08, 0.08), align = "v") # 2 summary bars

annotated_multipanel_with_overview_bars <- ggdraw(multipanel_with_overview_bars) +
  # draw_plot(labsplot, .65, .65, .35, .35) +
  draw_plot(ov_legends, 0.05, 0.1, .1, .2)

# copied from barplot section above --------------------------------------------

###########
# save
###########
ggsave(filename = paste(pathtoout, "/", model_name, "_multipanel_barplot_overviewbars_not_annotated.pdf", sep = ""), plot = annotated_multipanel_with_overview_bars, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)

```