---
title: "GDM Threshold considerations"
author: 'Noëlle Schenk'
date: "5/27/2021"
output: html_document
---

requirements : 
- TODO user : connect to planteco and nsch to get input files
- GDM output as .rds, e.g. gdm_EFnestedness_LUI_input.Rds
    - find at planteco/.../BetaDivMultifun/analysis/output_datasets
- cluster output as .rds, e.g. gdm_EFnestedness_LUI_permutation.Rds
- run `results_nonpublic.R` : load pathtodata, nicenames and model_names
  - pathtodata : path to location where gdm input of model is stored. points to a folder, where the folder cluster witht the model input is. *note : was pathtoout before.*
  - nicenames : helper file to get nice names of functions and trophic levels
  - model_names : helper file for automation. Contains 3 columns : `name`  contains the all model names for `model_name`, `luiinput` contains either "lui" or "components" as LUI input for `luiinput` and `permut` is either TRUE or FALSE.
- `analysis_nonpublic.R` loads `EFmaster_all_thresholds.Rds`


# create sensitivity comparison: deviance explained, sdev
```{r, eval = F}
#############################
#
# treshold selection
#
#############################

# models with function calculated with threshold : 
# thresholds between 0.1 and 0.9.
# need selection : by (new) : sdev of input. old : percent deviance explained.

# VARIABILITY OF RESPONSE VARIABLE
# require : input data : EFmaster_all_thresholds.Rds : load from script "analysis_nonpublic.R"
EFmaster_sd <- apply(EFmaster[, -c("Var1", "Var2")], 2, function(x) sd(x, na.rm = F))
EFmaster_var <- apply(EFmaster[, -c("Var1", "Var2")], 2, function(x) var(x, na.rm = F))
EFmaster <- data.table(name = names(EFmaster)[-(1:2)], sd = EFmaster_sd, var = EFmaster_var)
EFmaster[, type := "EFdistance"]
EFmaster[grep("nestedness", EFmaster$name), type := "nestedness"]
EFmaster[grep("turnover", EFmaster$name), type := "turnover"]
EFmaster[grep("beta", EFmaster$name), type := "beta"]

ggplot(EFmaster[type != "beta"], aes(x = name, y = sd, fill = type)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


# PERCENT DEVIANCE EXPLAINED
# require : 
# model_results : the model results saved manually from the models
setorder(model_results, cols = "name")
model_results[, EF := factor(EF, ordered = T, levels = c("distance", "singleEF", "abund", "beta", "turnover", "nestedness"))]
str(model_results)
model_results <- model_results[!is.na(dev_expl_by_GDM)]

devexpl <- ggplot(model_results, aes(x = name, y = dev_expl_by_GDM, fill = EF)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(filename = paste(pathtoout, "/plot_GDM_EFbeta_tresholds_var_expl.pdf", sep = ""), 
       plot = devexpl, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)

#TODO can add nice names at some point
# STANDARD DEVIATION AND PERCENT DEVIANCE EXPLAINED
model_results[, name := gsub("_LUI", "", model_results$name)]
model_results[, name := gsub("gdm_", "", model_results$name)]
model_results <- merge(model_results, EFmaster, by = "name")

sdplot <- ggplot(model_results, aes(x = name, y = sd, fill = EF)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
threshold_selection <- plot_grid(sdplot, devexpl, nrow = 2)

ggsave(filename = paste(pathtoout, "/plot_GDM_EFbeta_tresholds_var_expl_and_sd.pdf", sep = ""), 
       plot = threshold_selection, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```



# thresholds EF heatmap

```{r, eval = F}
# dataset is stored, next chunk reads it in
mod_names <- model_results[EF %in% c("distance", "nestedness", "turnover") & luiinput == "lui", name]
mod_names <- mod_names[-grep("0.75", mod_names)]

i <- mod_names[1]
a <- paste(pathtodata, "/analysis/output_datasets/", i, "_output.Rds", sep = "")
gdm_output <- readRDS(a)
# get maxsplines
exSplines <- gdm::isplineExtract(gdm_output)
# take maximum of each variable for barplot
maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
restab <- data.table(model = i, data.frame(as.list(maxsplines)))

for(i in mod_names[-1]){
  a <- paste(pathtodata, "/analysis/output_datasets/", i, "_output.Rds", sep = "")
  gdm_output <- readRDS(a)
  # get maxsplines
  exSplines <- gdm::isplineExtract(gdm_output)
  # take maximum of each variable for barplot
  maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
  restab_newline <- data.table(model = i, data.frame(as.list(maxsplines)))
  # append to restab
  restab <- rbindlist(list(restab, restab_newline), use.names = T)
  rm(restab_newline); rm(gdm_output)
}
rm(i); rm(exSplines); rm(a); rm(maxsplines)

# make restab column names nice
setnames(restab, old = as.character(nicenames$names),  new = as.character(nicenames$nicenames), skip_absent = T)
# transpose data.table
restab <- dcast(melt(restab, id.vars = "model"), variable ~ model)
setnames(restab, old = "variable", new = "nicenames")
restab <- merge(nicenames, restab, by = "nicenames")

# save
saveRDS(restab, file = "vignettes/out/plot_GDM_results_EFthresholds.rds")
```

```{r}
# Load data :
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))
# need to add component column
restab <- merge(restab, nicenames[, .(names, nicenames, ground, component)], by = c("names", "nicenames", "ground"))
fwrite(restab, file = paste(pathtoout, "/all_LUI_mods_results_overview.csv", sep = ""), sep = ",") # write for supplementary
```


create heatmap
```{r, eval = F}
include <- c("nicenames", names(restab)[grep("gdm", names(restab))])
# pal <- rev(RColorBrewer::brewer.pal(9, "RdYlBu"))
pal <- RColorBrewer::brewer.pal(9, "YlOrRd")
# pal <- c("white", RColorBrewer::brewer.pal(9, "YlOrRd"))[-2]
# colgrad <- melt(restab2[, ..include], id.vars = c("nicenames", "n_affected_fun"))
# colgrad <- colgrad$value

df <- restab[ground == "a", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA] # 0 values are set to NA to make them appear grey in the plot

a <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 9, hjust = 1),
    plot.margin = margin(l = 50, r = 20),
    axis.line.x=element_blank(),
    axis.text.y = element_text(size=9, angle = 0),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.title = element_blank(),
    legend.position = "top") +
  scale_x_discrete(position = "top")

df <- restab[ground == "b", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
b <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


df <- restab[ground == "x", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
c <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


heatmap <- plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, 
                     nrow = 3, rel_heights = c(0.52, 0.39, 0.09), align = "v")

ggsave(filename = paste(pathtoout, "/plot_GDM_thresholds_heatmap.pdf", sep = ""), plot = heatmap, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```

clustered heatmap
```{r, eval = F}
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))

include <- c("nicenames", names(restab)[grep("gdm", names(restab))])
include <- include[-which(include %in% c("gdm_EFdistance_LUI", "gdm_EFturnover_median_LUI", "gdm_EFnestedness_median_LUI"))]

restab1 <- restab[, ..include]
restab1 <- as.matrix(restab1[, -1])
rownames(restab1) <- restab$nicenames
colnames(restab1) <- gsub("EFnestedness_", "NES ", gsub("EFturnover_", "TO ", 
                                                        gsub("gdm_", "", gsub("_LUI", "", colnames(restab1)))))

heatmap(restab1, hclustfun = hclust)
# manually saved as "plot_GDM_thresholds_heatmap_cluster.pdf
```
Create overviewbars
```{r, eval = F}
# Load data :
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))
# need to add component column
restab <- merge(restab, nicenames[, .(names, nicenames, ground, component)], by = c("names", "nicenames", "ground"))
# select the relevant columns
include <- grep("median", names(restab), value = T, invert = T)
restab <- restab[, ..include]
rm(include)

a <- create_single_funs_overviewbars(restab2 = restab, 
                                     rel_colnames = grep("gdm_", names(restab), value = T))
# bring to ggplot format
ov_ab_allthresholds <- melt(a$above_below, id.vars = c("ground", "color"))
ov_tn_allthresholds <- melt(a$turnover_nestedess, id.vars = c("component", "color"))
# create plots
p_ov_ab_thres <- create_single_funs_overviewbar_plot(singleF_restab = ov_ab_allthresholds, pos = "dodge")
p_ov_tn_thres <- create_single_funs_overviewbar_plot(singleF_restab = ov_tn_allthresholds, pos = "dodge")

p_ov_abtn_thres <- plot_grid(p_ov_ab_thres, p_ov_tn_thres, labels = c("A", "B"))
# saved as : <date>_GDM_multifun_thresholds_allthresholds_overview_bars<_dodge>
```



# calc average effects over all thresholds
```{r, eval = F}
# created dataset which is read in in next chunk
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds.rds", sep = ""))
include <- c("nicenames", names(restab)[grep("gdm", names(restab))])
include <- include[-which(include %in% c("gdm_EFdistance_LUI", "gdm_EFturnover_median_LUI", "gdm_EFnestedness_median_LUI"))]
restab <- restab[, ..include]
# long format for aggregate
restab1 <- melt(restab, id.vars = "nicenames")
restab1[grep("turnover", variable), component := "EFturnover"]
restab1[grep("nestedness", variable), component := "EFnestedness"]

# calculate average and sd
EFmean <- data.table(aggregate(value ~ component + nicenames, restab1, function(x) mean(x, na.rm = T)))
setnames(EFmean, old = "value", new = "mean")
# EFsd <- data.table(aggregate(value ~ component + nicenames, restab1, sd))
EFsd <- data.table(aggregate(value ~ component + nicenames, restab1, function(x) sd(x)/sqrt(length(x))))
setnames(EFsd, old = "value", new = "sd")

# combine to new data table
EFmean <- merge(EFmean, EFsd, by = c("component", "nicenames"))
rm(EFsd)
t1 <- EFmean[component == "EFturnover", -"component"]
setnames(t1, old = c("mean", "sd"), new = c("EFturnover_mean", "EFturnover_sd"))
t2 <- EFmean[component == "EFnestedness", -"component"]
setnames(t2, old = c("mean", "sd"), new = c("EFnestedness_mean", "EFnestedness_sd"))
EFmean <- merge(t1, t2, by = "nicenames")
rm(t1); rm(t2)

# add to threshold dataset
restab <- merge(restab, EFmean, by = "nicenames")
# save dataset
saveRDS(restab, file = "vignettes/out/plot_GDM_results_EFthresholds_mean_se.rds")
# note : there is also "vignettes/out/plot_GDM_results_EFthresholds_mean_sd.rds" with standard deviation
```

## Heatmap
```{r, eval = F}
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/plot_GDM_results_EFthresholds_mean_se.rds", sep = ""))
restab <- merge(nicenames[, .(nicenames, ground)], restab, by = "nicenames")
# OPTION : select the variables to include in heatmap
#    OPTION 2 : take out "white" from color pal
include <- colnames(restab)[-which(colnames(restab) == "ground")]
# include <- c("nicenames", "EFturnover_mean", "EFturnover_sd", "EFnestedness_mean", "EFnestedness_sd")

# pal <- c("white", RColorBrewer::brewer.pal(9, "YlOrRd"))
pal <- c(RColorBrewer::brewer.pal(9, "YlOrRd"))

# below : code copied from heatmap above -----------------------------
df <- restab[ground == "a", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA] # 0 values are set to NA to make them appear grey in the plot

a <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 9, hjust = 1),
    plot.margin = margin(l = 50, r = 20),
    axis.line.x=element_blank(),
    axis.text.y = element_text(size=9, angle = 0),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.title = element_blank(),
    legend.position = "top") +
  scale_x_discrete(position = "top")

df <- restab[ground == "b", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
b <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


df <- restab[ground == "x", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames"))
df[value == 0, value := NA]
c <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))
# above : code copied from heatmap above -----------------------------

heatmap <- plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, 
                     nrow = 3, rel_heights = c(0.48, 0.42, 0.1), align = "v")

# manually saved small heatmap with only mean and sd (no other thesholds) 
#    as : GDM_thresholds_heatmap_mean_sd_only
ggsave(filename = paste(pathtoout, "/plot_GDM_thresholds_heatmap_mean_se.pdf", sep = ""), plot = heatmap, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```




## average effects over all thresholds weighted by model sd

from Gossner 2016
We also calculated a bootstrapped P value
for each term in the full GDM, using the gdm.varImp function in the gdm library
(Supplementary Table 5-2). Additionally we estimated uncertainty for the GDM
plots by using 100 bootstraps for each model, each time removing 30% of the plot
pairs and then fitting a GDM and extracting the predictions. We then calculated the
s.d. of the predictions and added this (±-) to the fitted line (Extended Data Fig. 6).

runned permutations on the cluster with following parameters :
`plotUncertainty_getsd(gdm_model, sampleSites = 0.3, bsIters = 100, geo = T, splineCol = "black", cores = 2)`
results are stored at :
planteco/.../BetaDivMultifun/analysis/output_datasets/uncertainty_calc

reading in data
```{r}
# Note : go through code and either chose to work with EFnestedness, or EFturnover. (by outcommenting respective lines)

# modnames <- paste(paste("gdm_EFturnover_", seq(0.1, 0.9, 0.1), sep = ""), "_LUI", sep = "")
modnames <- paste(paste("gdm_EFnestedness_", seq(0.1, 0.9, 0.1), sep = ""), "_LUI", sep = "")
i <- 1
res <- readRDS(paste(paste(paste(pathtodata, "/analysis/output_datasets/uncertainty_calc/", sep = ""), modnames[i], sep = ""), "_uncertainty.Rds", sep = ""))
res[, component := modnames[i]]

for(i in 2:9){
  t <- readRDS(paste(paste(paste(pathtodata, "/analysis/output_datasets/uncertainty_calc/", sep = ""), modnames[i], sep = ""), "_uncertainty.Rds", sep = ""))
  t[, component := modnames[i]]
  res <- rbindlist(list(res, t))
}
```
plotting
```{r, eval = F}
a <- ggplot(res[varname %in% nicenames[ground == "a", names]], aes(x = varname, y = yatmax, fill = component)) +
  geom_bar(stat = "identity", position="dodge") +
  coord_flip() +
  geom_errorbar(aes(ymin = yatmax - sdy, ymax = yatmax + sdy), 
                  linetype = "solid",
                position=position_dodge(width=0.9)) +
  scale_fill_brewer(palette="Set3")
b1 <- ggplot(res[varname %in% nicenames[ground == "b", names][1:10]], aes(x = varname, y = yatmax, fill = component)) +
  geom_bar(stat = "identity", position="dodge") +
  coord_flip() +
  geom_errorbar(aes(ymin = yatmax - sdy, ymax = yatmax + sdy), 
                  linetype = "solid",
                position=position_dodge(width=0.9)) +
  scale_fill_brewer(palette="Set3")
b2 <- ggplot(res[varname %in% nicenames[ground == "b", names][11:20]], aes(x = varname, y = yatmax, fill = component)) +
  geom_bar(stat = "identity", position="dodge") +
  coord_flip() +
  geom_errorbar(aes(ymin = yatmax - sdy, ymax = yatmax + sdy), 
                  linetype = "solid",
                position=position_dodge(width=0.9)) +
  scale_fill_brewer(palette="Set3")
x <- ggplot(res[varname %in% nicenames[ground == "x", names]], aes(x = varname, y = yatmax, fill = component)) +
  geom_bar(stat = "identity", position="dodge") +
  coord_flip() +
  geom_errorbar(aes(ymin = yatmax - sdy, ymax = yatmax + sdy), 
                  linetype = "solid",
                position=position_dodge(width=0.9)) +
  scale_fill_brewer(palette="Set3")
# saved by hand and combined to 1 pdf with the name :
#  "GDM_multifun_thresholds_weighted_avg_over_all_thresholds_plotUncertainty_test_intermediate.pdf"
```
Calculate weighted average
$$wa = \frac{e_{1}*w_{1} + e_{2}*w_{2} + ... + e_{9}*w_{9}}{w_{1}+w_{2} + ... + w_{9}} = \frac{\sum_{i=1}^{n} e_{n}+\frac{1}{sd_{n}}}{\sum_{i=1}^{n} \frac{1}{sd_{n}}} ;  w_{n} = \frac{1}{sd_{n}} $$
uncertainty-weighted average effect size.
```{r, eval = F}
######
# calculate WEIGHTED AVERAGE
res[, weights := sdy ^ (-1)]
# if sdy is zero, set weight to 0
res[sdy == 0, weights := 0]
res[, weighted_yatmax := yatmax * weights]
# every variable has its own weight - calculate sum of weights for each variable:
sum_of_weights <- aggregate(weights ~ varname, res, sum)

wres <- data.table(aggregate(weighted_yatmax ~ varname, res, sum))

wres <- merge(wres, sum_of_weights, by = "varname")
wres[, wres := weighted_yatmax / weights]

# test one variable and check results
i <-  "bacterivore.protist.beta.sim"
sum(res[varname == i, yatmax] * (1 / res[varname == i, sdy])) / sum((1 / res[varname == i, sdy]))
# True for Geographic, autotroph.beta.sim, bacterivore.protist.beta.sim

# merge with nicenames to get stuff for barplots
setnames(wres, old = "varname", new = "names")
wres <- merge(wres[, .(names, wres)], nicenames, by = "names")

# saveRDS(wres, file = "vignettes/out/weighted_avg_over_all_thresholds_EFturnover.Rds")
# saveRDS(wres, file = "vignettes/out/weighted_avg_over_all_thresholds_EFnestedness.Rds")
```
prepare for plotting
```{r}
# restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/weighted_avg_over_all_thresholds_EFturnover.Rds", sep = ""))
restab <- readRDS(paste(pathtodata, "/analysis/output_datasets/weighted_avg_over_all_thresholds_EFnestedness.Rds", sep = ""))
setnames(restab, old = c("wres"), new = c("maxsplines"))
model_name <- "EFturnover_WEIGTHED_mean_across_thresholds"
# model_name <- "EFnestedness_WEIGTHED_mean_across_thresholds"
luiinput <- "lui"
permut <- F

p <- create_bio_aboveground_barplot() + ylim(c(0, 0.15))
b <- create_bio_belowground_barplot() + ylim(c(0, 0.15))
q <- create_abio_barplot() + ylim(c(0, 0.15))

# go to plot_GDM, and run the part under "# Barplots ". Start with section "produce overview bars" (leave out the upper part)
# reset the ylims : for EFnestedness : + ylim(c(0, 0.1)), for EFturnover : + ylim(c(0, 0.15))
#TODO ERROR plotting does not work any more --> make sure it can be plotted normally with plot_GDM. At the moment, create_restab2() is not working.


# Additional : loaded both EFturnover and EFnestedness, merged together and saved as : 
# "EFnestedness_turnover_WEIGTHED_mean_across_thresholds_maxsplines.csv"
```



