---
title: "how to use this package"
author: "noelle"
date: "November 8, 2018"
output: html_document
---

# installing this package as a user
1. Download the whole folder from [github](https://github.com/allanecology/BetaDivMultifun).
2. set the R working directory to the parent folder of the downloaded package. (e.g. Downloads if you stored the package folder in Downloads.)
The devtools::install function also installs packages on which `BetaDivMultifun` depends on.
```{r, eval=F}
setwd("<parent-directory-of-the-package>")
devtools::install("BetaDivMultifun")
```

Set wd *back* to the package
```{r}
setwd("/home/noelleschenk/Desktop/BetaDivMultifun/")
```



# install the package as a developer
## Literature
To create the package, I used the blogposts of: [Hilary Parker](https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/), [the book "R packages"](http://r-pkgs.had.co.nz/) and the idea of using `R packages` for an analysis is coming from Robert M. Flight's blogposts, where he explains [why](http://rmflight.github.io/post/analyses-as-packages/) and [how](http://rmflight.github.io/post/vignette-analysis/) to use `R package` for an analysis.

## Setup
Only if you like to work on the package (change it). I worked with following packages and RStudio, which is extremely well suited for package development.

If you are linux user, install the following packages with `sudo apt-get install <packagename>`
- `libssl-dev`
- `libxml2 libxml2-dev`
- `libcurl4-gnutls-dev` (previously `libcurl`)

If you are not using Linux and there was any requirement, R will tell you in its error message as soon as you try to install the below mentioned packages.
```{r, eval=F}
# running these lines will take some time ...
install.packages("devtools")
library(devtools)
devtools::install_github("klutometis/roxygen")
library(roxygen2)
```


## changing code
After you generated/changed code of the package, the newly generated code needs to be loaded. (E.g. if you add a function and later want to use it.) Therefore, the function code as well as the functions documentation needs to be loaded as described below. 

- create *documentation*
from roxygen comments, with Ctrl-Shift-D or
```{r,eval=F}
devtools::document()
```


- *load functions* which are encoded in the package. `devtools::load_all()` or with Ctrl-Shift-L (while the cursor is placed in the console)

- *install and restart* the package :  Ctrl-Shift-B (console `R CMD INSTALL --no-multiarch --with-keep.source BetaDivMultifun`)

- to *rebuild the source package* .tar.gz file newly : RStudio : Build > Build Source Package `devtools::build()`

### check code
RStudio : Build > Check
--> note : don't confuse with "test", this is something else.

### test code
develop your own tests to prevent introducing bugs.



# creating `assembled_functions` dataset
is described in 3 scripts in another github folder : 
https://github.com/biodiversity-exploratories-synthesis/2019_grassland_functions

In this package, the required dataset is just read in and an analysis is performed. As the dataset was constructed from within this directory, the commits can be looked up here.

After installing the package and being connected to the "planteco" drive at IPS:
* find the file `nonpublic.R`, it is stored at the "planteco" drive where the data is.
* run `nonpublic.R`
* run `1read_raw_datasets.Rmd` then `2calc_raw_dataset.R`
* the file `"~/Desktop/december2018_assembled_functions_dataset.csv"` will be written on your Desktop.


# knit to html
By hand
```{r, eval = F}
rmarkdown::render('<pathtothisdocument>/1read_raw_dataset.Rmd')
rmarkdown::render('vignettes/function_imputation.Rmd', output_format = "html_document")
```

