---
title: "how to use this package"
author: "noelle"
date: "November 8, 2018"
output: html_document
---

# Resources used
To create the package, I used the blogposts of: [Hilary Parker](https://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/), [the book "R packages"](http://r-pkgs.had.co.nz/) and the idea of using `R packages` for an analysis is coming from Robert M. Flight's blogposts, where he explains [why](http://rmflight.github.io/post/analyses-as-packages/) and [how](http://rmflight.github.io/post/vignette-analysis/) to use `R package` for an analysis.

# install required packages
If you are linux user, install the following packages with `sudo apt-get install`
- `libcurl`
- `libssl-dev`
- `libxml2 libxml2-dev`

If you are not using Linux and there was any requirement, R will tell you in its error message as soon as you try to install the below mentioned packages.
```{r, eval=F}
install.packages("devtools")
library(devtools)
devtools::install_github("klutometis/roxygen")
library(roxygen2)
```


# installing this package
1. Download the whole folder of github.
2. set the R working directory to the parent folder of the downloaded package. (e.g. Downloads if you stored the package folder in Downloads.)
```{r, eval=F}
setwd("<parent-directory-of-the-package>")
install("multiFunBetadivLuiGDM")
```

Set wd back to the package
```{r}
setwd("/home/noelleschenk/Desktop/github/allanecology/BetaDivMultifun/")
```


# working on the package
## create documentation
from roxygen comments, with Ctrl-Shift-D or
```{r,eval=F}
devtools:document()
```

## load
with Ctrl-Shift-L or `devtools::load_all()`



