---
title: "Introductory Plots"
author: 'Noëlle Schenk'
date: "3/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Requirements : 
Load the following variables from `results_nonpublic.R`:
- load package (as usual)
- pathtodata
- master_alpha
- trlevels
- plotNAset
- LUI
- EFmastser
- gdm_EFdistance_LUI_input

from `results_nonpublic.R`
- nicenames

# Content
- Diversity
    - ... 


# Clean dataset
```{r}
require(cowplot)
# masterdiversity <- data.table::data.table(masterdiversity)
masterbeta <- data.table::data.table(masterbeta)
LUI <- data.table::data.table(LUI)


# only selected plot set
# masterdiversity <- masterdiversity[Var1 %in% plotNAset & Var2 %in% plotNAset, ]
masterbeta <- masterbeta[Var1 %in% plotNAset & Var2 %in% plotNAset,]

full_master <- merge(EFmaster, masterbeta, by = c("Var1", "Var2"))
any(is.na(full_master))
full_master <- merge(full_master, distLUI, by = c("Var1", "Var2"), all.x = T)
any(is.na(full_master))
temp <- data.table::copy(LUI)
setnames(temp, old = "Plotn", new = "Var1")
full_master <- merge(full_master, temp, by = "Var1")
setnames(temp, old = "Var1", new = "Var2")
full_master <- merge(full_master, temp, by = "Var2")
rm(temp)

# add mean LUI
full_master[, meanLUI := rowMeans(full_master[, .(LUI.x, LUI.y)])]

ggfull_master <- melt(full_master, id.vars = c("Var1", "Var2"))

# get colours for trophic levels
#TODO put right color for each level
# longpallette <- unique(nicenames$color)
#TODO nicenames are shown in a strange way...
longpallette <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#000000", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```



# LUI distance
```{r}
ggplot(full_master, aes(x = LUI.x, y = LUI.y, colour = distLUI)) +
  geom_point() +
  scale_color_gradient(low="blue", high="red") +
  labs(subtitle = c("distLUI is ambiguous, per distLUI value, two possible LUI values."))

# EFdistance and LUI
ggplot(full_master, aes(x = LUI.x, y = LUI.y, colour = EFdistance)) +
  geom_point() + scale_color_gradient(low="blue", high="red")
a <- ggplot(full_master, aes(x = distLUI, y = EFdistance, colour = LUI.y)) +
  geom_point() + scale_color_gradient(low="blue", high="red")
b <- ggplot(full_master, aes(x = distLUI, y = EFdistance, colour = LUI.x)) +
  geom_point() + scale_color_gradient(low="blue", high="red")
plot_grid(a, b, nrow = 1)


# delta LUI vs. mean LUI
ggplot(full_master, aes(x = meanLUI, y = distLUI)) +
  geom_point()
# there are many intermediate cases which we can not distinguish
a <- ggplot(full_master, aes(x = meanLUI, y = distLUI, colour = LUI.y)) +
  geom_point() + 
  scale_color_gradient(low="blue", high="red")
b <- ggplot(full_master, aes(x = meanLUI, y = distLUI, colour = LUI.x)) +
  geom_point() + 
  scale_color_gradient(low="blue", high="red")
plot_grid(a, b, nrow = 2, align = T)
# mit diesen 3 informationen sollte jeder Wert eindeutig zuordenbar sein
# saved as data_assembly/plots/21-03-11_introductory_plots_meanLUI_distLUI_ambiguity.pdf
```



# Diversity

```{r}
master_alpha <- data.table::data.table(master_alpha)
master_alpha <- master_alpha[Var1 %in% plotNAset]
ggmaster_alpha <- melt(master_alpha, id.vars = "Var1")
```


## Trophic levels
```{r}
# edit bacteria.RNA : divide by 100
edit_ggmaster_alpha <- copy(ggmaster_alpha)
edit_ggmaster_alpha[variable == "bacteria.RNA.alpha", value := value / 100]
edit_ggmaster_alpha[variable == "bacteria.RNA.alpha", variable := "bacteria.RNA.alpha.div100"]
```


Example plot AEG01
```{r}
# ultra-high number of bacteria
# show 2 plots, one with all and one with just the small data
plotname <- "AEG48"
a <- ggplot(data = ggmaster_alpha[Var1 == plotname, .(variable, value)], 
            aes(x = variable, y = value, fill = variable)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values = longpallette) +
  labs(title = plotname, x = "", y = "alpha diversity") +
  theme(legend.position = "none")
legend <- cowplot::get_legend(a)
b <- a + ylim(c(0, 480))+ labs(title = "") + theme(legend.position = "none")
c <- a + ylim(c(0, 40)) + labs(title = "") + theme(legend.position = "none")
plot_grid(a, NULL, b, plot_grid(legend), c, NULL, nrow = 3, align = T, rel_widths = c(0.5, 0.2))
# saved as : 21-03-11_introductory_plots_alpha_diversity_plot_AEG01
```

Violin plot over all plots
```{r}
a <- ggplot(data = edit_ggmaster_alpha, aes(x = variable, y = value, fill = variable)) +
  geom_violin() +
  ylim(c(0, 540)) +
  theme(axis.text.x = element_blank(), legend.position = "none") +
  scale_fill_manual(values = longpallette) +
  labs(title = "alpha diversity over all plots", x = "", y = "alpha diversity")
legend <- cowplot::get_legend(a)
b <- a + ylim(c(0, 100)) + labs(title = "")
plot_grid(a, legend, b, NULL, rel_widths = c(0.6, 0.3))
# saved as : 21-03-11_introductory_plots_alpha_diversity_over_all_plots
```

# Functions

## single functions
```{r}
#TODO example plot like the one for alphadiversity for a given plot
#    do the same with functions
```

Violin plot over all plots
```{r}
#TODO
```

EFturnover and EFnestedness
```{r}
ggplot(full_master, aes(x = EFturnover_0.7, y = EFnestedness_0.7, colour = EFbeta_0.7)) +
  geom_point()
```


# LUI
```{r}

```


# case descriptions

## high betadiversity
```{r}
par(mfrow = c(2,1))
plot(full_master$autotroph.beta.sim, 
     main = "Autotroph betadiversity", sub = "turnover (above) and nestedness (below)", 
     xlab = "pairwise comparisons of plots", ylab = "autotroph betadiversity")
plot(full_master$autotroph.beta.sne,
     xlab = "pairwise comparisons of plots", ylab = "autotroph betadiversity")

ggplot(full_master, aes(x = autotroph.beta.sim, y = autotroph.beta.sne, colour = EFbeta_0.7)) +
  scale_color_gradient(low="blue", high="red") +
  geom_point()
# saved as : 21-03-11_introductory_plots_autotroph_betadiversity_turnover_nestedness_EFdist
```

High autotroph betadiversity turnover
```{r}
test <- full_master[autotroph.beta.sim > 0.95, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest autotroph beta turnover",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_autotroph_betadiversity_turnover_EFdist_LUI
```

High autotroph betadiversity nestedness
```{r}
test <- full_master[autotroph.beta.sne > 0.4, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest autotroph beta nestedness",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_autotroph_betadiversity_nestedness_EFdist_LUI
```



## high EFdist
```{r}
plot(full_master$EFdistance)

test <- full_master[EFdistance > 0.93, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest EFdistance",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_EFdist_LUI_selected_beta
```

## high EFnestedness
```{r}
plot(full_master$EFnestedness_0.7)

test <- full_master[EFnestedness_0.7 > 0.7, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- test[sample(1:nrow(test), 8), ]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest EFnestedness0.7",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_EFnestedness_LUI_selected_beta
```




# trying out
```{r}
ggplot(full_master, aes(x = distLUI, y = EFdistance, colour = autotroph.beta.sne)) +
  geom_point()

a <- ggplot(full_master, aes(x = distLUI, y = autotroph.beta.sne, colour = EFdistance)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
b <- ggplot(full_master, aes(x = distLUI, y = autotroph.beta.sim, colour = EFdistance)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```

moving window of EFdist for LUI autotroph beta connection
```{r}
a <- ggplot(full_master[EFdistance < 0.2], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0, 0.2}")
b <- ggplot(full_master[EFdistance > 0.2 & EFdistance < 0.4], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.2, 0.4}")
c <- ggplot(full_master[EFdistance > 0.4 & EFdistance < 0.6], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.4, 0.6}")
d <- ggplot(full_master[EFdistance > 0.6 & EFdistance < 0.75], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.6, 0.75}")
e <- ggplot(full_master[EFdistance > 0.75], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.75, 1}")

plot_grid(a, b, c, d, e, NULL, NULL, NULL, NULL, NULL, nrow = 2)
```

