---
title: "Introductory Plots"
author: 'Noëlle Schenk'
date: "3/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Requirements : 
Load the following variables from `analysis_nonpublic.R`:
- load package (as usual)
- pathtodata
- master_alpha
- trlevels
- plotNAset
- LUI
- distLUI
- EFmastser
- gdm_EFdistance_LUI_input

from `results_nonpublic.R`
- nicenames

# Content
- Diversity
    - ... 


# Clean dataset
```{r}
require(cowplot)
# masterdiversity <- data.table::data.table(masterdiversity)
masterbeta <- data.table::data.table(masterbeta)
LUI <- data.table::data.table(LUI)


# only selected plot set
# masterdiversity <- masterdiversity[Var1 %in% plotNAset & Var2 %in% plotNAset, ]
masterbeta <- masterbeta[Var1 %in% plotNAset & Var2 %in% plotNAset,]

full_master <- merge(EFmaster, masterbeta, by = c("Var1", "Var2"))
any(is.na(full_master))
full_master <- merge(full_master, distLUI, by = c("Var1", "Var2"), all.x = T)
any(is.na(full_master))
temp <- data.table::copy(LUI)
setnames(temp, old = "Plotn", new = "Var1")
full_master <- merge(full_master, temp, by = "Var1")
setnames(temp, old = "Var1", new = "Var2")
full_master <- merge(full_master, temp, by = "Var2")
rm(temp)

# add mean LUI
full_master[, meanLUI := rowMeans(full_master[, .(LUI.x, LUI.y)])]

ggfull_master <- melt(full_master, id.vars = c("Var1", "Var2"))

# get colours for trophic levels
#TODO put right color for each level
# longpallette <- unique(nicenames$color)
#TODO nicenames are shown in a strange way...
longpallette <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#000000", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```



# LUI distance
```{r}
ggplot(full_master, aes(x = LUI.x, y = LUI.y, colour = distLUI)) +
  geom_point() +
  scale_color_gradient(low="blue", high="red") +
  labs(subtitle = c("distLUI is ambiguous, per distLUI value, two possible LUI values."))

# EFdistance and LUI
ggplot(full_master, aes(x = LUI.x, y = LUI.y, colour = EFdistance)) +
  geom_point() + scale_color_gradient(low="blue", high="red")
a <- ggplot(full_master, aes(x = distLUI, y = EFdistance, colour = LUI.y)) +
  geom_point() + scale_color_gradient(low="blue", high="red")
b <- ggplot(full_master, aes(x = distLUI, y = EFdistance, colour = LUI.x)) +
  geom_point() + scale_color_gradient(low="blue", high="red")
plot_grid(a, b, nrow = 1)


# delta LUI vs. mean LUI
ggplot(full_master, aes(x = meanLUI, y = distLUI)) +
  geom_point()
# there are many intermediate cases which we can not distinguish
a <- ggplot(full_master, aes(x = meanLUI, y = distLUI, colour = LUI.y)) +
  geom_point() + 
  scale_color_gradient(low="blue", high="red")
b <- ggplot(full_master, aes(x = meanLUI, y = distLUI, colour = LUI.x)) +
  geom_point() + 
  scale_color_gradient(low="blue", high="red")
plot_grid(a, b, nrow = 2, align = T)
# mit diesen 3 informationen sollte jeder Wert eindeutig zuordenbar sein
# saved as data_assembly/plots/21-03-11_introductory_plots_meanLUI_distLUI_ambiguity.pdf
```



# Diversity

```{r}
master_alpha <- data.table::data.table(master_alpha)
master_alpha <- master_alpha[Var1 %in% plotNAset]
ggmaster_alpha <- melt(master_alpha, id.vars = "Var1")
```


## Trophic levels
```{r}
# edit bacteria.RNA : divide by 100
edit_ggmaster_alpha <- copy(ggmaster_alpha)
edit_ggmaster_alpha[variable == "bacteria.RNA.alpha", value := value / 100]
edit_ggmaster_alpha[variable == "bacteria.RNA.alpha", variable := "bacteria.RNA.alpha.div100"]
```


Example plot AEG01
```{r}
# ultra-high number of bacteria
# show 2 plots, one with all and one with just the small data
plotname <- "AEG48"
a <- ggplot(data = ggmaster_alpha[Var1 == plotname, .(variable, value)], 
            aes(x = variable, y = value, fill = variable)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_blank()) +
  scale_fill_manual(values = longpallette) +
  labs(title = plotname, x = "", y = "alpha diversity") +
  theme(legend.position = "none")
legend <- cowplot::get_legend(a)
b <- a + ylim(c(0, 480))+ labs(title = "") + theme(legend.position = "none")
c <- a + ylim(c(0, 40)) + labs(title = "") + theme(legend.position = "none")
plot_grid(a, NULL, b, plot_grid(legend), c, NULL, nrow = 3, align = T, rel_widths = c(0.5, 0.2))
# saved as : 21-03-11_introductory_plots_alpha_diversity_plot_AEG01
```

Violin plot over all plots
```{r}
a <- ggplot(data = edit_ggmaster_alpha, aes(x = variable, y = value, fill = variable)) +
  geom_violin() +
  ylim(c(0, 540)) +
  theme(axis.text.x = element_blank(), legend.position = "none") +
  scale_fill_manual(values = longpallette) +
  labs(title = "alpha diversity over all plots", x = "", y = "alpha diversity")
legend <- cowplot::get_legend(a)
b <- a + ylim(c(0, 100)) + labs(title = "")
plot_grid(a, legend, b, NULL, rel_widths = c(0.6, 0.3))
# saved as : 21-03-11_introductory_plots_alpha_diversity_over_all_plots
```

# Functions

## Violin plot overview all functions
```{r}
# summary stats over EFmaster
a <- apply(EFmaster, 2, max)
data.frame(a[3:30])


# % turnover and nestedness
restab_percent_to_nes <- EFmaster[, .(Var1, Var2)]
for(t in seq(0.1, 0.9, 0.1)){
  i <- grep(t, names(EFmaster), value = T)
  subset <- EFmaster[, ..i]
  setnames(subset, old = i, new = sub("_", "", sub(t, "", i)))
  subset[, paste("percent_turnover", t, sep = "_") := 100 / EFbeta * EFturnover]
  subset[, paste("percent_nestedness", t, sep = "_") := 100 / EFbeta * EFnestedness]
  subset[apply(subset, 2, is.nan)] <- 0
  # check
  subset[, check := 100 / EFbeta * EFturnover + 100 / EFbeta * EFnestedness]
  plot(subset$check) # all values are 100
  # rename and add to new dataset
  i <- c("EFbeta", "EFturnover", "EFnestedness", "check")
  subset[, (i) := NULL]
  restab_percent_to_nes <- cbind(restab_percent_to_nes, subset)
  rm(subset)
}
# summary stats
summary(restab_percent_to_nes)

# visualise results
ggplot(restab_percent_to_nes, aes(x = percent_turnover_0.1, y = percent_nestedness_0.1)) +
  geom_jitter(width = 8, shape = 1) +
  xlim(c(0, 100)) +
  ggtitle(label = "threshold = 0.1")

ggplot(restab_percent_to_nes, aes(x = percent_turnover_0.9, y = percent_nestedness_0.9)) +
  geom_jitter(width = 8, shape = 1) +
  xlim(c(0, 100)) +
  ggtitle(label = "threshold = 0.9")

plot(restab_percent_to_nes$percent_turnover_0.2, restab_percent_to_nes$percent_nestedness_0.2)
plot(restab_percent_to_nes$percent_turnover_0.1, restab_percent_to_nes$percent_nestedness_0.1)


restab_melted <- melt(restab_percent_to_nes, id.vars = c("Var1", "Var2"))
p <- ggplot(restab_melted, aes(x=variable, y=value)) + 
  geom_violin(fill = "#A4A4A4") +
  coord_flip()


restab_melted[, component := "nestedness"]
restab_melted[grep("turnover", restab_melted$variable), component := "turnover"]
table(restab_melted$component) # quick check
aggregate(value ~ component, restab_melted, mean)
aggregate(value ~ component, restab_melted, sd)
```

## correlations among beta-multifunctionalities
```{r}
a <- ggplot(EFmaster, aes(x = EFturnover_0.8, y = EFnestedness_0.8, colour = EFbeta_0.8)) +
  geom_jitter(aes(size = EFbeta_0.8)) +
  scale_color_gradient(low="lightseagreen", high="orangered3")

b <- ggplot(EFmaster, aes(x = EFturnover_0.8, y = EFnestedness_0.8, colour = EFdistance)) +
  geom_jitter(aes(size = EFdistance)) +
  scale_color_gradient(low="lightseagreen", high="orangered3")

plot_grid(a, b)
```

```{r}
M <- cor(EFmaster[, -"Var1"][, -"Var2"])
corrplot::corrplot(M,type="lower", addCoef.col = "black", method="color", diag=F, tl.srt=1, tl.col="black", mar=c(0,0,0,0), number.cex=0.4, order = "hclust", tl.cex = 0.6)

corrplot::corrplot(M,type="lower", addCoef.col = "black", method="color", diag=F, tl.srt=1, tl.col="black", mar=c(0,0,0,0), number.cex=0.4, tl.cex = 0.6)
```



## single functions present/ absent along LUI level
```{r}
# read in dataset
lui <- data.table(readRDS(paste(pathtodata, "/data_assembly/output_data/LUI.rds", sep = "")))
funs <- readRDS(paste(pathtodata, "/data_assembly/output_data/imputed_and_compound_grlfuns.rds", sep = ""))

# FUNCTION LEVELS
# create heatmap
dat <- merge(lui[, .(Plotn, LUI)], funs, by = "Plotn")
setorder(dat, LUI) # sort by LUI
mdat <- as.matrix(t(dat[, -"Plotn"]))
colnames(mdat) <- dat$Plotn
heatmap(mdat, scale = "row", Colv = NA, main = "Function levels along the LUI gradient")# , Rowv = NA)
# saved as : introductory_plots_heatmap_function_levels_along_LUI_gradient.pdf
#    went through differen thresholds and combined the output pdfs to 1 : introductory_plots_heatmap_functions_along_LUI_gradient

# FUNCTION PRESENCE-ABSENCE
# calc presence-absence based on given threshold
funs_pa <- data.table::copy(funs)
threshold <- 0.2
include <- names(funs_pa)[!names(funs_pa) %in% "Plotn"]
funs_pa[, (include) := lapply(.SD, function(c) calc_presenceabsence(c, threshold = threshold, type = "max5")), .SDcols = include]
funs_pa[, (include) := (.SD * 1), .SDcols = include]
# create heatmap
dat <- merge(lui[, .(Plotn, LUI)], funs_pa, by = "Plotn")
setorder(dat, LUI) # sort by LUI
mdat <- as.matrix(t(dat[, -"Plotn"]))
colnames(mdat) <- dat$Plotn
heatmap(mdat, scale = "none", Colv = NA, main = paste("threshold :", threshold), xlab = "LUI: AEG02 has highest LUI")
# heatmap(mdat, scale = "none")
# saved as : introductory_plots_heatmap_function_pa0.3_along_LUI_gradient
#    went through differen thresholds and combined the output pdfs to 1 : introductory_plots_heatmap_functions_along_LUI_gradient
```

## single functions change at high EFdist
Plot comparisons with high EFdist (> 0.8): 

High EFdistance, high turnover and low nestedness
SEG17 AEG02  1.0000000      0.8333333       0.01282051
HEG38 AEG02  0.8628104      0.8000000       0.03333333

High EFdistance, low turnover and high nestedness 
(nes goes only until 0.55, so this is very high)
SEG22 AEG49  0.8823974      0.3333333       0.26666667
SEG22 HEG30  0.8545400      0.3333333       0.26666667

Filter for high nestedness: 
if nes is at its max, turnover is always 0!
HEG11 AEG19  0.3787118              0        0.5555556
SEG22 HEG20  0.7917539              0        0.5555556
also if nestedness is > 0.4
--> high nes cases are 0 turnover cases!
```{r}
#TODO HERE : summary stats of these cases.
# note that somehow.

# EFmastser : read in from analysis_nonpublic.R
# find high distance comparisons
highdist <- EFmaster[EFdistance > 0.85, .(Var1, Var2, EFdistance, EFturnover_0.6, EFnestedness_0.6)]
highdist <- EFmaster[EFnestedness_0.6 > 0.5, .(Var1, Var2, EFdistance, EFturnover_0.6, EFnestedness_0.6)]
highdist <- EFmaster[EFnestedness_0.6 > 0.4 & EFnestedness_0.6 < 0.55, .(Var1, Var2, EFdistance, EFturnover_0.6, EFnestedness_0.6)]
unique(highdist$EFturnover_0.6)
# HEG38 AEG02  : 0.8628104 (EFdistance), EFturnover > 0.8, EFnes : 0.03 (very low)

a <- ggplot(EFmaster, aes(x = EFdistance, y = EFturnover_0.1, colour = EFnestedness_0.1)) +
  geom_jitter(aes(size = EFnestedness_0.1)) +
  scale_color_gradient(low="lightseagreen", high="orangered3")

b <- ggplot(EFmaster, aes(x = EFdistance, y = EFturnover_0.6, colour = EFnestedness_0.6)) +
  geom_jitter(aes(size = EFnestedness_0.6)) +
  scale_color_gradient(low="lightseagreen", high="orangered3")

c <- ggplot(EFmaster, aes(x = EFdistance, y = EFturnover_0.7, colour = EFnestedness_0.7)) +
  geom_jitter(aes(size = EFnestedness_0.7)) +
  scale_color_gradient(low="lightseagreen", high="orangered3")

d <- ggplot(EFmaster, aes(x = EFdistance, y = EFturnover_0.8, colour = EFnestedness_0.8)) +
  geom_jitter(aes(size = EFnestedness_0.8)) +
  scale_color_gradient(low="lightseagreen", high="orangered3")
plot_grid(a, b, c, d, nrow = 2)

data.table(readRDS(paste(pathtodata, "/data_assembly/output_data/LUI.rds", sep = "")))
funs <- readRDS(paste(pathtodata, "/data_assembly/output_data/imputed_and_compound_grlfuns.rds", sep = ""))

```



Violin plot over all plots
```{r}
#TODO
```

EFturnover and EFnestedness
```{r}
ggplot(full_master, aes(x = EFturnover_0.7, y = EFnestedness_0.7, colour = EFbeta_0.7)) +
  geom_point()
```


# LUI
```{r}

```


# case descriptions

## high betadiversity
```{r}
par(mfrow = c(2,1))
plot(full_master$autotroph.beta.sim, 
     main = "Autotroph betadiversity", sub = "turnover (above) and nestedness (below)", 
     xlab = "pairwise comparisons of plots", ylab = "autotroph betadiversity")
plot(full_master$autotroph.beta.sne,
     xlab = "pairwise comparisons of plots", ylab = "autotroph betadiversity")

ggplot(full_master, aes(x = autotroph.beta.sim, y = autotroph.beta.sne, colour = EFbeta_0.7)) +
  scale_color_gradient(low="blue", high="red") +
  geom_point()
# saved as : 21-03-11_introductory_plots_autotroph_betadiversity_turnover_nestedness_EFdist
```

High autotroph betadiversity turnover
```{r}
test <- full_master[autotroph.beta.sim > 0.95, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest autotroph beta turnover",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_autotroph_betadiversity_turnover_EFdist_LUI
```

High autotroph betadiversity nestedness
```{r}
test <- full_master[autotroph.beta.sne > 0.4, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest autotroph beta nestedness",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_autotroph_betadiversity_nestedness_EFdist_LUI
```



## high EFdist
```{r}
plot(full_master$EFdistance)

test <- full_master[EFdistance > 0.93, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest EFdistance",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_EFdist_LUI_selected_beta
```

## high EFnestedness
```{r}
plot(full_master$EFnestedness_0.7)

test <- full_master[EFnestedness_0.7 > 0.7, .(Var1, Var2, EFdistance, EFbeta_0.7, EFturnover_0.7, EFnestedness_0.7, distLUI, LUI.x, LUI.y, autotroph.beta.sim, autotroph.beta.sne, herbivore.arthropod.beta.sim, herbivore.arthropod.beta.sne, symbiont.soilfungi.beta.sim, symbiont.soilfungi.beta.sne)]
test[, comparison := paste(Var1, Var2, sep = "_")]
test[, Var1 := NULL]
test[, Var2 := NULL]
test <- test[sample(1:nrow(test), 8), ]
test <- melt(test, id.vars = "comparison")

ggplot(test, aes(x = variable, y = value, colour = comparison, group = comparison)) +
  geom_hline(yintercept = 1) +
  geom_line() +
  geom_point(aes(size = 3)) +
  scale_colour_manual(values = longpallette) +
  labs(title = "specific case descriptions", 
       subtitle = "extraction from dataset \n8 comparisons, selected by highest EFnestedness0.7",
       x = "", y = "distance or absolute value") +
  theme(axis.text.x = element_text(angle = 90))
# saved as : 21-03-11_introductory_plots_specific_case_high_EFnestedness_LUI_selected_beta
```




# trying out
```{r}
ggplot(full_master, aes(x = distLUI, y = EFdistance, colour = autotroph.beta.sne)) +
  geom_point()

a <- ggplot(full_master, aes(x = distLUI, y = autotroph.beta.sne, colour = EFdistance)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
b <- ggplot(full_master, aes(x = distLUI, y = autotroph.beta.sim, colour = EFdistance)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```

moving window of EFdist for LUI autotroph beta connection
```{r}
a <- ggplot(full_master[EFdistance < 0.2], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0, 0.2}")
b <- ggplot(full_master[EFdistance > 0.2 & EFdistance < 0.4], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.2, 0.4}")
c <- ggplot(full_master[EFdistance > 0.4 & EFdistance < 0.6], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.4, 0.6}")
d <- ggplot(full_master[EFdistance > 0.6 & EFdistance < 0.75], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.6, 0.75}")
e <- ggplot(full_master[EFdistance > 0.75], aes(x = distLUI, y = autotroph.beta.sim)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(subtitle = "EFdist in {0.75, 1}")

plot_grid(a, b, c, d, e, NULL, NULL, NULL, NULL, NULL, nrow = 2)
```

