---
title: "plot GDM"
author: "Noelle Schenk"
date: "August 20, 2019"
output: html_document
---
```{r, eval = T}
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(cowplot))
```

```{r}
#TODO :  fill in code to work with significances
#TODO : order of trophic groups :  once all models are read, use effect size sequence and not alphabetically
```


requirements : 
- TODO user : connect to planteco and nsch to get input files
- GDM output as .rds, e.g. gdm_EFnestedness_LUI_input.Rds
    - find at planteco/.../BetaDivMultifun/analysis/output_datasets
- cluster output as .rds, e.g. gdm_EFnestedness_LUI_permutation.Rds
    - TODO : not calculated yet (only for old)
- run `results_nonpublic.R` : load pathtodata, nicenames and model_names
  - pathtodata : path to location where gdm input of model is stored. points to a folder, where the folder cluster witht the model input is. *note : was pathtoout before.*
  - nicenames : helper file to get nice names of functions and trophic levels
  - model_names : helper file for automation. Contains 3 columns : `name`  contains the all model names for `model_name`, `luiinput` contains either "lui" or "components" as LUI input for `luiinput` and `permut` is either TRUE or FALSE.
- for threshold selection : `analysis_nonpublic.R` loads `EFmaster_all_thresholds.Rds`


## Load data

### User input
Either chose automaticrun T or F. If T, the model names and lui input are automatically read in from model_names in a loop, started from `results_nonpublic.R`.
```
for(i in nrow(model_names)){
  source("plot_GDM.R")
}
```

# manual input

If F, `luiinput` and `model_name` need to be given manually.
*note* that i <- 6 for testing.
```{r}
automaticrun <- F
    # only provide if automaticrun is F
luiinput <- "lui" # lui or components
# model_name <- "gdm_EFnestedness_0.5_LUI"
model_name <- "gdm_EFturnover_0.5_LUI"
permut <- F # T or F, depending wether permutations are available or not
```

# preparation
```{r}
# i <- 6 # i <- 6 for testing
if(automaticrun){
  print("automatic input")
  luiinput <- model_names$luiinput[i]
  model_name <- model_names$name[i]
  permut <- model_names$permut[i]
}
if(permut == T){
  gdmperm <- readRDS(paste(pathtoout, "/cluster/gdm_EFnestedness_LUI_permutation.Rds", sep = ""))
  sign <- gdmperm[[3]][,1]
}


###########
# read input data
###########
gdm_output <- readRDS(paste_gdm_input_path_together(pathtoout = pathtodata, name = model_name))
exSplines <- gdm::isplineExtract(gdm_output)

###########
# get model specifications
###########
# get model specs to print with plot
model_specs <- paste(
model_name,
paste("deviance of the fitted GDM model :", round(gdm_output$gdmdeviance, 3)),
paste("deviance of the null model : ", round(gdm_output$nulldeviance, 3)),
paste("percentage of null deviance \n explained by the fitted GDM model :", round(gdm_output$explained, 3)),
"3 splines used, knots at min, max and 50% (default). ",
paste("LUI knots at :", paste(round(gdm_output$knots[1:3], 3), collapse = ", ")),
sep = "\n")
capture.output(print(model_specs), file = paste(pathtoout, "/", model_name, "_GDM_model_specs.txt", sep = ""))
#TODO maybe remove capture output if done in other place

###########
# define plot sequence
###########
plotsequence_bio <- c("autotroph", "bacteria.RNA", "herbivore.arthropodsoillarvae", "secondary.consumer.arthropodsoillarvae", "herbivore.arthropod", "plantpathogen.fungi", "bacterivore.protist", "eukaryvore.protist", "omnivore.protist", "plantparasite.protist", "secondary.consumer", "decomposer.soilfungi", "pathotroph.soilfungi", "symbiont.soilfungi", "tertiary.consumer")
plotsequence_abio <- define_rownames_lui_or_components(luiinput = luiinput)
```



  

# line plots

naming convention : `model_name` + "_lineplot_all_modelspec.pdf"

common y axis maximum is the maximum of the current model
```{r}
###########
# creat plot table
###########
    # one column with xaxis values and one column with yaxis values, and one column
    # containing names
x <- data.table(exSplines$x)
x[, ids := as.numeric(seq(1, nrow(x)))] # ids are required for later matching with y
y <- data.table(exSplines$y)
y[, ids := as.numeric(seq(1, nrow(y)))]
x <- melt(x, measure.vars = colnames(x), value.name = "xaxis", id.vars = "ids", variable.name = "names") # get to long format
y <- melt(y, measure.vars = colnames(y), value.name = "value", id.vars = "ids", variable.name = "names")
    # add x and y values
lineplot_data <- merge(x, y, by = c("ids", "names"))
lineplot_data[, ids := NULL]
lineplot_data <- lineplot_data[names != "ids",]
lineplot_data[, names := as.character(names)] # somehow required for merging
lineplot_data <- merge(lineplot_data, nicenames, by = "names", all.x = T)


###########
# create lineplots
###########
# above, below and abiotic subplots
p <- create_gdm_lineplot(lineplot_data[ground == "a"], legend = T, ymax = max(lineplot_data$value))
q <- create_gdm_lineplot(lineplot_data[ground == "b"], legend = T, ymax = max(lineplot_data$value))

# abiotic is splitted into several plots due to x axes
if(luiinput == "components"){
  luidat <- lineplot_data[names %in% c("Gstd", "Mstd", "Fstd")]
  deltaluidat <- lineplot_data[names %in% c("deltaGstd", "deltaMstd", "deltaFstd")]
} else if(luiinput == "lui"){
  luidat <- lineplot_data[names %in% c("LUI")]
  deltaluidat <- lineplot_data[names %in% c("deltaLUI")]
}
l <- create_gdm_lineplot(luidat, legend = T, ymax = max(lineplot_data$value))
dl <- create_gdm_lineplot(deltaluidat, legend = T, ymax = max(lineplot_data$value))

lgeo <- create_gdm_lineplot(lineplot_data[names %in% c("Geographic")], legend = T, ymax = max(lineplot_data$value))
lsoil <- create_gdm_lineplot(lineplot_data[names %in% c("edis_soil")], legend = T, ymax = max(lineplot_data$value))
lisol <- create_gdm_lineplot(lineplot_data[names %in% c("plot_isolation")], legend = T, ymax = max(lineplot_data$value))

# r <- create_gdm_lineplot(lineplot_data[names %in% c("Geographic", "edis_soil", "plot_isolation")], legend = T)
#TODO outcomment this, because x axis don't match?

###########
# create additional plots
###########
# model specifications and legends
labsplot <- ggplot(NULL) + #annotate("text", x = 0.8, y = 0.5, size=8, label = model_name) +
  theme_void() + annotate("text", x = 0.4, y = 0.5, size = 4, label = model_specs)
biotic_legend <- get_nice_legend(type = "biotic") # plot with plot_grid(biotic_legend)
abiotic_legend <- get_nice_legend(type = "abiotic")



###########
# add plots together
###########
# suppressWarnings(
#   lin <- plot_grid(p, biotic_legend, q, labsplot, l, abiotic_legend,  
#                    labels = c('A', '', 'B', '', 'C', ''), nrow = 3,
#                    rel_widths = c(0.7, 0.3, 0.7, 0.3, 0.7, 0.3))
#   ) # suppresswarnings because of special characters
# 
# suppressWarnings(
# lin <- plot_grid(p, labsplot, q, NULL, l, 
#                  labels = c('A', '', 'B', '', 'C', ''), nrow = 3,
#                  rel_widths = c(0.7, 0.3))
# )
#TODO uncomment above, here just use quick solution for the seminar
suppressWarnings(
  lin <- plot_grid(p, q, l, labels = c("A", "B", "C"), nrow = 3,
                   align = T)
)
                   
lin_abiotic <- plot_grid(dl, lgeo, lsoil, lisol, 
                         labels = c("D", "", "", ""),
                         align = T)

# plot_grid(lin, lin_abiotic, nrow = 1)
#TODO shuffle well the plots on a page. Ev. remove legend

###########
# save
###########
# `model_name` + "_lineplot_all_modelspec.pdf"
cowplot::ggsave(plot = lin, filename = paste(pathtoout, "/", model_name, "_lineplot_no_modelspec.pdf", sep = ""),
                device=cairo_pdf, width = 8.27, height = 11.96, units = "in", dpi = 900)
cowplot::ggsave(plot = lin_abiotic, filename = paste(pathtoout, "/", model_name, "_lineplot_no_modelspec_abiotic.pdf", sep = ""),
                device=cairo_pdf, width = 11.96, height = 8.27, units = "in", dpi = 900)
cowplot::ggsave(plot = labsplot, filename = paste(pathtoout, "/", model_name, "_lineplot_only_modelspec_abiotic.pdf", sep = ""),
                device=cairo_pdf, width = 11.96, height = 8.27, units = "in", dpi = 900)
```






# create sensitivity comparison: deviance explained, sdev
```{r, eval = F}
#############################
#
# treshold selection
#
#############################

# models with function calculated with threshold : 
# thresholds between 0.1 and 0.9.
# need selection : by (new) : sdev of input. old : percent deviance explained.

# VARIABILITY OF RESPONSE VARIABLE
# require : input data : EFmaster_all_thresholds.Rds : load from script "analysis_nonpublic.R"
EFmaster_sd <- apply(EFmaster[, -c("Var1", "Var2")], 2, function(x) sd(x, na.rm = F))
EFmaster_var <- apply(EFmaster[, -c("Var1", "Var2")], 2, function(x) var(x, na.rm = F))
EFmaster <- data.table(name = names(EFmaster), sd = EFmaster_sd, var = EFmaster_var)
EFmaster[, type := "EFdistance"]
EFmaster[grep("nestedness", EFmaster$name), type := "nestedness"]
EFmaster[grep("turnover", EFmaster$name), type := "turnover"]
EFmaster[grep("beta", EFmaster$name), type := "beta"]

ggplot(EFmaster, aes(x = name, y = sd, fill = type)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


# PERCENT DEVIANCE EXPLAINED
# require : 
# model_results : the model results saved manually from the models
setorder(model_results, cols = "name")
model_results[, EF := factor(EF, ordered = T, levels = c("distance", "singleEF", "abund", "beta", "turnover", "nestedness"))]
str(model_results)
model_results <- model_results[!is.na(dev_expl_by_GDM)]

devexpl <- ggplot(model_results, aes(x = name, y = dev_expl_by_GDM, fill = EF)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggsave(filename = paste(pathtoout, "/plot_GDM_EFbeta_tresholds_var_expl.pdf", sep = ""), 
       plot = devexpl, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)

#TODO can add nice names at some point
# STANDARD DEVIATION AND PERCENT DEVIANCE EXPLAINED
model_results[, name := gsub("_LUI", "", model_results$name)]
model_results[, name := gsub("gdm_", "", model_results$name)]
model_results <- merge(model_results, EFmaster, by = "name")

sdplot <- ggplot(model_results, aes(x = name, y = sd, fill = EF)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
threshold_selection <- plot_grid(sdplot, devexpl, nrow = 2)

ggsave(filename = paste(pathtoout, "/plot_GDM_EFbeta_tresholds_var_expl_and_sd.pdf", sep = ""), 
       plot = threshold_selection, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```


  

# Barplots 
```{r}
###########
# create table
###########
maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE) # take maximum of each variable for barplot
restab <- create_restab0()


###########
# produce barplots
###########
p <- create_bio_aboveground_barplot()
b <- create_bio_belowground_barplot()
q <- create_abio_barplot()


###########
# produce overview bars
###########
    # ABOVE - BELOW - ABIOTIC
restab2 <- create_restab2()
restab2 <- restab2[type %in% c("unscaled", "averaged_scaled")]
df <- data.frame(restab2)
ov1 <- create_overview_above_below_abiotic_barplot()
ov1L <- create_overview_above_below_abiotic_barplot(legend = T) # legend only, plot with plot_grid()
    # TURNOVER - NESTEDNESS - ABIOTIC
restab3 <- create_restab_3()
restab3 <- restab3[type %in% c("unscaled", "averaged_scaled")]
df <- data.frame(restab3)
ov2 <- create_overview_turnover_nestedness_abiotic_barplot()
ov2L <- create_overview_turnover_nestedness_abiotic_barplot(legend = T) # only legend, plot with plot_grid()
ov_legends <- plot_grid(ov1L, ov2L, nrow = 2)

###########
# add plots together
###########
# multipanel_without_overviews <- plot_grid(p, b, q, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.31, 0.51, 0.18), align = "v")

# annotated_multipanel_without_overviews <- ggdraw(multipanel_without_overviews) +
  # draw_plot(labsplot, .65, .65, .35, .35) 

#TODO don't delete below, ev. nice for LUI components
# multipanel_with_overview_bars <- plot_grid(p, b, q, ov1, ov2, labels = c('A', '', 'B', 'C', ''), label_size = 12, nrow = 5, rel_heights = c(0.27, 0.37, 0.13, 0.075, 0.075), align = "v") # 2 summary bars

# LUI
multipanel_with_overview_bars <- plot_grid(p, b, q, ov1, ov2, labels = c('A', '', 'B', 'C', ''), label_size = 12, nrow = 5, rel_heights = c(0.23, 0.37, 0.14, 0.1, 0.1), align = "v") # 2 summary bars

annotated_multipanel_with_overview_bars <- ggdraw(multipanel_with_overview_bars) +
  # draw_plot(labsplot, .65, .65, .35, .35) +
  draw_plot(ov_legends, 0.05, 0.1, .1, .2)

extraplot <- plot_grid(labsplot, ov_legends)

###########
# save
###########
    # WITHOUT ANNOTATION
# ggsave(filename=paste(model_name, "_multipanel_barplot.pdf", sep = ""), plot = multipanel_without_overviews, device=cairo_pdf,
       # width = 11.96, height = 8.27, units = "in", dpi = 900)
# ggsave(filename = paste(model_name, "_multipanel_barplot_overviewbars.pdf", sep = ""), plot = multipanel_with_overview_bars, device=cairo_pdf,
       # width = 11.96, height = 8.27, units = "in", dpi = 900)

    # ANNOTATED PLOTS
# ggsave(filename=paste(model_name, "_multipanel_barplot_annotated.pdf", sep = ""), plot = annotated_multipanel_without_overviews, device=cairo_pdf,
       # width = 11.96, height = 8.27, units = "in", dpi = 900)
ggsave(filename = paste(pathtoout, "/", model_name, "_multipanel_barplot_overviewbars_not_annotated.pdf", sep = ""), plot = annotated_multipanel_with_overview_bars, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
ggsave(filename = paste(pathtoout, "/", model_name, "_multipanel_barplot_overviewbars_extraplot_annotations.pdf", sep = ""), plot = extraplot, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)

```



# get all legends
```{r, eval = T}
###########
# legends
###########
biotic_legend <- get_nice_legend(type = "biotic") # plot with plot_grid(biotic_legend)
abiotic_legend <- get_nice_legend(type = "abiotic")

all_legends <- plot_grid(biotic_legend, abiotic_legend, ov1L, ov2L, nrow = 4, rel_heights = c(0.3, 0.3, 0.1, 0.1))
all_legends <- plot_grid(NULL, NULL, NULL, NULL,  all_legends, NULL, NULL, NULL, nrow = 3, rel_heights = c(0.1, 0.8, 0.1))
ggsave(filename = paste(pathtoout, "/", "all_legends_incl_overview.pdf", sep = ""), plot = all_legends, device=cairo_pdf,
       width = 8.27, height = 11.96, units = "in", dpi = 900)
```





# single EFs heatmap

```{r, eval = F}
###########
# load data
###########
singleEFdist <- readRDS(paste(pathtodata, "/data_assembly/output_data/singleEFdist.rds", sep = "")) # already scaled
i <- 1
ni <- colnames(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")][i]
i <- paste(pathtoout, "/GDM/gdm_", ni, "_LUI_input.Rds", sep = "")
gdmin <- readRDS(i)
permut <- "F"

gdm_output <- gdm::gdm(gdmin, geo = F, splines = NULL, knots = NULL)
exSplines <- gdm::isplineExtract(gdm_output)

# take maximum of each variable for barplot
maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
# get significances
if(permut == T){sign <- gdmperm[[3]][,1]}

# create restults table
restab <- create_restab()
data.table::setnames(restab, old = "maxsplines", new = ni)
restab2 <- data.table::copy(restab)

i <- 2
for(i in 2:(length(colnames(singleEFdist))-2)){
  print(i)
  ni <- colnames(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")][i]
  i <- paste(pathtoout, "/GDM/gdm_", ni, "_LUI_input.Rds", sep = "")
  gdmin <- readRDS(i)
  permut <- "F"
  gdm_output <- gdm::gdm(gdmin, geo = F, splines = NULL, knots = NULL)
  exSplines <- gdm::isplineExtract(gdm_output)
  # take maximum of each variable for barplot
  maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
  # get significances
  if(permut == T){sign <- gdmperm[[3]][,1]}
  
  restab <- create_restab()
  restab <- data.table::data.table(restab)
  data.table::setnames(restab, old = "maxsplines", new = ni)
  include <- c(ni, "names")
  restab2 <- merge(restab2, restab[, ..include], by = "names", all.x = T)
}
restab2 <- data.table::data.table(restab2)

```

```{r, eval = F}
# Create heatmap

# pal <- RColorBrewer::brewer.pal(9, "YlOrRd")
pal <- rev(RColorBrewer::brewer.pal(9, "RdYlBu"))
include <- c("nicenames", colnames(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")])

df <- restab2[ground == "a", ]
df <- data.table::melt(df[, ..include], id.vars = "nicenames")

a <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 9, hjust = 1),
    plot.margin = margin(l = 50),
    axis.line.x=element_blank(),
    axis.text.y = element_text(size=9, angle = 0),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.title = element_blank(),
    legend.position = "top") + 
  scale_x_discrete(position = "top") 

df <- restab2[ground == "b", ]
df <- data.table::melt(df[, ..include], id.vars = "nicenames")
b <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


df <- restab2[ground == "x", ]
df <- data.table::melt(df[, ..include], id.vars = "nicenames")
c <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal) +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey"))


# plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.31, 0.51, 0.18), align = "v")
plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.5, 0.4, 0.07), align = "v")
# plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.58, 0.44, 0.10), align = "v") # grayscale with legend

# plot_grid(p, b, q, labels = c('aboveground', 'belowground', 'abiotic'), label_size = 12, nrow = 3, rel_heights = c(0.31, 0.51, 0.18), align = "v")
```
