---
title: "plot GDM"
author: "Noelle Schenk"
date: "August 20, 2019"
output: html_document
---
```{r, eval = T}
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(cowplot))
theme_set(theme_half_open())
```

```{r}
#TODO :  fill in code to work with significances, e.g. take out the non-significant maxsplines
#TODO : order of trophic groups : once all models are read, use effect size sequence and not alphabetically
```

# Aim
This scripts visualises the GDM output. It is runned for each model individually.


**Requirements** : 
- GDM output as .rds, e.g. gdm_EFnestedness_LUI_input.Rds
    - find at planteco/.../BetaDivMultifun/analysis/output_datasets
- cluster output as .rds, e.g. gdm_EFnestedness_LUI_permutation.Rds
    - TODO : not calculated yet (only for old)


## Load data
The model name 

# select input
Depending on the model intented to run, chose the appropriate variables below. The model is selected from the table model_names.

Load table `model_names`.
```{r}
# Load requirements from <..>nonpublic.R file
source("vignettes/analysis_nonpublic.R")
```
Chose model
```{r}
# select a model to run, from the table `model_names` which is part of the R package data (see how to load in `analysis_nonpublic.R`)
if(exists("model_names_selection")){
  # model_names_selection is given from outside the script
  print("using selection of the model from outside the script")
} else {
  print("model selection is not given, default is EFturnover_0.7")
  model_names_selection <- model_names[which(model_names$modelname == "gdm_EFturnover_0.7_LUI"), ]
  print(c("working on model :", paste(model_names_selection, collapse = ", ")))
}
```
Load requirements
```{r}
funs <- model_names_selection$funs
lui <- model_names_selection$lui
sections_to_be_loaded <- c("plotting", "results", "gdminput", "gdmoutput")
source("vignettes/analysis_nonpublic.R")
# will create "model_name" automatically
#    e.g. gdm_EFnestedness_0.7_LUI

permut <- F # T or F, depending wether permutations are available or not
#TODO : implement this
```




# preparation
```{r}
if(permut == T){
  stop("handling permutation values is currently not implemented")
  gdmperm <- readRDS(paste(paste(paste(pathtodata, "/analysis/output_datasets/pvalue_calculation/", sep = ""), model_name, sep = ""), "_perm.Rds", sep = ""))
  sign <- gdmperm[[3]][,1]
  sign <- data.table::data.table("names" = names(sign), "sign" = sign)
}


# read input data
gdm_output <- readRDS(paste_gdm_input_path_together(pathtoout = pathtodata, name = model_name))
exSplines <- gdm::isplineExtract(gdm_output)
maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)

# define plot sequence
plotsequence_bio <- c("autotroph", "bacteria.RNA", "herbivore.arthropodsoillarvae", "secondary.consumer.arthropodsoillarvae", "herbivore.arthropod", "plantpathogen.fungi", "bacterivore.protist", "eukaryvore.protist", "omnivore.protist", "plantparasite.protist", "secondary.consumer", "decomposer.soilfungi", "pathotroph.soilfungi", "symbiont.soilfungi", "tertiary.consumer")
plotsequence_abio <- c(model_names_selection$lui, 
                       paste("delta", lui, sep = ""), "soil", "isolation", "geo")
```

## Creating legend plots
Plots that are re-used in all the plots (Legend and model results)
```{r}
###########
# create additional plots
###########
# model specifications and legends
labsplot <- ggplot(NULL) + #annotate("text", x = 0.8, y = 0.5, size=8, label = model_name) +
  theme_void() + annotate("text", x = 0.4, y = 0.5, size = 3, 
                          label = paste(names(model_specs), model_specs, sep = ": ", collapse = "\n"))
biotic_legend <- get_nice_legend(type = "biotic") # plot with plot_grid(biotic_legend)
abiotic_legend <- get_nice_legend(type = "abiotic")
```



# line plots

naming convention : `model_name` + "_lineplot_all_modelspec.pdf"

common y axis maximum is the maximum of the current model
```{r, eval = T}
###########
# create plot table
###########
    # one column with xaxis values and one column with yaxis values, and one column
    # containing names
x <- data.table(exSplines$x)
x[, ids := as.numeric(seq(1, nrow(x)))] # ids are required for later matching with y
y <- data.table(exSplines$y)
y[, ids := as.numeric(seq(1, nrow(y)))]
x <- melt(x, measure.vars = colnames(x), value.name = "xaxis", id.vars = "ids", variable.name = "names") # get to long format
y <- melt(y, measure.vars = colnames(y), value.name = "value", id.vars = "ids", variable.name = "names")
    # add x and y values
lineplot_data <- merge(x, y, by = c("ids", "names"))
lineplot_data[, ids := NULL]
lineplot_data <- lineplot_data[names != "ids",]
lineplot_data[, names := as.character(names)] # somehow required for merging
lineplot_data <- merge(lineplot_data, nicenames, by = "names", all.x = T)


###########
# create lineplots
###########
# above, below and abiotic subplots
p <- create_gdm_lineplot(lineplot_data[ground == "a"], legend = F, ymax = max(lineplot_data$value))
q <- create_gdm_lineplot(lineplot_data[ground == "b"], legend = F, ymax = max(lineplot_data$value))

# abiotic subplots
# LUI
luidat <- lineplot_data[names %in% model_names_selection$lui,]
deltaluidat <- lineplot_data[names %in% paste("delta", model_names_selection$lui, sep = ""),]
l <- create_gdm_lineplot(luidat, legend = F, ymax = max(lineplot_data$value))
dl <- create_gdm_lineplot(deltaluidat, legend = F, ymax = max(lineplot_data$value))

lgeo <- create_gdm_lineplot(lineplot_data[names %in% c("Geographic")], legend = F, ymax = max(lineplot_data$value))
lsoil <- create_gdm_lineplot(lineplot_data[names %in% c("edis_soil")], legend = F, ymax = max(lineplot_data$value))
lisol <- create_gdm_lineplot(lineplot_data[names %in% c("plot_isolation")], legend = F, ymax = max(lineplot_data$value))


###########
# add plots together
###########
# create classic lineplot
suppressWarnings(
  lin <- plot_grid(p, biotic_legend, q, labsplot, l, abiotic_legend,
                   labels = c('A', '', 'B', '', 'C', ''), nrow = 3,
                   rel_widths = c(0.6, 0.5))
  ) # suppresswarnings because of special characters
lin_abiotic <- plot_grid(dl, lgeo, lsoil, lisol,
                         labels = c("D", "", "", ""),
                         align = T, nrow = 1)
all_lines <- plot_grid(lin, lin_abiotic, nrow = 2, rel_heights = c(0.8, 0.2))


# Alternative plottings
# lin <- plot_grid(p, q, l, labels = c("A", "B", "C"), nrow = 3,
#                    align = T)
# suppressWarnings(
# lin <- plot_grid(p, labsplot, q, NULL, l,
#                  labels = c('A', '', 'B', '', 'C', ''), nrow = 3,
#                  rel_widths = c(0.7, 0.3))
# )
# p_annot <- plot_grid(biotic_legend, plot_grid(abiotic_legend, labsplot, NULL, nrow = 3))
# all_lines <- plot_grid(p, p_annot, q, l, dl, lgeo, lsoil, lisol,
#           labels = c("A", "B", "C", "D"),
#           nrow = 4)



###########
# save
###########
# `model_name` + "_lineplot_all_modelspec.pdf"
# ggplot2::ggsave(plot = lin, 
#                 filename = paste(pathtoout, "/", model_name, "_lineplot.pdf", sep = ""),
#                 device=cairo_pdf, width = 8.27, height = 11.96, units = "in", dpi = 900)

ggplot2::ggsave(plot = all_lines, 
                filename = paste(pathtoout, "/", model_name, "_lineplot.pdf", sep = ""),
                device=cairo_pdf, width = 8.27, height = 11.96, units = "in", dpi = 900)

# ggplot2::ggsave(plot = lin_abiotic, filename = paste(pathtoout, "/", model_name, "_lineplot_no_modelspec_abiotic.pdf", sep = ""),
#                 device=cairo_pdf, width = 11.96, height = 8.27, units = "in", dpi = 900)
# ggplot2::ggsave(plot = labsplot, filename = paste(pathtoout, "/", model_name, "_lineplot_only_modelspec_abiotic.pdf", sep = ""),
#                 device=cairo_pdf, width = 11.96, height = 8.27, units = "in", dpi = 900)
```


# Barplots 
```{r, eval = T}
###########
# create table
###########
# maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE) # take maximum of each variable for barplot # is created above
restab <- create_restab0()


###########
# produce barplots
###########
p <- create_bio_aboveground_barplot(type = "stacked") # type is  in c("stacked", "grouped")
b <- create_bio_belowground_barplot(type = "stacked")
q <- create_abio_barplot(type = "stacked") # here it does not matter if "stacked" or "grouped" is given.


###########
# produce overview bars
###########
# # # # #
# Produce data for plotting
df <- create_overviewbar_restab(restab, fun = "mean")
df2 <- create_overviewbar_restab(restab, fun = "sum")
df <- rbindlist(list(df, df2), use.names = T, fill = T)
rm(df2)
saveRDS(df, file = paste(pathtoout, "/", model_names_selection$modelname, "_overviewbar_data.RDS", sep = ""))
#TODO check if it's a problem that instead of "averaged_scaled", it's called "mean"

# # # # #
# Produce plot
ov <- create_overview_barplot(df)
ovL <- create_overview_barplot(df, legend = T)
# plot_grid(ovL) # to show the legend individually


###########
# add plots together
###########
multipanel_with_overview_bars <- plot_grid(p, b, q, ov, 
                                           labels = c('A', '', 'B', 'C'), 
                                             label_size = 12, nrow = 4,
                                             rel_heights = c(0.21, 0.32, 0.2, 0.15), 
                                           align = "v") # 2 summary bars


annotated_multipanel_with_overview_bars <- ggdraw(multipanel_with_overview_bars) +
  # draw_plot(labsplot, .65, .65, .35, .35) +
  draw_plot(ovL, 0.04, 0.02, .1, .2) + # (left, below, x, x)
  draw_plot(labsplot, 0.6, 0.6, 0.6, 0.6)

# extraplot <- plot_grid(labsplot, ov_legends)

###########
# save
###########
# -- probably delete below
#     # WITHOUT ANNOTATION
# # ggsave(filename=paste(model_name, "_multipanel_barplot.pdf", sep = ""), plot = multipanel_without_overviews, device=cairo_pdf,
#        # width = 11.96, height = 8.27, units = "in", dpi = 900)
# # ggsave(filename = paste(model_name, "_multipanel_barplot_overviewbars.pdf", sep = ""), plot = multipanel_with_overview_bars, device=cairo_pdf,
#        # width = 11.96, height = 8.27, units = "in", dpi = 900)
# 
#     # ANNOTATED PLOTS
# # ggsave(filename=paste(model_name, "_multipanel_barplot_annotated.pdf", sep = ""), plot = annotated_multipanel_without_overviews, device=cairo_pdf,
#        # width = 11.96, height = 8.27, units = "in", dpi = 900)
# --

ggsave(filename = paste(pathtoout, "/", model_name, "_barplot.pdf", sep = ""), plot = annotated_multipanel_with_overview_bars, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)

# ggsave(filename = paste(pathtoout, "/", model_name, "_multipanel_barplot_overviewbars_extraplot_annotations.pdf", sep = ""), plot = extraplot, device=cairo_pdf,
#        width = 11.96, height = 8.27, units = "in", dpi = 900)
```
Note : the barplot of weighted average is done in the script `GDM_multifun_thresholds.Rmd`, together with pieces of code from this script.
