---
title: "Plot unique GDM"
author: "N. Schenk"
date: "2022-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```







#TODO HERE
the below part is mostly appendix figures. 
Probably move that?
Also here : only include figures that are done model-wise and that have an input data format which matches the
examples here.
Check if the output format of the average over threshold models can be used as input here.
Check if the ouptut format of singleEFdist models can be used as input here (probably could be, but we anyway show results in a summary) Maybe just use summary functions used here for those as well.
--> Philosophy : use more scripts, that contain similar things.

- Move the creation of the average-threshold computation to a separate file and put that to the group of scripts that only need to be runned once.









# Once-in lifetime plots (only run once)

## get all legends
```{r, eval = F}
###########
# legends
###########
biotic_legend <- get_nice_legend(type = "biotic") # plot with plot_grid(biotic_legend)
abiotic_legend <- get_nice_legend(type = "abiotic")

all_legends <- plot_grid(biotic_legend, abiotic_legend, ovL, nrow = 4, rel_heights = c(0.3, 0.3, 0.1, 0.1))
all_legends <- plot_grid(NULL, NULL, NULL, NULL,  all_legends, NULL, NULL, NULL, nrow = 3, rel_heights = c(0.1, 0.8, 0.1))
ggsave(filename = paste(pathtoout, "/", "all_legends_incl_overview.pdf", sep = ""), plot = all_legends, device=cairo_pdf,
       width = 8.27, height = 11.96, units = "in", dpi = 900)
```



## single EFs heatmap

```{r, eval = F}
###########
# load data
###########
singleEFdist <- readRDS(paste(pathtodata, "/data_assembly/output_data/singleEFdist.rds", sep = "")) # already scaled
i <- 1
ni <- colnames(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")][i]
i <- paste(pathtodata, "/analysis/output_datasets/gdm_", ni, "_LUI_input.Rds", sep = "")
gdmin <- readRDS(i)
permut <- "F"

gdm_output <- gdm::gdm(gdmin, geo = F, splines = NULL, knots = NULL)
exSplines <- gdm::isplineExtract(gdm_output)

# take maximum of each variable for barplot
maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
# get significances
if(permut == T){sign <- gdmperm[[3]][,1]}

# create restults table
restab <- create_restab()
data.table::setnames(restab, old = "maxsplines", new = ni)
restab2 <- data.table::copy(restab)

i <- 2
for(i in 2:(length(colnames(singleEFdist))-2)){
  print(i)
  ni <- colnames(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")][i]
  i <- paste(pathtodata, "/analysis/output_datasets/gdm_", ni, "_LUI_input.Rds", sep = "")
  gdmin <- readRDS(i)
  permut <- "F"
  gdm_output <- gdm::gdm(gdmin, geo = F, splines = NULL, knots = NULL)
  exSplines <- gdm::isplineExtract(gdm_output)
  # take maximum of each variable for barplot
  maxsplines <- apply(exSplines$y, 2, max, na.rm = TRUE)
  # get significances
  if(permut == T){sign <- gdmperm[[3]][,1]}
  
  restab <- create_restab()
  restab <- data.table::data.table(restab)
  data.table::setnames(restab, old = "maxsplines", new = ni)
  include <- c(ni, "names")
  restab2 <- merge(restab2, restab[, ..include], by = "names", all.x = T)
}
restab2 <- data.table::data.table(restab2)

# report the number of functions affected by each group / affected by above- belowground / abiotic
singleEFnames <- names(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")]
x <- apply(restab2[, ..singleEFnames], 1, function(x) sum(x > 0))
restab2[, n_affected_fun := x]
rm(x)
```

Create heatmap
```{r, eval = F}
# Create heatmap
include <- c("nicenames", colnames(singleEFdist)[!colnames(singleEFdist) %in% c("Var1", "Var2")], "n_affected_fun")

# pal <- rev(RColorBrewer::brewer.pal(9, "RdYlBu"))
pal <- RColorBrewer::brewer.pal(9, "YlOrRd")
# pal <- c("white", RColorBrewer::brewer.pal(9, "YlOrRd"))[-2]
# colgrad <- melt(restab2[, ..include], id.vars = c("nicenames", "n_affected_fun"))
# colgrad <- colgrad$value  

df <- restab2[ground == "a", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames", "n_affected_fun"))
df[value == 0, value := NA] # 0 values are set to NA to make them appear grey in the plot

a <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 9, hjust = 1),
    plot.margin = margin(l = 50, r = 20),
    axis.line.x=element_blank(),
    axis.text.y = element_text(size=9, angle = 0),
    axis.ticks.x = element_blank(),
    panel.grid.major.x = element_line(color = "grey"),
    axis.title = element_blank(),
    legend.position = "top") +
  scale_x_discrete(position = "top") +
  coord_cartesian(clip = 'off') +  
  geom_text(data = unique(df[, .(nicenames, n_affected_fun)]), aes(label = n_affected_fun, x = length(unique(df$variable)) + 1, y = nicenames), color="black", size=3, inherit.aes = FALSE)

df <- restab2[ground == "b", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames", "n_affected_fun"))
df[value == 0, value := NA]
b <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey")) +
    coord_cartesian(clip = 'off') +  
  geom_text(data = unique(df[, .(nicenames, n_affected_fun)]), aes(label = n_affected_fun, x = length(unique(df$variable)) + 1, y = nicenames), color="black", size=3, inherit.aes = FALSE)


df <- restab2[ground == "x", ]
df <- data.table::melt(df[, ..include], id.vars = c("nicenames", "n_affected_fun"))
df[value == 0, value := NA]
c <- ggplot(data = df, aes(x = variable, y = nicenames, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = pal, na.value = "grey80") +
  theme(legend.position = "none", 
        axis.title = element_blank(),
        axis.text.y = element_text(size=9, angle = 0),
        plot.margin = margin(l = 50, r = 20),
        axis.text.x = element_blank(),
        axis.line.x=element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_line(color = "grey")) +
      coord_cartesian(clip = 'off') +  
  geom_text(data = unique(df[, .(nicenames, n_affected_fun)]), aes(label = n_affected_fun, x = length(unique(df$variable)) + 1, y = nicenames), color="black", size=3, inherit.aes = FALSE)

# plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.31, 0.51, 0.18), align = "v")
heatmap <- plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.49, 0.42, 0.09), align = "v")
# plot_grid(a, b, c, labels = c('A', '', 'B'), label_size = 12, nrow = 3, rel_heights = c(0.58, 0.44, 0.10), align = "v") # grayscale with legend

# plot_grid(p, b, q, labels = c('aboveground', 'belowground', 'abiotic'), label_size = 12, nrow = 3, rel_heights = c(0.31, 0.51, 0.18), align = "v")

ggsave(filename = paste(pathtoout, "/plot_GDM_singleEFdist_heatmap.pdf", sep = ""), plot = heatmap, device=cairo_pdf,
       width = 11.96, height = 8.27, units = "in", dpi = 900)
```

Create overview bars
```{r}
# requires : restab2, created in chunk 1 under "# single EFs heatmap"
a <- create_single_funs_overviewbars(restab2 = restab2, rel_colnames = singleEFnames)

# bring to ggplot format
ov_ab_singleEFmods <- melt(a$above_below, id.vars = c("ground", "color"))
ov_tn_singleEFmods <- melt(a$turnover_nestedess, id.vars = c("component", "color"))
# create plots
p_ov_ab <- create_single_funs_overviewbar_plot(singleF_restab = ov_ab_singleEFmods, pos = "dodge")
p_ov_tn <- create_single_funs_overviewbar_plot(singleF_restab = ov_tn_singleEFmods, pos = "dodge")

p_ov_abtn <- plot_grid(p_ov_ab, p_ov_tn, labels = c("A", "B"))
# saved as : <date>_plot_GDM_singleEFdist_overview_bars
```

Summary numbers about number of average biotic/ abiotic/ LUI drivers driving
single functions
```{r}
# use vector "singleEFnames" to get names of single functions to address each column separately

# add column to restab2 with needed categories
restab2[, bio_abio_lui := lui_ground]
restab2[bio_abio_lui == "a", bio_abio_lui := "b"]

calc_avg_number_of_biotic_abiotic_drivers_of_single_functions(restab2 = restab2,
                                                              singleEFnames = singleEFnames)
# total available categories
table(restab2$bio_abio_lui)
```

