---
title: "prepare and run GDM"
author: "Noelle Schenk"
date: "1/24/2020"
output: html_document
---
This script runs the GDM models. In the first code section, the desired GDM model is chosen manually. The script writes the GDM input and ouput files and the summary output.

# Dependencies
## Data
- colorder_gdm_input : the order of columns which is used for GDM input data. Can be given manually, but is very tedious
- EFmaster : ecosystem function dissimilarities
- masterbeta
- LUI
- predictors, predictors_soil
- pathtoout
see `analysis_nonpublic.R` file.

## packages
- BetaDivMultifun (as usual)
    - the pacakges used by betadivmultifun (see DESCRIPTION file)

# output
- gdm input .CSV and .Rds
- gdm output as .Rds
- gdm summary output as .txt

# select input
depending on the model intented to run, chose the appropriate variables below.
```{r}
#TODO : write automatic input for all models
#TODO : capture deviance explained from all models and store in table
# ECOSYSTEM FUNCTION
funs <- "EFdistance"
# funs <- "soilNitrateflxs"
# funs <- "EFbeta_median"

# LUI
## name of the chosen LUI
compon_lui <- "LUI"
# compon_lui <- "components"

## column titles of the chosen LUI
lui <- c("LUI", "deltaLUI")
# lui <- c("Gstd", "Mstd", "Fstd", "deltaGstd", "deltaMstd", "deltaFstd")

## special case : ABUNDANCE-based betadiversity
# model name : gdm_EFdistance_abundancebeta_LUI
colorder_gdm_input <- gsub(pattern = ".beta.sim", replacement = ".beta.bal", colorder_gdm_input)
colorder_gdm_input <- gsub(pattern = ".beta.sne", replacement = ".beta.gra", colorder_gdm_input)
```
chose selected input from mastertables
```{r}
modelname <- paste("gdm", funs, compon_lui, sep = "_")

include <- c("Var1", "Var2", funs)
EFmaster <- EFmaster[, ..include]
rm(include)

LUI <- LUI[, c("Plotn", lui)]
LUI <- data.table(LUI)

# create table containing geographic distances
geodist <- predictors[, c("Plotn", "Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")]
geodist <- data.table(geodist)
predictors <- data.table(predictors)
predictors <- predictors[, .(Plotn, plot_isolation)]
predictors_soil <- data.table(predictors_soil)
```


# prepare GDM input table

## ecosystem functions
Distance in ecosystem functions is called "distance"
```{r}
setnames(EFmaster, old = funs, new = "distance")
gdminput <- data.table::copy(EFmaster)
gdminput[, weights := 1]
rm(EFmaster)
```

## geographic distance
xcoord = Latitude
ycoord = Longitude
```{r}
setcolorder(geodist, c("Plotn", "Latitude_Dec_Plotcenter", "Longitude_Dec_Plotcenter"))
setnames(geodist, old = c("Plotn", "Latitude_Dec_Plotcenter","Longitude_Dec_Plotcenter"), new = c("Var1", "s1.xCoord", "s1.yCoord"))
gdminput <- merge(gdminput, geodist, by = c("Var1"))
setnames(geodist, old = c("Var1", "s1.xCoord", "s1.yCoord"), new = c("Var2", "s2.xCoord", "s2.yCoord"))
gdminput <- merge(gdminput, geodist, by = "Var2")
```

### variant : by hand geographic distance calculation
Geographic distance can also be calculated with geosphere package and normally added to gdminput, when geo = F in gdm.
```{r, eval = F}
gdminput[, s1.geosphere_distgeo := geosphere::distGeo(p1 = as.matrix(gdminput[, .(s1.yCoord, s1.xCoord)]), 
                                                   p2 = as.matrix(gdminput[, .(s2.yCoord, s2.xCoord)]))]
gdminput[, s2.geosphere_distgeo := 0]
```


## betadiversities
Prepare for each betadiversity two columns : one "s1.<name_of_diversity>" containing the betadiversity values and one "s2.<name_of_diversity>" with 0.
GDM expects two values per predictor, but as dissimilarities for betadivs are already calculated, the input is just the distance between the actual distance (beta diversity) and 0.
```{r}
# add betadiversities
colnames(masterbeta)[-c(1,2)] <- paste("s1.", colnames(masterbeta)[-c(1,2)], sep = "")
gdminput <- merge(gdminput, masterbeta, by = c("Var1", "Var2"))
# add 0 columns
zeronames <- gsub("s1.", "s2.", colnames(masterbeta)[-c(1,2)])
gdminput[, (zeronames) := 0]
```

## LUI
```{r}
setnames(LUI, old = c("Plotn", colnames(LUI)[-1]), new = c("Var1", paste("s1.", colnames(LUI)[-1], sep = "")))
gdminput <- merge(gdminput, LUI, by = "Var1")
setnames(LUI, old = c("Var1", colnames(LUI)[-1]), new = c("Var2", gsub("s1.", "s2.", colnames(LUI)[-1])))
gdminput <- merge(gdminput, LUI, by = "Var2")
```


## covariates
```{r}
# create predictors table with edis_soil
setnames(predictors_soil, old = "edis_soil", new = "s1.edis_soil")
predictors_soil[, s2.edis_soil := 0]

# add plot isolation
setnames(predictors, old = c("Plotn" ,"plot_isolation"), new = c("Var1" ,"s1.plot_isolation"))
predictors_soil <- merge(predictors_soil, predictors, by = "Var1")
setnames(predictors, old = c("Var1" ,"s1.plot_isolation"), new = c("Var2" ,"s2.plot_isolation"))
predictors_soil <- merge(predictors_soil, predictors, by = "Var2")
rm(predictors)
gdminput <- merge(gdminput, predictors_soil, by = c("Var1", "Var2"))
```


# finalising dataset for GDM
fix column order as required by gdm. The required columns, followed by s1.betadiversities, s1.LUI, s1.predictors, s2.betadiversities, s2.LUI, s2.predictors.
convert to data.frame and assign rownames to be the plot combinations separated by dot.

GDM requires a well-defined table. Below, the names of the columns and their content are listed:
- distance : distance between ecosystem functions
- weights : column containing all 1 (because we weigh all comparisons the same)
- s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord : 4 columns which will be ignored as I will set geo = F in the gdm() function. 
- s1.pred1, ..., s1.predn : here I fill in my predictor columns, (all have s1. at the beginning)
- s2.pred1, ..., s2.predn : here I fill in columns of zero to imitate I would not have given distance but two values.
```{r, eval = F}
### variant : by hand geographic distance calculation
# adjust colorder so it includes s1.geosphere_distgeo and s2.geosphere_distgeo
colorder_gdm_input <- append(colorder_gdm_input, "s1.geosphere_distgeo", after=max(grep("s1.", colorder_gdm_input)))
colorder_gdm_input <- append(colorder_gdm_input, "s2.geosphere_distgeo", after = max(grep("s2.", colorder_gdm_input)))
```

```{r}
# select respective LUI measures from colorder
grep(lui[1], colorder_gdm_input)
if(lui[1] == "LUI"){
  colorder_gdm_input <- colorder_gdm_input[-grep(paste(c("Gstd", "Mstd", "Fstd"), collapse = "|"), colorder_gdm_input)]
} else {
  colorder_gdm_input <- colorder_gdm_input[-grep("LUI", colorder_gdm_input)]
}
```


```{r}
# # --------- intermediate solution start
# #TODO change input file, fix this
# # adapt for LUi
# test <- read.csv("colord.csv", header = T)
# colorder_gdm_input <- as.vector(t(test))
# rm(test)
# #TODO maybe add tertiary consumers
# colorder_gdm_input[36]
# # grep("tertiary", colnames(gdminput), value = T)
# colorder_gdm_input <- append(colorder_gdm_input, c("s1.tertiary.consumer.beta.sne"), after = 36)
# colorder_gdm_input[22]
# colorder_gdm_input <- append(colorder_gdm_input, c("s1.tertiary.consumer.beta.sim"), after = 22)
# colorder_gdm_input[56]
# colorder_gdm_input <- append(colorder_gdm_input, "s2.tertiary.consumer.beta.sim", after = 56)
# colorder_gdm_input[71]
# colorder_gdm_input <- append(colorder_gdm_input, "s2.tertiary.consumer.beta.sne", after = 71)
# # --------- intermediate solution end

# check if all names of gdminput are in colorder_gdm_input and vice versa
all(colnames(gdminput) %in% colorder_gdm_input)
all(colorder_gdm_input %in% colnames(gdminput))
length(grep("geosphere_distgeo", colorder_gdm_input)) == 0
length(grep("geosphere_distgeo", colnames(gdminput))) == 0


gdminput <- data.table::setcolorder(gdminput, colorder_gdm_input)

gdminput[, rown := paste(Var1, Var2, sep = ".")]
gdminput[,c("Var1","Var2") := NULL]
gdminput <- as.data.frame(gdminput)
rownames(gdminput) <- gdminput$rown
gdminput$rown <- NULL

gdminput <- gdm::formatsitepair(bioData = gdminput, bioFormat = 4, predData = gdminput)
```




# run GDM
```{r}
gdm_output <- gdm::gdm(gdminput, geo = T)
```

# save output
```{r}
outname <- paste("gdm", funs, compon_lui, "output_summary.txt", sep = "_")
outnameRDS <- paste("gdm", funs, compon_lui, "output.Rds", sep = "_")

# save input and output
saveRDS(gdminput, file = paste("vignettes/out/", paste("gdm", funs, compon_lui, "input.Rds", sep = "_"), sep = ""))
write.csv(gdminput, file = paste("vignettes/out/", paste("gdm", funs, compon_lui, "input.CSV", sep = "_"), sep = ""))
saveRDS(gdm_output, paste("vignettes/out/", outnameRDS, sep = ""))
capture.output(summary(gdm_output), file = paste("vignettes/out/", outname, sep = ""))

# get model specifications
model_specs <- paste(
modelname,
paste("deviance of the fitted GDM model :", round(gdm_output$gdmdeviance, 3)),
paste("deviance of the null model : ", round(gdm_output$nulldeviance, 3)),
paste("percentage of null deviance \n explained by the fitted GDM model :", round(gdm_output$explained, 3)),
"3 splines used, knots at min, max and 50% (default). ",
paste("LUI knots at :", paste(round(gdm_output$knots[1:3], 3), collapse = ", ")),
sep = "\n")
capture.output(print(model_specs), file = paste(pathtoout, "/", modelname, "_GDM_model_specs.txt", sep = ""))
```


# next steps
- check the input table for correlations and produce linear model and variance inflation `check_GDM_input.Rmd`
- plot models
- run permutations (cluster scripts)