---
title: "prepare and run GDM"
author: "Noelle Schenk"
date: "1/24/2020"
output: html_document
---
This script runs the GDM models. In the first code section, the desired GDM model is chosen manually. The script writes the GDM input and output files and the summary output.

# Dependencies
## Data
- colorder_gdm_input : the order of columns which is used for GDM input data. Can be given manually, but is very tedious
- datasets : 
  - EFmaster
  - masterbeta
  - LUI
  - predictors, predictors_soil
- pathtoout : path to save output to
see `analysis_nonpublic.R` file.

```{r}
sections_to_be_loaded <- c("functions_dissimilarity", "betadiversity", "LUI", "prepared_covariates")
source("vignettes/analysis_nonpublic.R")
```

# output
- gdm input .CSV and .Rds (the dataset fed into GDM models)
- gdm output as .Rds (the direct output of a GDM model)
- gdm summary output as .txt (the model summary as given by the `gdm()` funciton)


# select input
Depending on the model intented to run, chose the appropriate variables below.
```{r}
# select a model to run, from the table `model_names` which is part of the R package data (see how to load in `analysis_nonpublic.R`)
if(exists("model_names_selection")){
  # model_names_selection is given from outside the script
  print("using selection of the model from outside the script")
} else {
  print("model selection is not given, default is EFturnover_0.7")
  model_names_selection <- model_names[which(model_names$modelname == "gdm_EFturnover_0.7_LUI"), ]
  print(c("working on model :", paste(model_names_selection, collapse = ", ")))
}
```

# Implementation of input choice
chose selected input from mastertables
```{r}
## column titles of the chosen LUI
lui <- c(model_names_selection$lui, paste("delta", model_names_selection$lui, sep = ""))
# can be e.g. : "deltaLUI", "Gstd", "Mstd", "Fstd", "deltaGstd", "deltaMstd", "deltaFstd"

# ecosystem funcitons
include <- c("Var1", "Var2", model_names_selection$funs)
EFmaster <- EFmaster[, ..include]
rm(include)

# lui
LUI <- LUI[, c("Plotn", lui)]
LUI <- data.table(LUI)

# geographic distances
geodist <- predictors[, c("Plotn", "Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")]
geodist <- data.table(geodist)

# soil covariates
predictors <- data.table(predictors)
predictors <- predictors[, .(Plotn, plot_isolation)]
predictors_soil <- data.table(predictors_soil)
```


# prepare GDM input table

## ecosystem functions
Distance in ecosystem functions is called "distance"
```{r}
setnames(EFmaster, old = model_names_selection$funs, new = "distance")
gdminput <- data.table::copy(EFmaster)
gdminput[, weights := 1]
rm(EFmaster)
```

## geographic distance
xcoord = Latitude
ycoord = Longitude
```{r}
setcolorder(geodist, c("Plotn", "Latitude_Dec_Plotcenter", "Longitude_Dec_Plotcenter"))
setnames(geodist, old = c("Plotn", "Latitude_Dec_Plotcenter","Longitude_Dec_Plotcenter"), new = c("Var1", "s1.xCoord", "s1.yCoord"))
gdminput <- merge(gdminput, geodist, by = c("Var1"))
setnames(geodist, old = c("Var1", "s1.xCoord", "s1.yCoord"), new = c("Var2", "s2.xCoord", "s2.yCoord"))
gdminput <- merge(gdminput, geodist, by = "Var2")
```

### variant : by hand geographic distance calculation
Geographic distance can also be calculated with geosphere package and normally added to gdminput, when geo = F in gdm.
```{r, eval = F}
gdminput[, s1.geosphere_distgeo := geosphere::distGeo(p1 = as.matrix(gdminput[, .(s1.yCoord, s1.xCoord)]), 
                                                   p2 = as.matrix(gdminput[, .(s2.yCoord, s2.xCoord)]))]
gdminput[, s2.geosphere_distgeo := 0]
```
Note : based on sensitivity analysis (comparing by-hand and automatic distance calculation), this variant was not used.
#TODO include conclusion why.
Note that this variant has code below as well.

## betadiversities
Prepare for each betadiversity two columns : one "s1.<name_of_diversity>" containing the betadiversity values and one "s2.<name_of_diversity>" with 0.
GDM expects two values per predictor, but as dissimilarities for betadivs are already calculated, the input is just the distance between the actual distance (beta diversity) and 0.
Note that this is not the intented use of GDM and causes issues in permutation that need to be solved later on.
```{r}
# add betadiversities
colnames(masterbeta)[-c(1,2)] <- paste("s1.", colnames(masterbeta)[-c(1,2)], sep = "")
gdminput <- merge(gdminput, masterbeta, by = c("Var1", "Var2"))
# add 0 columns
zeronames <- gsub("s1.", "s2.", colnames(masterbeta)[-c(1,2)])
gdminput[, (zeronames) := 0]
```

## LUI
```{r}
setnames(LUI, old = c("Plotn", colnames(LUI)[-1]), new = c("Var1", paste("s1.", colnames(LUI)[-1], sep = "")))
gdminput <- merge(gdminput, LUI, by = "Var1")
setnames(LUI, old = c("Var1", colnames(LUI)[-1]), new = c("Var2", gsub("s1.", "s2.", colnames(LUI)[-1])))
gdminput <- merge(gdminput, LUI, by = "Var2")
```


## covariates
```{r}
# create predictors table with edis_soil
setnames(predictors_soil, old = "edis_soil", new = "s1.edis_soil")
predictors_soil[, s2.edis_soil := 0]

# add plot isolation
setnames(predictors, old = c("Plotn" ,"plot_isolation"), new = c("Var1" ,"s1.plot_isolation"))
predictors_soil <- merge(predictors_soil, predictors, by = "Var1")
setnames(predictors, old = c("Var1" ,"s1.plot_isolation"), new = c("Var2" ,"s2.plot_isolation"))
predictors_soil <- merge(predictors_soil, predictors, by = "Var2")
rm(predictors)
gdminput <- merge(gdminput, predictors_soil, by = c("Var1", "Var2"))
```


# finalising dataset for GDM
fix column order as required by gdm. The required columns, followed by s1.betadiversities, s1.LUI, s1.predictors, s2.betadiversities, s2.LUI, s2.predictors.
convert to data.frame and assign rownames to be the plot combinations separated by dot.

GDM requires a well-defined table. Below, the names of the columns and their content are listed:
- distance : distance between ecosystem functions
- weights : column containing all 1 (because we weigh all comparisons the same)
- s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord : 4 columns containting geographic location. GDM either includes them in modelling (geo = T) or excludes (geo = F).
- s1.pred1, ..., s1.predn : here I fill in my predictor columns, diversity, LUI and covariates of the first plot. (all have s1. at the beginning)
- s2.pred1, ..., s2.predn : here I fill in LUI and covariate values of plot 2, and columns of zero for betadiversity to imitate I would have 2 values.
```{r, eval = F}
### variant : by hand geographic distance calculation
# adjust colorder so it includes s1.geosphere_distgeo and s2.geosphere_distgeo
colorder_gdm_input <- append(colorder_gdm_input, "s1.geosphere_distgeo", after=max(grep("s1.", colorder_gdm_input)))
colorder_gdm_input <- append(colorder_gdm_input, "s2.geosphere_distgeo", after = max(grep("s2.", colorder_gdm_input)))
```

```{r}
# select respective LUI measures from colorder
# grep(lui[1], colorder_gdm_input)
if(lui[1] == "LUI"){
  colorder_gdm_input <- colorder_gdm_input[-grep(paste(c("Gstd", "Mstd", "Fstd"), collapse = "|"), colorder_gdm_input)]
} else {
  colorder_gdm_input <- colorder_gdm_input[-grep("LUI", colorder_gdm_input)]
}
```


```{r}
# check if all names of gdminput are in colorder_gdm_input and vice versa
all(colnames(gdminput) %in% colorder_gdm_input)
all(colorder_gdm_input %in% colnames(gdminput))
length(grep("geosphere_distgeo", colorder_gdm_input)) == 0
length(grep("geosphere_distgeo", colnames(gdminput))) == 0

gdminput <- data.table::setcolorder(gdminput, colorder_gdm_input)

gdminput[, rown := paste(Var1, Var2, sep = ".")]
gdminput[,c("Var1","Var2") := NULL]
gdminput <- as.data.frame(gdminput)
rownames(gdminput) <- gdminput$rown
gdminput$rown <- NULL

gdminput <- gdm::formatsitepair(bioData = gdminput, bioFormat = 4, predData = gdminput)
```


# run GDM
```{r}
gdm_output <- gdm::gdm(gdminput, geo = T)
```


# save output
```{r}
outname <- paste("gdm", model_names_selection$funs, lui[1], "output_summary.txt", sep = "_")
outnameRDS <- paste("gdm", model_names_selection$funs, lui[1], "output.Rds", sep = "_")

# save input and output
saveRDS(gdminput, file = paste("vignettes/out/", paste("gdm", model_names_selection$funs, lui[1], "input.Rds", sep = "_"), sep = ""))
write.csv(gdminput, file = paste("vignettes/out/", paste("gdm", model_names_selection$funs, lui[1], "input.CSV", sep = "_"), sep = ""))
saveRDS(gdm_output, paste("vignettes/out/", outnameRDS, sep = ""))
capture.output(summary(gdm_output), file = paste("vignettes/out/", outname, sep = ""))

# get model specifications
# model_specs <- paste(
# model_names_selection$modelname,
# paste("deviance of the fitted GDM model :", round(gdm_output$gdmdeviance, 3)),
# paste("deviance of the null model : ", round(gdm_output$nulldeviance, 3)),
# paste("percentage of null deviance \n explained by the fitted GDM model :", round(gdm_output$explained, 3)),
# "3 splines used, knots at min, max and 50% (default). ",
# paste("LUI knots at :", paste(round(gdm_output$knots[1:3], 3), collapse = ", ")),
# sep = "\n")
# capture.output(print(model_specs), file = paste(pathtoout, "/", model_names_selection$modelname, "_GDM_model_specs.txt", sep = ""))

model_specs <- data.table(modelname = model_names_selection$modelname,
                           "deviance of the fitted GDM model" = round(gdm_output$gdmdeviance, 3),
                           "deviance of the null model" = round(gdm_output$nulldeviance, 3),
                           "percentage of null dev expl" = round(gdm_output$explained, 3),
                           "3 splines knots at min, max and 50% (default)" = paste(round(gdm_output$knots[1:3], 3), collapse = ", ")
)
fwrite(model_specs, paste(pathtoout, "/", model_names_selection$modelname, "_GDM_model_spects.csv", sep = ""))
```


# next steps
- check the input table for correlations and produce linear model and variance inflation `check_GDM_input.Rmd`
- later : plot models, run permutations (cluster scripts)
