---
title: "run GDM"
author: "Noelle Schenk"
date: "June 21, 2019"
output: html_document
---

note : packages used : gdm, car, geosphere, viridis and colrobrewer

# Dependencies
This script relies on loading the following datasets:
- EFmaster : dissimilarities of functions
- singleEFdist : dissimilarities of single functions
TODO : is singleEFdist really needed?
- masterbeta : dissimilarities of trophic groups
- LUI : LUI (and components) and delta LUI (and delta components)
- predictors : values of predictors (plot isolation, geographic position)
- predictors soil : euclidean distance of soil pca

# outputs COLLECT THEM BY HAND AT THE MOMENT!
- correlation plot
- variable inflations
- gdm models


# Select Variables
Select Variables for the models
- Ecosystem Functions : either EFdistance, EFturnover, EFnestedness or single functions
- LUI : either LUI and delta LUI or the individual components
- xaxis_gdm : is LUI or a component
```{r}
# print("user input required here!")
source('/run/user/1001/gvfs/smb-share:server=ipssmonstera.unibe.ch,share=planteco/DATA/BE_Synthesis/BetaDivMultifun/analysis_nonpublic.R', echo=F)

##  PSEUDO-AUTOMATIC RUNNING
# specify what is required here
# can be run from `analysis_nonpublic.R`:

## MODELS
funs <- "EFdistance"
# funs <- "EFabund_bray_gra"

## TURNOVER AND NESTEDNESS
# with several thresholds
# funs <- paste("EFturnover_", c(seq(0.1, 0.9, 0.1), "median"), sep = "")[1] # from 1 to 10
# funs <- paste("EFnestedness_", c(seq(0.1, 0.9, 0.1), "median"), sep = "")[1] # from 1 to 9

## SINGLE FUNCTIONS
# funs <- colnames(EFmaster)[11] # from 3 to 19

# LUI is one of the below
# lui <- c("LUI", "deltaLUI")
lui <- c("Gstd", "Mstd", "Fstd", "deltaGstd", "deltaMstd", "deltaFstd")

# xaxis_gdm <- "LUI"
xaxis_gdm <- "Fstd"
```

# Prepare for GDM
```{r}
lui <- lui[!lui %in% xaxis_gdm]
modelname <- paste("gdm", funs, xaxis_gdm, sep = "_")

include <- c("Var1", "Var2", funs)
EFmaster <- EFmaster[, ..include]

xLUI <- LUI[, c("Plotn", xaxis_gdm)]
LUI <- LUI[, c("Plotn", lui)]

# Group dissimilarity matrices (long format):
# combine functions, betadiversity and predictors
fordistpreds <- merge(predictors_soil, masterbeta, by = c("Var1", "Var2"))
fordistpreds <- merge(fordistpreds, EFmaster, by = c("Var1", "Var2"))
rm(predictors_soil, masterbeta); gc()
```

Calc dissimilarities from covariates which are not dissimilarities yet
```{r}
forpredData <- merge(predictors, LUI, by = "Plotn")
rownames(forpredData) <- forpredData$Plotn
forpredData$Plotn <- NULL

# calculate euclidean distance for all predictors except for geographic distances
foreuclid <- names(forpredData)[!names(forpredData) %in% c("Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")]
for(i in foreuclid){
  data <- as.matrix(vegan::vegdist(forpredData[, i, drop=F], method = "euclid"))
  data[!lower.tri(data)] <- NA
  data <- reshape2::melt(data, value.name=i)
  fordistpreds <- merge(fordistpreds, data, value.name = i, by = c("Var1", "Var2"))
}
# take out variables for which euclidean distance was calculated, such that only geographic distance stays
forpredData <- forpredData[, !names(forpredData) %in% foreuclid]

# calculate geographic distance
for(i in 1:length(rownames(fordistpreds))){
  new <- geosphere::distGeo(forpredData[fordistpreds[i, 1], c("Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")], forpredData[fordistpreds[i, 2], c("Longitude_Dec_Plotcenter", "Latitude_Dec_Plotcenter")])
  fordistpreds[i, "geo_dist"] <- new
}

# standardization of non-betadiversities to be between 0 and 1
# apply(fordistpreds, 2, range)
forstand <- c(foreuclid,"geo_dist")
fordistpreds[, names(fordistpreds) %in% forstand] <- apply(fordistpreds[, names(fordistpreds) %in% forstand], 2, scale01)

rm(foreuclid); rm(forstand); rm(i)
```

# add LUI values
add the LUI value selected by `xaxis_gdm` as geographic coordinate to the gdm data.
```{r}
setnames(xLUI, old = c("Plotn", xaxis_gdm), new = c("Var1", "s1.xCoord"))
fordistpreds <- merge(fordistpreds, xLUI, by = "Var1", all.x = T)
setnames(xLUI, old = c("Var1", "s1.xCoord"), new = c("Var2", "s2.xCoord"))
fordistpreds <- merge(fordistpreds, xLUI, by = "Var2", all.x = T)
```



# Check Correlation
```{r, eval = T}
# name : modelname "_corrplot_before_gdm.pdf"
m <- cor(fordistpreds[, -c(1:2, 36)])
pdf(paste("out/", modelname, "_corrplot_before_gdm.pdf", sep = ""), paper = "a4r")
corrplot::corrplot(m, type = "lower", addCoef.col = "black", method = "color", diag = F, tl.srt = 50, tl.col = "black", order = "hclust", tl.cex = 0.4, cl.cex = 0.3, number.cex = 0.3)
dev.off()
```
Variance inflation
```{r}
# generate model
mod <- paste(funs, " ~ " ,paste(colnames(fordistpreds)[!colnames(fordistpreds) %in% c(funs, "Var1", "Var2")], collapse = " + "))
m1 <- lm(formula = formula(mod), data = fordistpreds)
vif_output <- car::vif(m1)
vif_output[which(vif_output > 5)]
# vif_output[which(vif_output[, "GVIF"] > 5),]
capture.output(mod,
               vif_output[which(vif_output > 5)],
               file = paste("out/", modelname, "_vif_before_gdm.txt", sep = ""))
```
We will leave the inflated variables in the dataset, but perform a sensitivity analysis later. --> The inflated variables are not distinguishable, maybe the effect is partitioned among them.


# Prepare GDM Table
GDM requires a well-defined table. Below, the names of the columns and their content are listed:
- distance : distance between ecosystem functions
- weights : column containing all 1 (because we weigh all comparisons the same)
- s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord : 4 columns which will be ignored as I will set geo = F in the gdm() function. 
- s1.pred1, ..., s1.predn : here I fill in my predictor columns, (all have s1. at the beginning)
- s2.pred1, ..., s2.predn : here I fill in columns of zero to imitate I would not have given distance but two values.

```{r}
meltset <- data.table::as.data.table(fordistpreds)
# delete first 2 columns as not used any more
plotorder <- meltset[,.(Var1, Var2)] # backup for plots, in case I need it again
meltset[,c("Var1","Var2") := NULL]

# rearrange column order
colorder <- funs
colorder1 <- colnames(meltset)[!colnames(meltset) %in% colorder]
data.table::setcolorder(meltset, colorder)

# define the new colnames
names(meltset)[names(meltset) %in% funs] = "distance" # functions column is named "distance"
# all predictors are having s1. at the beginning
newnames <- paste("s1.", colorder1[!colorder1 %in% c("s1.xCoord", "s2.xCoord")], sep = "")
names(meltset)[!names(meltset) %in% c("distance", "s1.xCoord", "s2.xCoord")] <- newnames

# adding new columns at the right positions
newcols <- c("weights", "s1.yCoord", "s2.yCoord")
meltset[, (newcols) := 0]
meltset[, weights := 1]

# order in right position
colorder <- c("distance", newcols[1], "s1.xCoord", newcols[2], "s2.xCoord", newcols[3], newnames)
data.table::setcolorder(meltset, colorder)

# add empty s2. columns at the end
newcols2 <- gsub("1", "2", newnames)
meltset[, (newcols2) := 0]

# gdm wants meltset to be a data.frame
meltset <- as.data.frame(meltset)

# row names of meltset must be the plot combinations
rownames(meltset) <- paste(plotorder$Var1, plotorder$Var2, sep=".")

# formatsitepair has been used, but does not do anything to the data if runned like this
meltset <- gdm::formatsitepair(bioData=meltset, bioFormat = 4, predData=meltset) 
```

# Run GDM
The model name is automatically generated from `funs` and `lui`.
```{r}
outname <- paste("gdm", funs, xaxis_gdm, "output_summary.txt", sep = "_")
outnameRDS <- paste("gdm", funs, xaxis_gdm, "output.Rds", sep = "_")
gdm_output <- gdm::gdm(meltset, geo = T)
# summary(gdm_output)

# save input and output
saveRDS(meltset, file = paste("out/", paste("gdm", funs, xaxis_gdm, "input.Rds", sep = "_"), sep = ""))
saveRDS(gdm_output, paste("out/", outnameRDS, sep = ""))
capture.output(summary(gdm_output), file = paste("out/", outname, sep = ""))
```

go to cluster scripts and plot.