---
title: "NEW threshold analysis"
author: 'Noëlle Schenk'
date: "5/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# single function effects along thresholds

Questions : 
- which threshold to include?
    - functions levels are not normally distributed. Therefore, when putting a threshold, the amount of plots passing the threshold varies per function. E.g. at threshold 0.6, most plots have high levels of N leaching but only few plots have high levels of herbivory.
        - --> the multifuncitonality at threshold 0.6 is probably driven by the functions with intermediate number of plots that pass the threshold. 
     - possibly, there is a threshold where no function has a high effect - which is too "full of functions" / "lacking any functions" that anything can have an effect. We can imagine such a situation at an extremely high threshold like e.g. 0.99 : barely any plot would pass it --> so few variance that we can not explain multifuncitonality with it --> should not be included because it would only add noise.

- is beta-multifunctionality driven by the effect of only few functions, or the complementary effect of many functions?
    - with this analysis, we can answer that at different thresholds and maybe in general



## Analysis plan

Correlate the single functions (binary, presence-absence) and multifunctionality.
--> logistic regression-type correlation needed
- **multiple model** like : multifun ~ f1{0, 1} + f2{0, 1} + f3{0, 1} + ... + f16{0, 1}
    - pro : do everythin in one go
        - functions are corrected for each other, so the correlated ones will be weaker than with just pairwise
            - the group of functions with many plots passing the threshold are correlated. (also the group with barley any plot passing the threshold) --> will they have very weak effects because they will be corrected for each other?
        - tell us which functions drive multifun for each threshold and help us interpret the results ad kind of "replace" the single functions analysis

    - gives one value per function per threshold (all plots versus all plots)

- alternative : **single correlations**, each function individually with beta-multifun at each threshold.
  - pro : easier, could use Pearson or Spearman
  - contra : more work, functions are not corrected for each other

- chosing a multifunctionality : 
    - EFturnover, because 1. we need a multifunctionality which was calculated on the thresholds and 2. because nestedness is less releveant


# Analysis
Read data

need : 
- single functions levels
    - apply threshold
      create dataset of sigle functions ? (Intermediate step in calc_multifun --> save it)
- EFmaster for EFturnover at different thresholds
```{r}
t01 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.1.Rds", sep = ""))
t02 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.2.Rds", sep = ""))
t03 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.3.Rds", sep = ""))
t04 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.4.Rds", sep = ""))
t05 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.5.Rds", sep = ""))
t06 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.6.Rds", sep = ""))
t07 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.7.Rds", sep = ""))
t08 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.8.Rds", sep = ""))
t09 <- readRDS(paste(pathtodata, "/data_assembly/output_data/single_functions_passing_threshold_0.9.Rds", sep = ""))
```

- merge each treshold with EFmaster
    - loop through all thresholds I guess, start with t06
- apply model
- save vector result to result data.table
- do with other thresholds --> add result vector to result data.table
- heatmap of result data.table


