---
title: "Conductor script"
author: "N. Schenk"
date: "2022-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Aim
The analysis consists of various small (drone) scripts. The order is indicated, but instead of calling them by hand, giving the model names as inputs, this script does this job for you, as the hive does the coordination job for the drones. ([source](https://www.theserverside.com/opinion/Master-slave-terminology-alternatives-you-can-use-right-now) for hive-drone terminology).

The conductor script here calls the drone scripts of the analysis in order to run the full analysis from one place.

# Requirements

The script `vignettes/analysis_nonpublic.R` loads the necessary requirements, specifying the specific requirements by variable `sections_to_be_loaded`. Each section includes an option to load the requirements.

Requires access to the Plant Ecology Group (Fischer) drive "planteco" at the IPS Bern. See header of the `analysis_nonpublic.R` file.


# Sequence of analysis


## Prepare data
```{r}
#TODO add early scripts
```


## Run modelwise
Scripts depending on output of each other, need to run per model.

Requirements
```{r}
if(!exists("model_names")){
  sections_to_be_loaded <- c()
  source("vignettes/analysis_nonpublic.R")
}
```


Chose model by hand
```{r}
# select a given model
model_names_selection <- model_names[which(model_names$modelname == "gdm_EFturnover_0.7_LUI"), ]
#
# run analysis and create results
ksource("vignettes/prepare_and_run_GDM.Rmd") # default model : EF turnover 0.7 LUI
ksource("vignettes/check_GDM_input.Rmd") # produce nice plots/ results about correlations among input variables
ksource("vignettes/plot_modelwise_GDM.Rmd") # produce bar and line plots of the current model
```

Loop through all models. Dont forget to transfer the output from vignettes/out to `pathtodata`.
```{r, eval = F}
source("vignettes/analysis_nonpublic.R")

# run all single function and multifunctionality models
for(mod in model_names[model_class %in% c("multifun", "singlefun") & lui == "LUI", modelname]){
  model_names_selection <- model_names[which(model_names$modelname == mod), ]
  # run analysis and create results
  ksource("vignettes/prepare_and_run_GDM.Rmd") 
  ksource("vignettes/check_GDM_input.Rmd") # produce nice plots/ results about correlations among input variables
  ksource("vignettes/plot_modelwise_GDM.Rmd") # produce bar and line plots of the current model
}
```


## Run unique
Note : some scripts only have to be runned once, after the above scripts have been runned for all models. 
```{r}
ksource("vignettes/check_GDM_diagnostics.Rmd") # 1 script for all model diagnostics
ksource("vignettes/plot_unique_GDM.Rmd") # plot single EF models
ksource("vignettes/summarise_GDM_results.Rmd") # create summary tables of all model results (deviance explained and isplines)
```

## Threshold overview


Run the thresholds part of `cluster_scripts_documentation.Rmd` on the cluster..
```{r}
# ksource("cluster_scripts_documentation.Rmd") # Note : this script is not runned here, but on a cluster.
#    in case no cluster is available, run here, but be aware the computation time is high.
#TODO threshold stuff

#TODO probably add script : GDM_multifun_thresholds.Rmd
```

